### 2019-10-23 paper/paper.Rmd
first

---
title             : "The ``Standard Model'' of Early Language Learning"
shorttitle        : "Standard Model of Language Learning"

author: 
  - name          : "George Kachergis"
    affiliation   : "1"
    corresponding : no 
    address       : "420 Serra Mall"
    email         : "kachergis@stanford.edu"
  - name          : "Michael C. Frank"
    affiliation   : "1"
    corresponding : yes 
    address       : "420 Serra Mall"
    email         : "mcfrank@stanford.edu"

affiliation:
  - id            : "1"
    institution   : "Stanford University"

authornote: |
  Department of Psychology

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

In physics, the "Standard Model" is a widely-accepted theory completed in the 1970s that describes all known elementary particles, along with three of the four known fundamental forces in the universe, explaining a wide variety of empirical phenomena. 
What fundamental building blocks and mechanisms are needed to construct an analogous standard model of early language learning?

Understanding children's language environment has been of central interest to both theoretical and applied and research on early language learning.
In particular, word learning has come to be viewed as central to measuring and influencing language learning.
Important applied area, interest in sociodemographic variation

Word learning is one of the rare areas in developmental work in which the construct of individual differences is also seen as an epistemic construct:

# The Standard Model

To build a "standard model", we begin by specifying the minimal assumptions made by any model of word learning.
First, assume that words accumulate through repeated exposure at a rate controlled by a variety of factors, including 1) learning rate, 2) frequency of exposure, 3) diversity of exposure, and 4) leveraged learning.

Combine to create a "macro-economics of early language" 
Input distributional patterns that match corpus dynamics (Montag etc)
Identify critical uptake patterns to aim for (CDI quartiles?)

## Extensions
ambiguity and disambiguation
Cross-situational extensions (Smith, Blythe) - most word learning situations are ambiguous and require disambiguation across time (using mechanisms of learning and memory)
Socio-pragmatic extensions - most word learning instances are disambiguated by social and pragmatic cues (and are unambiguous)
Experimental evidence from the “associations vs. hypotheses” debate

# Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2019-10-24 paper/paper.Rmd
working on app

@@
 Socio-pragmatic extensions - most word learning instances are disambiguated by social and pragmatic cues (and are unambiguous)
-Experimental evidence from the “associations vs. hypotheses” debate
+Experimental evidence from the "associations vs. hypotheses" debate
 

### 2019-10-24 app.R

@@ -11,39 +11,74 @@ library(shiny)
 
 # Define UI for application that draws a histogram
 ui <- fluidPage(
-
-    # Application title
-    titlePanel("Old Faithful Geyser Data"),
+    titlePanel("Standard Model of Early Word Learning"),
 
     # Sidebar with a slider input for number of bins 
     sidebarLayout(
         sidebarPanel(
-            sliderInput("bins",
-                        "Number of bins:",
-                        min = 1,
-                        max = 50,
-                        value = 30)
+            selectInput("distro", "Word difficulty distribution:",
+                        list("Zipfian" = "zipf", 
+                             "Gaussian" = "gaussian", 
+                             "Uniform" = "uniform")),
+            sliderInput("vocab_size", "Vocabulary size:", 
+                        min=100, max=10000, value=1000, step=100),
+            sliderInput("n_learners", "Number of learners:", 
+                        min=1, max=1000, value=10),
+            sliderInput("input_rate", "Input rate (words/day):", 
+                        min=0, max=10000, value=1000),
+            sliderInput("ELI_thresh", "Mean effective learning instances needed to learn:", 
+                        min=1, max=100, value=10),
+            sliderInput("max_age", "Age range (months):", 
+                        min=0, max=48, value=12),
+            sliderInput("proc_speed", "Mean processing speed:", 
+                        min = 0, max = 1, value = 0.5, step= 0.1),
+            checkboxInput("proc_facilitates", "Processing facilitates acquisition", FALSE)
         ),
 
         # Show a plot of the generated distribution
         mainPanel(
-           plotOutput("distPlot")
+           plotOutput("ageVocab")
         )
     )
 )
 
-# Define server logic required to draw a histogram
+# need to add proc_speed term and proc_facilitates interaction term
+simulate <- function(vocab_size, distro, input_rate, n_learners, ELI_thresh, max_age) {
+    if(distro=="zipf") {
+        probs = 1:vocab_size / sum(1:vocab_size)
+    } else if(distro=="uniform") {
+        probs = rep(1/vocab_size, vocab_size)
+    }
+    
+    word_occs = matrix(0, nrow=n_learners, ncol=vocab_size) # number of times each word has appeared per learner
+    known_words = matrix(0, nrow=n_learners, ncol=max_age) # 
+    
[…]

### 2019-10-25 app.R
first version - no distributions

@@ -2,34 +2,31 @@
 # This is a Shiny web application. You can run the application by clicking
 # the 'Run App' button above.
 #
-# Find out more about building applications with Shiny here:
-#
-#    http://shiny.rstudio.com/
-#
 
 library(shiny)
+require(tidyverse)
 
-# Define UI for application that draws a histogram
+# Define UI
 ui <- fluidPage(
     titlePanel("Standard Model of Early Word Learning"),
 
-    # Sidebar with a slider input for number of bins 
     sidebarLayout(
         sidebarPanel(
             selectInput("distro", "Word difficulty distribution:",
                         list("Zipfian" = "zipf", 
-                             "Gaussian" = "gaussian", 
                              "Uniform" = "uniform")),
             sliderInput("vocab_size", "Vocabulary size:", 
-                        min=100, max=10000, value=1000, step=100),
+                        min=100, max=10000, value=2000, step=200),
             sliderInput("n_learners", "Number of learners:", 
-                        min=1, max=1000, value=10),
+                        min=1, max=1000, value=100, step=50),
             sliderInput("input_rate", "Input rate (words/day):", 
-                        min=0, max=10000, value=1000),
-            sliderInput("ELI_thresh", "Mean effective learning instances needed to learn:", 
-                        min=1, max=100, value=10),
+                        min=0, max=10000, value=1000, step=100),
+            sliderInput("threshold", "Number of occurrences needed to learn a word:", 
+                        min=0, max=1000, value=100, step=10),
             sliderInput("max_age", "Age range (months):", 
-                        min=0, max=48, value=12),
+                        min=0, max=48, value=24, step=3),
+            sliderInput("learning_rate", "Mean learning rate:", 
+                        min = 0, max = 1, value = 0.5, step= 0.1),
             sliderInput("proc_speed", "Mean processing speed:", 
                         min = 0, max = 1, value = 0.5, step= 0.1),
             checkboxInput("proc_facilitates", "Processing facilitates acquisition", FALSE)
@@ -43,15 +40,17 @@ ui <- fluidPage(
 )
 
 # need to add proc_speed term and proc_facilitates interaction term
-simulate <- function(vocab_size, distro, input_rate, n_learners, ELI_thresh, max_age) {
[…]

### 2019-10-26 app.R
add individual learning rate variability

@@ -5,6 +5,7 @@
 
 library(shiny)
 require(tidyverse)
+library(quantreg)
 
 # Define UI
 ui <- fluidPage(
@@ -12,13 +13,13 @@ ui <- fluidPage(
 
     sidebarLayout(
         sidebarPanel(
-            selectInput("distro", "Word difficulty distribution:",
+            selectInput("distro", "Word frequency distribution:",
                         list("Zipfian" = "zipf", 
                              "Uniform" = "uniform")),
             sliderInput("vocab_size", "Vocabulary size:", 
-                        min=100, max=10000, value=2000, step=200),
+                        min=200, max=10000, value=5000, step=200),
             sliderInput("n_learners", "Number of learners:", 
-                        min=1, max=1000, value=100, step=50),
+                        min=10, max=500, value=100, step=10),
             sliderInput("input_rate", "Input rate (words/day):", 
                         min=0, max=10000, value=1000, step=100),
             sliderInput("threshold", "Number of occurrences needed to learn a word:", 
@@ -26,7 +27,7 @@ ui <- fluidPage(
             sliderInput("max_age", "Age range (months):", 
                         min=0, max=48, value=24, step=3),
             sliderInput("learning_rate", "Mean learning rate:", 
-                        min = 0, max = 1, value = 0.5, step= 0.1),
+                        min = 1, max = 10, value = 1, step= 1),
             sliderInput("proc_speed", "Mean processing speed:", 
                         min = 0, max = 1, value = 0.5, step= 0.1),
             checkboxInput("proc_facilitates", "Processing facilitates acquisition", FALSE)
@@ -39,10 +40,14 @@ ui <- fluidPage(
     )
 )
 
+sigmoid <- function(x) {
+    return( 1 / (1 + exp(-x)) )
+}
+
 # need to add proc_speed term and proc_facilitates interaction term
-#simulate(1000, "zipf", 100, 10, 15, 12)
-simulate <- function(vocab_size, distro, input_rate, n_learners, threshold, max_age) {
-    # could add learning_rate = rnorm(n_learners, mean_learning_rate) # individual learning rates
+simulate <- function(vocab_size, distro, input_rate, n_learners, threshold, max_age, mean_learning_rate) {
+    learning_rate = rnorm(n_learners, mean=mean_learning_rate, sd=1) # individual learning rates
[…]

### 2019-10-26 app.R
minor changes

@@ -17,15 +17,15 @@ ui <- fluidPage(
                         list("Zipfian" = "zipf", 
                              "Uniform" = "uniform")),
             sliderInput("vocab_size", "Vocabulary size:", 
-                        min=200, max=10000, value=5000, step=200),
+                        min=500, max=10000, value=5000, step=500),
             sliderInput("n_learners", "Number of learners:", 
                         min=10, max=500, value=100, step=10),
             sliderInput("input_rate", "Input rate (words/day):", 
-                        min=0, max=10000, value=1000, step=100),
+                        min=100, max=10000, value=1000, step=100),
             sliderInput("threshold", "Number of occurrences needed to learn a word:", 
                         min=0, max=1000, value=100, step=10),
             sliderInput("max_age", "Age range (months):", 
-                        min=0, max=48, value=24, step=3),
+                        min=0, max=48, value=36, step=3),
             sliderInput("learning_rate", "Mean learning rate:", 
                         min = 1, max = 10, value = 1, step= 1),
             sliderInput("proc_speed", "Mean processing speed:", 
@@ -72,7 +72,7 @@ simulate <- function(vocab_size, distro, input_rate, n_learners, threshold, max_
     }
     # return per individual per month 
     # reshape to long
-    known_words = cbind(rep(0,s), known_words)
+    known_words = cbind(rep(0,n_learners), known_words)
     known_words = data.frame(known_words)
     names(known_words) = 0:max_age
     known_words$id = 1:nrow(known_words)
@@ -90,7 +90,7 @@ server <- function(input, output) {
         #sim = simulate(5000, "uniform", 500, 10, 15, 12, 1)
         gd <- sim %>% group_by(month, id) %>% summarise(mean=mean(words), sd=sd(words))
         qs <- c(0.10,0.25,0.50,0.75,0.90)
-        ggplot(sim, aes(x=month, y=words)) + geom_jitter(width=.1, alpha=.2) + geom_point(alpha=.2) + # position="jitter", 
+        ggplot(sim, aes(x=month, y=words)) + geom_jitter(width=.1, alpha=.2) + geom_point(alpha=.2) +  
             geom_quantile(quantiles=qs, formula=y ~ poly(x, 2), aes(colour = as.factor(..quantile..))) + 
             labs(colour="Quantile") +
             geom_abline(intercept=0, slope=input$vocab_size/input$max_age, linetype="dashed", color="grey", size=1) + 

### 2019-10-26 app.R
tweaks

@@ -16,17 +16,17 @@ ui <- fluidPage(
             selectInput("distro", "Word frequency distribution:",
                         list("Zipfian" = "zipf", 
                              "Uniform" = "uniform")),
-            sliderInput("vocab_size", "Vocabulary size:", 
+            sliderInput("vocab_size", "Total vocabulary size:", 
                         min=500, max=10000, value=5000, step=500),
             sliderInput("n_learners", "Number of learners:", 
-                        min=10, max=500, value=100, step=10),
-            sliderInput("input_rate", "Input rate (words/day):", 
-                        min=100, max=10000, value=1000, step=100),
+                        min=10, max=200, value=50, step=10),
+            sliderInput("input_rate", "Input rate (tokens/day):", # HR welfare: 616, prof: 2153
+                        min=100, max=6000, value=1000, step=100),
             sliderInput("threshold", "Number of occurrences needed to learn a word:", 
                         min=0, max=1000, value=100, step=10),
             sliderInput("max_age", "Age range (months):", 
                         min=0, max=48, value=36, step=3),
-            sliderInput("learning_rate", "Mean learning rate:", 
+            sliderInput("learning_rate", "Mean learning rate (scales value of occurrence; truncated at .1):", 
                         min = 1, max = 10, value = 1, step= 1),
             sliderInput("proc_speed", "Mean processing speed:", 
                         min = 0, max = 1, value = 0.5, step= 0.1),
@@ -47,7 +47,7 @@ sigmoid <- function(x) {
 # need to add proc_speed term and proc_facilitates interaction term
 simulate <- function(vocab_size, distro, input_rate, n_learners, threshold, max_age, mean_learning_rate) {
     learning_rate = rnorm(n_learners, mean=mean_learning_rate, sd=1) # individual learning rates
-    learning_rate[which(learning_rate<0)] = 0.00001 # >0
+    learning_rate[which(learning_rate<0)] = 0.1 # >0
     if(distro=="zipf") {
         probs = 1:vocab_size / sum(1:vocab_size)
     } else if(distro=="uniform") {
@@ -90,9 +90,9 @@ server <- function(input, output) {
         #sim = simulate(5000, "uniform", 500, 10, 15, 12, 1)
         gd <- sim %>% group_by(month, id) %>% summarise(mean=mean(words), sd=sd(words))
[…]

### 2020-02-14 app.R
no more learning rate -- only processing speed

@@ -3,7 +3,7 @@ library(shiny)
 library(tidyverse)
 library(quantreg)
 library(DT)
-
+library(shinyWidgets)
 
 source("model-nonsampling.R")
 
@@ -18,26 +18,37 @@ ui <- fluidPage(
                         list("Zipfian" = "zipf",
                              "log(Zipfian)" = "logzipf",
                              "Uniform" = "uniform")),
-            sliderInput("input_rate", "Input rate (tokens/hour):", 
+            chooseSliderSkin("Modern"),
+            setSliderColor(c("DarkSlateGrey", "DarkSlateGrey", # Input
+                             "DeepSkyBlue", "DeepSkyBlue", # Word Threshold
+                             "DeepPink", "DeepPink", # Proc Speed Adult Asymptote
+                             "DarkRed", "DarkRed"), # Proc Speed Rate of Development
+                           c(1,2, 3,4, 5,6, 7,8)), # color code param means and SDs
+            sliderInput("input_rate", "Input rate mean (tokens/hour):", 
                         min=100, max=6000, value=1000, step=100), 
-            sliderInput("input_rate_sd", "Standard deviation of input rate:", 
+            sliderInput("input_rate_sd", "Input rate standard deviation (SD):", 
                         min=0, max=1000, value=100, step=10), 
-            helpText("e.g., Hart & Risley low SES: 616/hr; high SES: 2153/hr, and let's assume 12 waking hours per day"),
-            sliderInput("threshold", "Mean occurrences needed to learn a word:", 
+            helpText("e.g., Hart & Risley low SES: 616/hr; high SES: 2153/hr; we assume 12 waking hours/day"),
+            
+            sliderInput("threshold", "Threshold mean (occurrences needed to learn a word):", 
                         min=100, max=6000, value=1000, step=100),
-            sliderInput("threshold_sd", "Standard deviation of threshold distribution:", 
+            sliderInput("threshold_sd", "Threshold standard deviation:", 
                         min=0, max=1000, value=300, step=20),
             helpText("McMurray (2007) used a mean of 4000 and a large SD."),
+            
             #sliderInput("learning_rate", "Mean learning rate (scales value of occurrence; truncated at .1):", 
             #            min = .5, max = 10, value = 1, step=.5),
-            sliderInput("learning_rate", "Mean adult processing speed asymptote (a; scales value of occurrence; truncated at .01):", 
+            
[…]

### 2020-02-14 grid-check-acceleration.R

@@ -8,22 +8,40 @@
 #  mean_learning_rate, threshold_sd, proc_facilitates, proc_speed_dev)
 source("model-nonsampling.R")
 
-start_t = Sys.time()
-dat = simulate(vocab_size=10000, distro="zipf", 1000, n_learners=100, 500, max_age=24, 1, .5, 10, F, .72)
-stop_t = Sys.time()
-stop_t - start_t # 1.2s
+ex_parms = list(distro="uniform",
+                input_rate = 1000,
+                input_rate_sd = 100,
+                threshold = 1000,
+                threshold_sd = 100,
+                mean_learning_rate = .5, # proc_speed_asymp
+                learning_rate_sd = .1,
+                proc_facilitates = T,
+                proc_speed_dev = .72, 
+                proc_speed_dev_sd = .1
+              )
 
-input_rates = c(500, 1000, 1500, 2000, 2500, 3000)
-thresholds = c(500, 1000, 1500, 2000, 2500, 3000, 4000)
-threshold_sds = c(0, 10, 100, 200, 300)
-learning_rates = c(.5, 1, 2, 3)
-learning_rate_sds = c(0, .5, 1, 2, 3)
+start_t = Sys.time()
+dat = simulate(ex_parms)
+stop_t = Sys.time()
+stop_t - start_t # 1.1s
+
+input_rates = seq(500, 3000, 500)
+thresholds = seq(500, 6000, 1000)
+threshold_sds = c(0, 100)
+#learning_rates = c(.5, 1, 2, 3)
+#learning_rate_sds = c(0, .5, 1, 2, 3)
+proc_speed_asymps = c(.3, .6, .9) # adult = .56
+proc_speed_asymp_sds = c(.1)
 # distro="uniform" / "zipf"
 # processing_facilitates T / F (if T: 
-proc_speeds = seq(.1,1,.2)
+proc_speed_devs = c(.4, .7, 1) # .72
+proc_speed_dev_sds = c(.1)
 
-length(input_rates)*length(thresholds)*length(learning_rates)*length(threshold_sds)*length(learning_rate_sds)
-# 4200 * 2 * 2 = 16800 * 1.2 / 60 = 336
+length(input_rates)*
+  length(thresholds)*length(threshold_sds)*
+  length(proc_speed_asymps)*length(proc_speed_asymp_sds)*
+  length(proc_speed_devs)*length(proc_speed_dev_sds)
[…]

### 2020-02-14 model-nonsampling.R

@@ -38,6 +38,31 @@ make_long_df <- function(df, n_learners, max_age) {
 #  max_age=48, mean_learning_rate, learning_rate_sd, threshold_sd, 
 #  proc_facilitates, proc_speed_dev, proc_speed_dev_sd) {
 
+get_proc_speed <- function(print_plot=F) {
+  a = rnorm(n_learners, mean=.56, sd=.1) # individual adult asymptotes for proc speed
+  c = rnorm(n_learners, mean=.72, sd=.1)
+  b = 1.04
+  proc_speed = matrix(0, nrow=n_learners, ncol=max_age)
+  #proc_speed = apply(get_proc_speed)
+  for(t in 1:max_age) {
+    proc_speed[,t] = a + b*exp(-c*log(t+1)) #/ vocab_size 
+  }
+
+  if(print_plot) {
+    proc_speed_long = make_long_df(proc_speed, n_learners, max_age)
+    gg <- ggplot(proc_speed_long, aes(x=month, y=words)) + 
+      geom_line(aes(group=id), alpha=.1) +  
+      labs(colour="Quantile") + geom_smooth() + 
+      xlab("Age (months)") + ylab("Response Time (seconds)") 
+    print(gg)
+  }
+  
+  return(proc_speed)
+}
+
+#get_proc_speed(print_plot=T)
+
+
 # need to add proc_speed term and proc_facilitates interaction term
 simulate <- function(parms) {
   #probs = wf_distros[[distro]] # "zipf" or "uniform"
@@ -72,7 +97,7 @@ simulate <- function(parms) {
   proc_speed = matrix(0, nrow=n_learners, ncol=max_age) # do we want per word instead of per learner?
   
   # processing speed - each of these parameters could have individual variability
-  # Kail (1991): RT slopes follow an exponential, such that Y(i) = a + b*exp(−c*i), where Y=predicted var, i=age. 
+  # Kail (1991): RT slopes follow an exponential, such that Y(i) = a + b*exp(−c*i), where Y=predicted var, i=age (mos? year?). 
   #a = 0.56 # eventual (adult) asymptote VARIABILITY HERE
   b = 1.04 # multiplier for the (infant) intercept - SES or preemie VARIABILITY HERE
   #c = proc_speed_dev #rate of development (0.72 in Frank, Lewis, & MacDonald, 2016)
@@ -87,7 +112,8 @@ simulate <- function(parms) {
[…]

### 2020-02-19 paper/paper.Rmd
no more learning rate -- only processing speed
model fits to wordbank data

@@
   One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-
   Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
   One sentence clearly stating the **general problem** being addressed by  this particular study.
-
   One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
   Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
   One or two sentences to put the results into a more **general context**.
-
   Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
@@
 Input distributional patterns that match corpus dynamics (Montag etc)
-Identify critical uptake patterns to aim for (CDI quartiles?)
+Identify critical uptake patterns to aim for (CDI quantiles?)
 

### 2020-02-19 app.R

@@ -108,7 +108,7 @@ server <- function(input, output) {
             labs(colour="Quantile") + geom_smooth() + 
             #geom_abline(intercept=0, slope=input$vocab_size/input$max_age, linetype="dashed", color="grey", size=1) + 
             xlab("Age (months)") + ylab("Response Time (seconds)") + 
-            xlim(1, max_age) + ylim(0,1.5)
+            xlim(1, max_age) + ylim(0,1.7)
         #print(sim_data()$proc_speed)
     })
     

### 2020-02-19 fit_quantiles.R

@@ -1,0 +1,73 @@
+# fit wordbank quantile data
+require(tidyverse)
+require(DEoptim)
+
+source("model-nonsampling.R")
+load("wordbank_quantiles.Rdata") # quantiles from wordbank EN WG+WS productive vocab
+
+max_age = 30 # months
+n_learners = 500
+
+qs <- c(0.10,0.25,0.50,0.75,0.90)
+
+ex_parms = list(distro="uniform",
+                input_rate = 1000,
+                input_rate_sd = 100,
+                threshold = 1000,
+                threshold_sd = 100,
+                mean_learning_rate = .5, # proc_speed_asymp
+                learning_rate_sd = .1,
+                proc_facilitates = T,
+                proc_speed_dev = .72, 
+                proc_speed_dev_sd = .1
+)
+
+
+fitSSE <- function(parms, proc_facilitates=T, graph=F) {
+  full_parms = list(distro="logzipf",
+                  input_rate = parms[1],
+                  input_rate_sd = parms[2],
+                  threshold = parms[3],
+                  threshold_sd = parms[4],
+                  mean_learning_rate = parms[5], # proc_speed_asymp
+                  learning_rate_sd = parms[6],
+                  proc_facilitates = proc_facilitates,
+                  proc_speed_dev = parms[7], 
+                  proc_speed_dev_sd = parms[8]
+  )
+  simdat = simulate(full_parms)$known_words
+  sim_quants <- simdat %>%
+    group_by(month) %>%
+    group_modify(~{
+      quantile(.x$words, probs=qs) %>%
+        tibble::enframe()
+    }) %>% spread(name, value)
+  
+  if(graph) {
+    theme_set(theme_classic())
+    ggplot(simdat, aes(x=month, y=words)) + geom_line(aes(group=id), alpha=.05) +  
+      geom_quantile(quantiles=qs, formula=y ~ poly(x, 2), aes(colour = as.factor(..quantile..))) + 
+      labs(colour="Quantile") + xlab("Age (months)") + ylab("Vocabulary Size") + xlim(1, max_age) 
+    ggsave("model_quantiles.pdf", width=6, height=5)
+  }
+  
+  return(sum((sim_quants[8:max_age,] - quants)^2))
+}
+
[…]

### 2020-02-19 grid-check-acceleration.R

@@ -25,23 +25,23 @@ dat = simulate(ex_parms)
 stop_t = Sys.time()
 stop_t - start_t # 1.1s
 
-input_rates = seq(500, 3000, 500)
-thresholds = seq(500, 6000, 1000)
+input_rates = seq(500, 4000, 500)
+thresholds = seq(500, 7000, 1000)
 threshold_sds = c(0, 100)
 #learning_rates = c(.5, 1, 2, 3)
 #learning_rate_sds = c(0, .5, 1, 2, 3)
-proc_speed_asymps = c(.3, .6, .9) # adult = .56
+proc_speed_asymps = c(.1, .3, .6, .9) # adult = .56
 proc_speed_asymp_sds = c(.1)
 # distro="uniform" / "zipf"
 # processing_facilitates T / F (if T: 
-proc_speed_devs = c(.4, .7, 1) # .72
+proc_speed_devs = c(.1, .4, .7, 1) # .72
 proc_speed_dev_sds = c(.1)
 
 length(input_rates)*
   length(thresholds)*length(threshold_sds)*
   length(proc_speed_asymps)*length(proc_speed_asymp_sds)*
   length(proc_speed_devs)*length(proc_speed_dev_sds)
-# 648 * 2 * 2 = 2592 * 1.1 / 60 = 47.5/60 = .79hrs
+# 1792 * 2 * 3 = 10752 * 1.1 / 60 = 132 mins 
 
 acceleration_test <- function(dat) {
   d = dat$known_words %>% filter(month>11) %>% group_by(month) %>% summarize(mean=mean(words)) 
@@ -57,7 +57,9 @@ acceleration_test <- function(dat) {
 
 do_grid <- function(distro, processing_facilitates) {
   dat = data.frame(distro=NA, input_rate=NA, threshold=NA, 
-                   threshold_sd=NA, learning_rate=NA, learning_rate_sd=NA, proc_facilitates=NA, proc_speed=NA, acceleration=NA)
+                   threshold_sd=NA, proc_speed_asymp=NA, 
+                   proc_speed_asymp_sd=NA, proc_facilitates=NA, 
+                   proc_speed_dev=NA, proc_speed_dev_sd=NA, acceleration=NA)
   for(i in input_rates) {
     for(t in thresholds) {
       for(lr in proc_speed_asymps) { 
@@ -78,7 +80,9 @@ do_grid <- function(distro, processing_facilitates) {
                 sim = simulate(parms)
[…]

### 2020-02-19 model-nonsampling.R

@@ -152,4 +152,4 @@ ex_parms = list(distro="uniform",
              proc_speed_dev_sd = .1
   )
 
-simulate(ex_parms)
+#simulate(ex_parms)

### 2020-02-19 fit_quantiles.R
sherlock jobs

@@ -23,7 +23,7 @@ ex_parms = list(distro="uniform",
 )
 
 
-fitSSE <- function(parms, proc_facilitates=T, graph=F) {
+fitSSE <- function(parms, proc_facilitates=T, graph="") {
   full_parms = list(distro="logzipf",
                   input_rate = parms[1],
                   input_rate_sd = parms[2],
@@ -43,31 +43,30 @@ fitSSE <- function(parms, proc_facilitates=T, graph=F) {
         tibble::enframe()
     }) %>% spread(name, value)
   
-  if(graph) {
+  if(graph!="") {
     theme_set(theme_classic())
     ggplot(simdat, aes(x=month, y=words)) + geom_line(aes(group=id), alpha=.05) +  
       geom_quantile(quantiles=qs, formula=y ~ poly(x, 2), aes(colour = as.factor(..quantile..))) + 
       labs(colour="Quantile") + xlab("Age (months)") + ylab("Vocabulary Size") + xlim(1, max_age) 
-    ggsave("model_quantiles.pdf", width=6, height=5)
+    ggsave(paste0(graph,".pdf"), width=6, height=5)
   }
   
   return(sum((sim_quants[8:max_age,] - quants)^2))
 }
 
-set.seed(1234)
-fit <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .01, .01, .01, .01), 
-               upper=c(9000, 9000, 9000, 9000, 1, 2, 1, 2), 
-        control = DEoptim.control(NP = 80, itermax = 100)) # , F = 1.2, CR = 0.7
-# 256282.537500 bestmemit:  
-pars = c(377.67, 2091.39, 8034.29, 2638.25, 0.402713, 0.015136, 0.627340, 0.109277)
-fitSSE(pars, graph=T) # SSE= 265834
-#fitSSE(fit$optim$bestmem, graph=T) 
 
-# fix the processing speed parameters
-fit2 <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .555, .01, .715, .01), 
-               upper=c(9000, 9000, 9000, 9000, .564, 2, .724, 2), 
-               control = DEoptim.control(NP = 80, itermax = 100))
+#set.seed(1234)
[…]

### 2020-02-19 fitmodel2.sbatch.sh

@@ -1,0 +1,32 @@
+#!/bin/bash
+#
+#SBATCH --job-name=standardmodel2
+#
+#SBATCH -p hns
+#SBATCH --time=10:00:00
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --mem-per-cpu=2G
+
+# load the module
+ml R
+
+R --no-save << EOF
+
+source("fit-quantiles.R")
+set.seed(1234)
+
+start_time <- Sys.time()
+
+fit <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .01, .01, .01, .01), 
+               upper=c(9000, 9000, 9000, 9000, 1, 2, 1, 2), proc_facilitates=F,
+               control = DEoptim.control(NP = 80, itermax = 200)) 
+
+end_time <- Sys.time()
+print(end_time - start_time)
+
+fitSSE(fit$optim$bestmem, graph="unconstrained_proc_facil_FALSE") 
+
+save.image(file='fit_unconstrained_proc_facil_FALSE.RData')
+save(fit, file="fit_uncon_pfF.RData")
+EOF

### 2020-02-19 fitmodel3.sbatch.sh

@@ -1,0 +1,32 @@
+#!/bin/bash
+#
+#SBATCH --job-name=standardmodel3
+#
+#SBATCH -p hns
+#SBATCH --time=10:00:00
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --mem-per-cpu=2G
+
+# load the module
+ml R
+
+R --no-save << EOF
+
+source("fit-quantiles.R")
+set.seed(1234)
+
+start_time <- Sys.time()
+
+fit <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .555, .01, .715, .01), 
+               upper=c(9000, 9000, 9000, 9000, .564, 2, .724, 2), 
+               control = DEoptim.control(NP = 80, itermax = 200))
+
+end_time <- Sys.time()
+print(end_time - start_time)
+
+fitSSE(fit$optim$bestmem, graph="constrained_proc_facil_TRUE")
+
+save.image(file='fit_constrained_proc_facil_TRUE.RData')
+save(fit, file="fit_con_pfT.RData")
+EOF

### 2020-02-19 fitmodel4.sbatch.sh

@@ -1,0 +1,32 @@
+#!/bin/bash
+#
+#SBATCH --job-name=standardmodel4
+#
+#SBATCH -p hns
+#SBATCH --time=10:00:00
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=1
+#SBATCH --mem-per-cpu=2G
+
+# load the module
+ml R
+
+R --no-save << EOF
+
+source("fit-quantiles.R")
+set.seed(1234)
+
+start_time <- Sys.time()
+
+fit <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .555, .01, .715, .01), 
+               upper=c(9000, 9000, 9000, 9000, .564, 2, .724, 2), proc_facilitates=FALSE,
+               control = DEoptim.control(NP = 80, itermax = 200))
+
+end_time <- Sys.time()
+print(end_time - start_time)
+
+fitSSE(fit$optim$bestmem, graph="constrained_proc_facil_FALSE")
+
+save.image(file='fit_constrained_proc_facil_FALSE.RData')
+save(fit, file="fit_con_pfF.RData")
+EOF

### 2020-05-29 model_interventions.Rmd
adding input data

---
title: "Modeling Interventions"
output: html_notebook
---

We'd like to be able to model interventions (both past and proposed) with the standard model. 
How would that work? Let's consider Gilkerson et al. (2017), in which 72 parents of typically developing children aged 9 to 21 months were assigned to immediate- or delayed-treatment (control) conditions. 
During treatment, parents were recorded 10 times over a 3-month period while engaging in a web-based intervention offering strategies for increasing talk and interaction. 
During a 9-month follow-up, parents completed additional recordings and language assessments. 
Overall, no differences were found in language behaviors between immediate-treatment versus delayed-treatment groups. 
However, parents who were at first below-average on automated language measures showed significant postintervention increases which held longitudinally, and children showed significant elevations in language ability.
Problem: Gilkerson 2017 only reports standardized scores (mean = 100 and SD = 15) for AWC (median split: bottom half M=84.09, upper half M=116.01), which are calculated from the original LENA norms (Gilkerson & Richards, 2008), where they report overall M=12,297 (SD=6462) per 12-hour session (so 1025 words/hour).

```{r}
require(tidyverse)
source("model-nonsampling.R")

# Gilkerson et al. 2017 Table 2: standardized AWC

# Gilkerson et al. 2018 Table 2: 
tribble(
  ~age_mos, ~AWC_mean, ~AWC_sd,
  "2-17",   14484, 6022,
  "18-24",  11768, 4159,
  "25+",   12928, 4342
) # total sample (N=146): 13218, sd=5079, range 3372-30794

# should look at Ramírez-Esparza, García-Sierra, and Kuhl (2014, 2017)
# ADS / CDS 30s LENA segments

# on average, Bergelson2018 (North American) children heard
# 11.36 min of CDS per hour (SD = 4.24)
# average of 7.3 min of ADS per hour (SD = 6.4)
# on average 65% of the speech infants heard from adults sounded child directed (SD = 0.22); varied widely, with one infant hearing only 17% CDS while another heard 100% CDS
# CDS quantity remained largely stable with child age, ADS quantity decreased with child age
# "Shneidman et al. (2013) finds 69% CDS in North American homes of 14- to 42 month olds while we find 65% on average across 3–20 months"

# Bergelson et al. 2018 Table S1: 
Bergelson_s1 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CDS_mean, ~CDS_sd, ~N, 
  "Schneidman & Goldin-Meadow 2013", 
  "Schneidman & Goldin-Meadow 2013", 
  "Weisleder & Fernald 2013", 
)

# Table S2: prev studies' LENA results, along with the 59 of 61 recordings in the current dataset that are >8hrs
Bergelson_s2 <- tribble(
  ~study, ~AWC_mean, ~AWC_sd, ~CVC_mean, ~CVC_sd, ~N,
  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, , # these are full-day recordings (12hrs)
  "Greenwood et al. 2011", 13142, 5562, 1714, 850, , # AWC range: 631-36563, CVC range: 11-5611
  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, ,
  "Zimmerman et al. 2009", 12800, 4400, NA, NA, ,
  "Bergelson et al. 2018", 16510, 8718, 1432, 764, 59,
)

Bergelson_s2 <- Bergelson_s2 %>% mutate(AWChr_mean = AWC_mean / 12,
                                        AWChr_sd = AWC_sd / 12)

# Sperry et al. Figure 1: HR = Hart & Risley 1995
# words/hour spoken by primary caregivers to children
Sperry <- tribble(
  ~group, ~CDS_mean, ~N, ~AllCaregiverPctIncr, ~AmbientEnvPctIncr, #AllCaregiver_mean, AmbientEnv_mean,
  "HR: Kansas professionals", 2153, 13, NA, NA, #NA, NA,
  "Black Belt Poor", 1838, 11, 1.58, 2.02, # NA, 3203,
  "Longwood Middle Class", 1491, 6, 1.30, 2.04, # 1777, NA
  "HR: Kansas Middle Class", 1400, 10, NA, NA,
  "Daly Park Working Class", 1351, 7, 1.17, 1.75,
  "HR: Kansas Lower Class", 1137, 13, NA, NA,
  "South Baltimore Poor", 1061, 3, 1.21, 1.54,
  "Jefferson Working Class", 1048, 15, 1.53, 3.10,
  "Kansas Welfare", 616, 6, NA, NA
)
# final columns from Table 2 let us recreate Figure 2, which breaks down into Primary Caregiver / All Caregivers / Ambient Environment
# e.g., Black Belt Poor Ambient Env = 1838 * 2.02 should be 3203 words/hr (as reported in text..) but is 3713..

Sperry <- Sperry %>% mutate(AllCaregiver = CDS_mean * AllCaregiverPctIncr,
                  AmbientEnv = CDS_mean * AmbientEnvPctIncr)
# "Finally, children in the poor Black Belt community heard an astounding 3,203 words per hour in their ambient environment, a 102% increase over the number of words addressed to them by their primary caregivers."
```

```{r}

```

### 2020-05-29 compare_fits.R

@@ -1,4 +1,5 @@
 require(tidyverse)
+source("fit_quantiles.R") # for fitSSE function, which also runs the model and generates plot
 
 fits = list.files("fits")
 fl = list()
@@ -10,6 +11,7 @@ for(f in fits) {
   f_ = str_replace(f, ".RData", "")
   fl[[ f_ ]] = fit$optim
   df = rbind(df, c(f_, fit$optim$bestval, fit$optim$bestmem))
+  fitSSE(fit$optim$bestmem, graph=paste0("figures/",f_))
 }
 
 df = na.omit(df)

### 2020-05-29 fit_quantiles.R

@@ -3,7 +3,7 @@ require(tidyverse)
 require(DEoptim)
 
 source("model-nonsampling.R")
-load("wordbank_quantiles.Rdata") # quantiles from wordbank EN WG+WS productive vocab
+load("data/wordbank_quantiles.Rdata") # quantiles from wordbank EN WG+WS productive vocab
 
 max_age = 30 # months
 n_learners = 500

### 2020-06-03 model_interventions.Rmd
use CHILDES word frequencies

@@
   ~study, ~AWC_mean, ~AWC_sd, ~CVC_mean, ~CVC_sd, ~N,
-  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, , # these are full-day recordings (12hrs)
-  "Greenwood et al. 2011", 13142, 5562, 1714, 850, , # AWC range: 631-36563, CVC range: 11-5611
-  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, ,
-  "Zimmerman et al. 2009", 12800, 4400, NA, NA, ,
+  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, 72, # these are full-day recordings (12hrs)
+  "Greenwood et al. 2011", 13142, 5562, 1714, 850, 30, # mid- to upper-SES; AWC range: 631-36563, CVC range: 11-5611
+  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, 11, # mostly shorter recordings (6-8hrs)
+  "Zimmerman et al. 2009", 12800, 4400, NA, NA, 275, # cross-sectional; only 71 longitudinal
   "Bergelson et al. 2018", 16510, 8718, 1432, 764, 59,
@@
 
+Let's use the hourly rates from individual children from Sperry et al. (2018), Hart & Risley (1995), and Soderstrom & Wittebolle (2013) to estimate the parameters of a normal distribution.
+
 ```{r}
+df <- read.csv("data/hourly_tokens_Sperry_HartRisley.csv")
+df_agg <- df %>% summarise(mean = mean(mother_child_tokens_hr),
+                           sd = sd(df$mother_child_tokens_hr))
+# mean: 1404, sd: 859
+breaks = seq(0, 3650, by = 120) 
+ggplot(df, aes(x = mother_child_tokens_hr)) + 
+    geom_histogram(aes(y =..density..), breaks = breaks, 
+                   color = "black", fill = "gray") +
+    stat_function(fun = dnorm, args = list(mean = df_agg$mean, sd = df_agg$sd)) +
+  geom_vline(xintercept=df_agg$mean, lty="dotted")
+ggsave("figures/CDS_hourly_tokens_Sperry_HartRisley.pdf", width=4.5,height=4)
 
+# but this should actually be a Poisson distribution, not normal
+ggplot(transform(data.frame(x=0:3650), y=dpois(x, df_agg$mean)), aes(x, y)) + 
+    geom_bar(stat="identity")
 ```

### 2020-06-03 app.R

@@ -4,12 +4,15 @@ library(tidyverse)
 library(quantreg)
 library(DT)
 library(shinyWidgets)
+library(shinythemes)
 
-source("model-nonsampling.R")
+#source("model-nonsampling.R")
+source("model-childesfreqs.R")
 
 theme_set(theme_classic())
 # Define UI
 ui <- fluidPage(
+    theme = shinytheme("spacelab"),
     titlePanel("Standard Model of Early Word Learning"),
 
     sidebarLayout(
@@ -29,10 +32,10 @@ ui <- fluidPage(
                         min=1, max=8, value=1, step=1), 
             # best-fitting parms for start_age=1 logzipf: c(204, 2436, 6937, 2127, 0.56, 0.03, 0.72, 0.21)
             sliderInput("input_rate", "Input rate mean (tokens/hour):", 
-                        min=100, max=2500, value=500, step=100), # 1000 is reasonable
+                        min=0, max=2500, value=1000, step=50), 
             helpText("e.g., Hart & Risley low SES: 616/hr; high SES: 2153/hr; we assume 12 waking hours/day"),
             sliderInput("input_rate_sd", "Input rate standard deviation (SD):", 
-                        min=0, max=3000, value=2000, step=100), 
+                        min=0, max=3000, value=900, step=100), 
             helpText("Gilkerson et al. 2017 daily SD range: 4,100-8,200"),
             
             sliderInput("threshold", "Threshold mean (occurrences needed to learn a word):", 
@@ -60,9 +63,24 @@ ui <- fluidPage(
         # Show a plot of the generated distribution
         mainPanel(
             tabsetPanel(type = "tabs", id="tabs", 
-                tabPanel("Vocabulary Growth by Age", plotOutput("ageVocab"), textOutput("acceleration")),
-                tabPanel("Processing Speed by Age", plotOutput("ageRT")),
-                tabPanel("Vocabulary Growth Table", textOutput("summary"), DT::dataTableOutput("mytable"))
+                tabPanel("Vocabulary Growth", 
+                         plotOutput("ageVocab"), 
+                         br(),
+                         textOutput("acceleration")
+                    ),
+                tabPanel("Processing Speed", 
+                         plotOutput("ageRT")
+                    ),
+                tabPanel("Vocabulary Growth Table", 
+                         textOutput("summary"), 
+                         br(),
+                         downloadButton("download_table", "Download Table",
+                                        class = "btn-default btn-xs"),
+                         br(),
[…]

### 2020-06-03 get_childes_freqs.R

@@ -1,0 +1,209 @@
+require(childesr)
+require(tidyverse)
+require(ggrepel)
+
+wf <- read_csv("data/childes_english.csv") %>%
+  select(-final_count, -solo_count, -mean_sent_length) %>% # 45598
+  filter(word_count>10, !is.na(word)) %>%# 27736 (21441 > twice, 14376 >5)
+  arrange(desc(word_count)) %>%
+  filter(!is.element(word, c("www", "w", "r", "e", "g", "y"))) # frequent non-words
+# 10236 words
+
+# problem: Mommy and mommy are different; should combine some of these counts
+combine_word_counts <- function(wf, sources, target) {
+  targ_idx = which(wf$word==target)
+  if(length(targ_idx)==0) {
+    print(paste("Target (",target,") not found in dataframe"))
+    return(wf)
+  }
+  for(source in sources) {
+    targ_idx = which(wf$word==target) # need to do this again in case it changes after removing source word
+    source_idx = which(wf$word==source)
+    if(length(source_idx)==0) {
+      print(paste("Source (",source,") not found in dataframe"))
+    } else {
+      print(paste("source count:",wf[source_idx,]$word_count, 
+                  "target count:",wf[targ_idx,]$word_count))
+      wf[targ_idx,]$word_count = wf[targ_idx,]$word_count + wf[source_idx,]$word_count
+      wf = wf[-source_idx,]
+    }
+  }
+  return(wf)
+}
+
+wf = combine_word_counts(wf, c("Mommy","mom","Mom","mama","momma","Mama","Momma"), "mommy") # 
+wf = combine_word_counts(wf, c("Daddy", "Dad", "dad", "dada", "papa","Papa","Dada"), "daddy")
+wf = combine_word_counts(wf, c("Grandpa","grandfather","Grandfather","granddaddy"), "grandpa")
+wf = combine_word_counts(wf, c("Grandma","grandmother","Grandmother","granny"), "grandma")
+wf = combine_word_counts(wf, "have_to", "hafta")
+wf = combine_word_counts(wf, "all_gone", "gone")
+wf = combine_word_counts(wf, "in", "inside")
+wf = combine_word_counts(wf, "kleenex", "tissue")
[…]

### 2020-06-03 model-childesfreqs.R

@@ -1,7 +1,10 @@
 # without loss of generality parameters (fixed):
 n_learners = 100
 max_age = 48 # months
-vocab_size = 10000
+
+wf = read.csv("data/childes_english_word_freq_cleaned.csv")
+
+vocab_size = nrow(wf) # 10190
 waking_hours_per_day = 12
 
 
@@ -69,9 +72,10 @@ simulate <- function(parms) {
   if(parms$distro=="uniform") {
     probs = rep(1/vocab_size, vocab_size)
   } else {
-    probs = 1 / (1:vocab_size + 2.7) # f(r) = 1 / (r+beta)^alpha, alpha=1, beta=2.7 (Mandelbrot, 1953, 1962)
-    probs = probs/sum(probs)
-    probs = sample(probs, length(probs)) 
+    #probs = 1 / (1:vocab_size + 2.7) # f(r) = 1 / (r+beta)^alpha, alpha=1, beta=2.7 (Mandelbrot, 1953, 1962)
+    #probs = probs/sum(probs)
+    #probs = sample(probs, length(probs)) 
+    probs = wf$word_count / sum(wf$word_count) # based on CHILDES WF distro
   }
   if(parms$distro=="logzipf") {
     lp = -log(probs)
@@ -93,7 +97,10 @@ simulate <- function(parms) {
   #if(threshold_varies) threshold = rnorm(vocab_size, mean=threshold, sd=20)
   threshold = rnorm(vocab_size, mean=parms$threshold, sd=parms$threshold_sd)
   threshold[which(threshold<1)] = 1 # minimum 1 occurrence to learn
+  # child x word
   cumulative_word_occs = matrix(0, nrow=n_learners, ncol=vocab_size) # number of times each word has appeared per learner
+  colnames(cumulative_word_occs) = wf$word
+  # child x age
   known_words = matrix(0, nrow=n_learners, ncol=max_age) # known words per individual (row) per month (col)
   proc_speed = matrix(0, nrow=n_learners, ncol=max_age) # do we want per word instead of per learner?
   
@@ -106,14 +113,18 @@ simulate <- function(parms) {
   # how to model clinically-important variables, e.g., they tell parents to follow-in (e.g., trains not animals for ASD)
   # can we get a true late-talker by 1) screwing around with input factors, and/or 2) screwing around with child-level variables
 
+  # tokens per month for each child
[…]

### 2020-06-03 fit_quantiles.R
new batches

@@ -2,7 +2,8 @@
 require(tidyverse)
 require(DEoptim)
 
-source("model-nonsampling.R")
+#source("model-nonsampling.R")
+source("model-childesfreqs.R")
 load("data/wordbank_quantiles.Rdata") # quantiles from wordbank EN WG+WS productive vocab
 
 max_age = 30 # months

### 2020-06-03 fitmodel2.sbatch.sh

@@ -3,7 +3,7 @@
 #SBATCH --job-name=standardmodel2
 #
 #SBATCH -p hns
-#SBATCH --time=36:00:00
+#SBATCH --time=48:00:00
 #SBATCH --ntasks=1
 #SBATCH --cpus-per-task=1
 #SBATCH --mem-per-cpu=2G
@@ -21,12 +21,11 @@ start_time <- Sys.time()
 fit <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .01, .01, .01, .01, 1), 
                upper=c(9000, 9000, 9000, 9000, 1, 2, 1, 2, 9), 
                proc_facilitates=F, distro="uniform",
-               control = DEoptim.control(NP = 120, itermax = 120)) 
+               control = DEoptim.control(NP = 140, itermax = 140)) 
 
 end_time <- Sys.time()
 print(end_time - start_time)
 
-#save.image(file='fit_unif_unconstrained_proc_facil_FALSE.RData')
-save(fit, file="fit_uncon_pfF.RData")
+save(fit, file="fits/fit_uncon_pfF.RData")
 fitSSE(fit$optim$bestmem, graph="unif_unconstrained_proc_facil_FALSE")
 EOF

### 2020-06-03 fitmodel3.sbatch.sh

@@ -3,7 +3,7 @@
 #SBATCH --job-name=standardmodel3
 #
 #SBATCH -p hns
-#SBATCH --time=20:00:00
+#SBATCH --time=48:00:00
 #SBATCH --ntasks=1
 #SBATCH --cpus-per-task=1
 #SBATCH --mem-per-cpu=2G
@@ -20,13 +20,12 @@ start_time <- Sys.time()
 
 fit <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .555, .01, .715, .01, 1), 
                upper=c(9000, 9000, 9000, 9000, .564, 2, .724, 2, 9), distro="uniform",
-               control = DEoptim.control(NP = 120, itermax = 120))
+               control = DEoptim.control(NP = 140, itermax = 140))
 
 end_time <- Sys.time()
 print(end_time - start_time)
 
 
-#save.image(file='fit_unif_constrained_proc_facil_TRUE.RData')
 save(fit, file="fit_unif_con_pfT.RData")
 fitSSE(fit$optim$bestmem, graph="unif_constrained_proc_facil_TRUE")
 EOF

### 2020-06-03 fitmodel4.sbatch.sh

@@ -3,7 +3,7 @@
 #SBATCH --job-name=standardmodel4
 #
 #SBATCH -p hns
-#SBATCH --time=20:00:00
+#SBATCH --time=48:00:00
 #SBATCH --ntasks=1
 #SBATCH --cpus-per-task=1
 #SBATCH --mem-per-cpu=2G
@@ -21,15 +21,12 @@ start_time <- Sys.time()
 fit <- DEoptim(fitSSE, lower=c(1, 1, 1, 1, .555, .01, .715, .01, 1), 
                upper=c(9000, 9000, 9000, 9000, .564, 2, .724, 2, 9), 
                proc_facilitates=FALSE, distro="uniform",
-               control = DEoptim.control(NP = 120, itermax = 120))
+               control = DEoptim.control(NP = 140, itermax = 140))
 
 end_time <- Sys.time()
 print(end_time - start_time)
 
-#fitSSE(fit$optim$bestmem, graph="unif_constrained_proc_facil_FALSE")
-
-#save.image(file='fit_unif_constrained_proc_facil_FALSE.RData')
-save(fit, file="fit_unif_con_pfF.RData")
+save(fit, file="fits/fit_unif_con_pfF.RData")
 fitSSE(fit$optim$bestmem, graph="unif_constrained_proc_facil_FALSE")
 
 EOF

### 2020-06-06 model_interventions.Rmd
integrated CHILDES word frequency distribution into model, and now plot learning of some individual words and total vs. CDI words known

--- 2020-05-29 model_interventions.Rmd
@@
 title: "Modeling Interventions"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
+
+```{r, libraries, include=F}
+require(tidyverse)
+source("model-nonsampling.R")
+```
 
@@
 
-```{r}
-require(tidyverse)
-source("model-nonsampling.R")
-
+```{r, summary-tables, echo=F}
 # Gilkerson et al. 2017 Table 2: standardized AWC
@@
   ~study, ~AWC_mean, ~AWC_sd, ~CDS_mean, ~CDS_sd, ~N, 
-  "Schneidman & Goldin-Meadow 2013", 
-  "Schneidman & Goldin-Meadow 2013", 
-  "Weisleder & Fernald 2013", 
+  "Schneidman & Goldin-Meadow 2013", 6254, NA, 2063, NA, 15,
+  "Schneidman & Goldin-Meadow 2013", NA, NA, 2404, NA, 15,
+  "Weisleder & Fernald 2013", 1084, 563, 598, 367, 29,
+  "Johnson et al. 2014", 1725, NA, NA, NA, 33, # also a group of 7 with AWC ~1000/hr
+  "Tamis-LeMonda et al. 2017", 2197, NA, NA, NA, 40,
+  "Pancsofar & Vernon-Feagans (2006)", 2239, NA, NA, NA, 92, # averaged together mom (2559) and dad (1919)
+  "Gilkerson et al. 2017", 1059, 356, NA, NA, 329,
+  "Hart & Risley", NA, NA, NA, NA, NA
 )
@@
   ~study, ~AWC_mean, ~AWC_sd, ~CVC_mean, ~CVC_sd, ~N,
-  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, , # these are full-day recordings (12hrs)
-  "Greenwood et al. 2011", 13142, 5562, 1714, 850, , # AWC range: 631-36563, CVC range: 11-5611
-  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, ,
-  "Zimmerman et al. 2009", 12800, 4400, NA, NA, ,
+  "Gilkerson et al. 2017", 12709, 4274, 1817, 787, 72, # these are full-day recordings (12hrs)
+  "Greenwood et al. 2011", 13142, 5562, 1714, 850, 30, # mid- to upper-SES; AWC range: 631-36563, CVC range: 11-5611
+  "Soderstrom & Wittebolle", 10125, 4890, 1744, 1058, 11, # mostly shorter recordings (6-8hrs)
+  "Zimmerman et al. 2009", 12800, 4400, NA, NA, 275, # cross-sectional; only 71 longitudinal
   "Bergelson et al. 2018", 16510, 8718, 1432, 764, 59,
 )
+# from Melanie and Elika and Alex Cristia:
+# N children = 264; N recordings = 851 AWC/hr across the recordings/participants:
+#    0%    25%    50%      75%     100%
+#  33.63  756    1087     1458    3440
 
@@
 # "Finally, children in the poor Black Belt community heard an astounding 3,203 words per hour in their ambient environment, a 102% increase over the number of words addressed to them by their primary caregivers."
+
+print(Bergelson_s2)
 ```
 
-```{r}
+Let's use the hourly rates from individual children from Sperry et al.
+(2018), Hart & Risley (1995), Soderstrom & Wittebolle (2013), and Weisleder & Fernald (2013) to estimate the parameters of hourly input rate distributions for both child-directed speech (CDS) and all adult speech.
 
+```{r, cds, echo=F, warning=F, caption="Distribution of hourly child-directed speech rates heard by individual children."}
+df <- read.csv("data/hourly_tokens_Sperry_HartRisley.csv")
+cds_agg <- df %>% summarise(mean = mean(mother_child_tokens_hr, na.rm=T),
+                           sd = sd(df$mother_child_tokens_hr, na.rm=T))
+# CDS mean: 1197.5, sd: 840.2
+breaks = seq(0, 3650, by = 120) 
+ggplot(df, aes(x = mother_child_tokens_hr)) + 
+    geom_histogram(aes(y =..density..), breaks = breaks, 
+                   color = "black", fill = "gray") +
+    stat_function(fun = dnorm, args = list(mean = cds_agg$mean, sd = cds_agg$sd)) +
+  geom_vline(xintercept=cds_agg$mean, lty="dashed") + xlab("Child-directed Speech (Tokens/Hour)") + theme_bw() +
+  annotate("text", x=cds_agg$mean +700, y=.0008, label=paste("Mean =",round(cds_agg$mean,0)))
+ggsave("figures/CDS_hourly_tokens_Sperry_HartRisley_WeislederFernald.pdf", width=4.5,height=4)
+
+# should this be a Poisson distribution? not normal
+#ggplot(transform(data.frame(x=0:3650), y=dpois(x, cds_agg$mean)), aes(x, y)) + 
+#    geom_bar(stat="identity")
 ```
+
+```{r, awc, echo=F, warning=F, caption="Distribution of hourly adult speech rates heard by individual children."}
+# all (adult) tokens Sperry + Soderstrom Wittebolle + Weisleder Fernald
+awc_agg <- df %>% summarise(mean = mean(all_tokens_hr, na.rm=T),
+                           sd = sd(df$all_tokens_hr, na.rm=T))
+# AWC: mean: 1797.8 sd=1130.2
+breaks = seq(0, 5450, by = 150) 
+ggplot(df, aes(x = all_tokens_hr)) + 
+    geom_histogram(aes(y =..density..), breaks = breaks, 
+                   color = "black", fill = "gray") + theme_bw() +
+    stat_function(fun = dnorm, args = list(mean = awc_agg$mean, sd = awc_agg$sd)) +
+  geom_vline(xintercept=awc_agg$mean, lty="dashed") + xlab("Adult Word Count (Tokens/Hour)") +
+  annotate("text", x=awc_agg$mean +700, y=.0008, label=paste("Mean =",round(awc_agg$mean,0)))
+ggsave("figures/all_hourly_tokens_Sperry_SoderstromWittebolle_WeislederFernald.pdf", width=5,height=4)
+
+nsubj_cds = sum(!is.na(df$mother_child_tokens_hr)) # 113 children
+nsubj_awc = sum(!is.na(df$all_tokens_hr)) # 83 children
+```
+
+Based on data from `r nsubj_cds` children, the mean number of child-directed tokens heard per hour is `r round(cds_agg$mean,0)` (sd=`r round(cds_agg$sd,0)`).
+Using data from `r nsubj_awc` children, the mean number of all adult tokens heard per hour is `r round(awc_agg$mean,0)` (sd=`r round(awc_agg$sd,0)`).
+
+## References
+
+Gilkerson, J., Richards, J. A., & Topping, K. (2017). Evaluation of a lena-based online intervention for parents of young children. Journal of Early Intervention, 1053815117718490.
+
+Greenwood, C.
+R., Thiemann-Bourque, K., Walker, D., Buzhardt, J., & Gilkerson, J.
+(2011).
+Assessing children’s home language environments using automatic speech recognition technology.
+Communication Disorders Quarterly, 32(2), 83-92.
+
+Soderstrom, M., & Wittebolle, K.
+(2013).
+When do caregivers talk?
+the influences of activity and time of day on caregiver speech and child vocalizations in two childcare environments.
+PloS one, 8(11), e80646.
+
+Weisleder, A., & Fernald, A. (2013). Talking to Children Matters: Early Language Experience Strengthens Processing and Builds Vocabulary. Psychological Science, 24(11), 2143–2152.
+
+Zimmerman, F.
+J., Gilkerson, J., Richards, J.
+A., Christakis, D.
+A., Xu, D., Gray, S., & Yapanel, U.
+(2009).
+Teaching by listening: The importance of adult-child conversations to language development.
+Pediatrics, 124(1), 342–349.

### 2020-06-06 app.R

@@ -80,7 +80,14 @@ ui <- fluidPage(
                          br(),
                          DT::dataTableOutput("mytable"),
                          br()
-                    )
+                    ),
+                tabPanel("Growth per Word",
+                         plotOutput("ageWord"),
+                         br()), # add selector(s) to show particular words
+                tabPanel("CDI vs. Full Vocab",
+                         plotOutput("cdi_vs_full"),
+                         br() # also show correlation of CDI and full, and mean words outside of CDI known (per age?)
+                         )
             )
         )
     )
@@ -115,14 +122,50 @@ server <- function(input, output) {
     output$ageVocab <- renderPlot({
         qs <- c(0.10,0.25,0.50,0.75,0.90)
         ggplot(sim_data()$known_words, aes(x=month, y=words)) + 
-            geom_line(aes(group = id), alpha = .1) + geom_smooth() + 
+            geom_line(aes(group = id), alpha = .1) + 
+            geom_smooth(aes(x=month, y=words), color="black") + 
             # geom_quantile(quantiles=qs, formula=y ~ poly(x, 2), aes(colour = as.factor(..quantile..))) + 
             # labs(colour="Quantile") + 
             xlab("Age (months)") + 
-            ylab("Vocabulary Size") + 
-            ylim(0, vocab_size) + xlim(1, max_age) 
+            ylab("Vocabulary Size") + #ylim(0, vocab_size) +
+            xlim(1, max_age) +
+            geom_line(aes(x = month, y = cdi_words, group = id), alpha=.1, color = "red") +
+            geom_smooth(aes(x = month, y = cdi_words), color="red")
             # geom_point(alpha=.1) +
             # geom_abline(intercept=0, slope=input$vocab_size/input$max_age, linetype="dashed", color="grey", size=1) 
+    })
+    
+    # show proportion of learners knowing each word over time
+    # (could also create a table of mean AoA per word)
+    output$ageWord <- renderPlot({
+        # maybe make this dataframe in model code..
+        dw <- sim_data()$prop_knowing_word # need to make this long
+        dl = data.frame(dw)
+        names(dl) = 1:max_age
+        dl$word = rownames(dw)
+        dl = gather(dl, "month", "prop_know", 1:max_age) 
+        dl$month = as.numeric(as.character(dl$month))
+        dl$on_cdi = wf$on_cdi
+        
[…]

### 2020-06-06 model-childesfreqs.R

@@ -4,6 +4,9 @@ max_age = 48 # months
 
 wf = read.csv("data/childes_english_word_freq_cleaned.csv")
 
+# indices of CDI items
+cdi_idx = which(wf$on_cdi==1)
+
 vocab_size = nrow(wf) # 10190
 waking_hours_per_day = 12
 
@@ -68,18 +71,16 @@ get_proc_speed <- function(print_plot=F) {
 
 # need to add proc_speed term and proc_facilitates interaction term
 simulate <- function(parms) {
-  #probs = wf_distros[[distro]] # "zipf" or "uniform"
   if(parms$distro=="uniform") {
     probs = rep(1/vocab_size, vocab_size)
-  } else {
+  } else { # "zipf"
     #probs = 1 / (1:vocab_size + 2.7) # f(r) = 1 / (r+beta)^alpha, alpha=1, beta=2.7 (Mandelbrot, 1953, 1962)
     #probs = probs/sum(probs)
     #probs = sample(probs, length(probs)) 
     probs = wf$word_count / sum(wf$word_count) # based on CHILDES WF distro
   }
   if(parms$distro=="logzipf") {
-    lp = -log(probs)
-    probs = lp / sum(lp) # max=.001, min=.003 (2x min(zipf))
+    probs = log(wf$word_count) / sum(log(wf$word_count))
   }
   start_age = parms$start_age # age (months) at which words start accumulating 
   input_rate = rnorm(n_learners, mean=parms$input_rate, sd=parms$input_rate_sd) # per-child variability in input rate
@@ -102,7 +103,12 @@ simulate <- function(parms) {
   colnames(cumulative_word_occs) = wf$word
   # child x age
   known_words = matrix(0, nrow=n_learners, ncol=max_age) # known words per individual (row) per month (col)
+  known_cdi_words = matrix(0, nrow=n_learners, ncol=max_age) # just the CDI words
   proc_speed = matrix(0, nrow=n_learners, ncol=max_age) # do we want per word instead of per learner?
+  
+  # word x age: proportion of children knowing each word per age
+  prop_knowing_word = matrix(0, nrow=vocab_size, ncol=max_age)
+  rownames(prop_knowing_word) = wf$word
   
   # processing speed - each of these parameters could have individual variability
   # Kail (1991): RT slopes follow an exponential, such that Y(i) = a + b*exp(−c*i), where Y=predicted var, i=age (mos? year?). 
@@ -140,14 +146,25 @@ simulate <- function(parms) {
[…]

### 2020-06-26 compare_fits.R
added refs, cleaned up model code a bit

@@ -1,13 +1,14 @@
 require(tidyverse)
 source("fit_quantiles.R") # for fitSSE function, which also runs the model and generates plot
 
-fits = list.files("fits")
+fits = list.files("fits/whole vocab fits/")
+#fits = list.files("fits/just CDI fits/")
 fl = list()
 df = data.frame(matrix(NA, nrow=1, ncol=10))
 names(df) = c("file", "SSE", paste0("par", 1:8))
 
 for(f in fits) {
-  load(paste0("fits/",f))
+  load(paste0("fits/whole vocab fits/",f))
   f_ = str_replace(f, ".RData", "")
   fl[[ f_ ]] = fit$optim
   df = rbind(df, c(f_, fit$optim$bestval, fit$optim$bestmem))

### 2020-06-26 model-childesfreqs.R

@@ -8,12 +8,9 @@ wf = read.csv("data/childes_english_word_freq_cleaned.csv")
 cdi_idx = which(wf$on_cdi==1)
 
 vocab_size = nrow(wf) # 10190
-waking_hours_per_day = 12
+waking_hours_per_day = 12 # eventually make normally-distributed (if we can find literature on sleeping time)
 
 
-sigmoid <- function(x) {
-  return( 1 / (1 + exp(-x)) )
-}
 
 acceleration_test <- function(dat) {
   d = dat$known_words %>% filter(month>11 & month<25) %>% group_by(month) %>% summarize(mean=mean(words)) 
@@ -44,6 +41,7 @@ make_long_df <- function(df, n_learners, max_age) {
 #  max_age=48, mean_learning_rate, learning_rate_sd, threshold_sd, 
 #  proc_facilitates, proc_speed_dev, proc_speed_dev_sd) {
 
+# call this once whenever proc_speed parms change?
 get_proc_speed <- function(print_plot=F) {
   a = rnorm(n_learners, mean=.56, sd=.1) # individual adult asymptotes for proc speed
   c = rnorm(n_learners, mean=.72, sd=.1)
@@ -73,14 +71,12 @@ get_proc_speed <- function(print_plot=F) {
 simulate <- function(parms) {
   if(parms$distro=="uniform") {
     probs = rep(1/vocab_size, vocab_size)
-  } else { # "zipf"
-    #probs = 1 / (1:vocab_size + 2.7) # f(r) = 1 / (r+beta)^alpha, alpha=1, beta=2.7 (Mandelbrot, 1953, 1962)
-    #probs = probs/sum(probs)
-    #probs = sample(probs, length(probs)) 
+  } else if(parms$distro=="zipf") {
     probs = wf$word_count / sum(wf$word_count) # based on CHILDES WF distro
-  }
-  if(parms$distro=="logzipf") {
+  } else if(parms$distro=="logzipf") {
     probs = log(wf$word_count) / sum(log(wf$word_count))
+  } else {
+    print("error")
   }
   start_age = parms$start_age # age (months) at which words start accumulating 
   input_rate = rnorm(n_learners, mean=parms$input_rate, sd=parms$input_rate_sd) # per-child variability in input rate

### 2020-10-22 paper/paper.Rmd
added refs, cleaned up model code a bit
wrote an abstract, added refs

@@
     corresponding : no 
-    address       : "420 Serra Mall"
+    address       : "420 Jane Stanford Way"
     email         : "kachergis@stanford.edu"
@@
     corresponding : yes 
-    address       : "420 Serra Mall"
+    address       : "420 Jane Stanford Way"
     email         : "mcfrank@stanford.edu"
@@
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-  One or two sentences to put the results into a more **general context**.
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
+  While there remains vigorous debate about the mechanisms by which children learn their first words, there is an emerging consensus about some criticcal factors that predict acquisition.
+  For example, although some words are more difficult to learn than others, words that are heard more often are generally learned earlier: word frequency matters.
+  Moreover, the overall amount of input heard by a child--whether child-directed or overheard speech--should influence uptake: a child hearing more hourly tokens should learn words earlier.
+  Building on previous efforts to model children's early word learning, we present a "standard model" that makes minimal, universally-agreed-upon assumptions about the learning process.
+  Using empirically-estimated input rate distributions and per-word frequencies from the CHILDES corpus, we produce simulated trajectories of vocabulary development.
+  Finally, we discuss implications of the standard model for designing and simulating interventions, as well as outstanding parameter identifiability issues and directions for further model develpment and testing.
 
@@
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["references.bib"]
 
@@
 
+How is children's linguistic input related to their uptake? 
+A wide range of computational models share the same basic accumulator architecture for word learning, in which hearing a word in the input leads to some increase in the probability of learning that word [@mcmurray2007; @hidaka2013; @mollica2017; @mayor2010].
+The goal of the current research is to unify these into a "standard" model -- analogous to the "Standard Model" of physics. 
+Our aim is to recapitulate the dynamics of how vocabulary grows, using words as the fundamental building blocks to understand the relationship between quantitative measures of language exposure and vocabulary size.
+
+We take a step towards this goal by considering Bayesian models that aim to predict children's vocabulary size using inferred parameter values. 
+We infer 1) the mean rate of input (tokens/hour) per child, and 2) a difficulty per word (the number of times a word must be heard to be learned). 
+In our baseline model, we assume that some children hear more hourly tokens than others, that these rates are normally-distributed across children, and that the difficulty of each word is also normally-distributed.
+A second model extends the baseline model to take into account varying frequency across words (estimated from the CHILDES corpus). 
+Both models assume that all children begin accumulating words at a particular *starting age*, which is to be inferred by the model.
+
 # The Standard Model
@@
 Identify critical uptake patterns to aim for (CDI quantiles?)
+
+@mcmurray2007 introduced a simple accumulator model of word learning to explain the vocabulary splurt...
 
@@
 ## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
 
@@
 ```{r create_r-references}
-r_refs(file = "r-references.bib")
+#r_refs(file = "references.bib")
 ```

### 2021-02-26 paper/paper.Rmd
moving CurrDirs ms from google doc to Rmd

--- 2020-02-19 paper/paper.Rmd
@@
 ---
-title             : "The ``Standard Model'' of Early Language Learning"
+title             : "Towards a ``Standard Model'' of Early Language Learning"
 shorttitle        : "Standard Model of Language Learning"
@@
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-  One or two sentences to put the results into a more **general context**.
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
+  A "standard model" is a theoretical framework that synthesizes observables into a quantitative consensus.
+We assess our progress towards such a synthesis for children’s early language learning.
+Several recent computational models of early vocabulary learning all share the intuition that individual words are learned through an accumulation of environmental input, a potential match with empirical work that emphasizes links between input and learning.
+However, models need to simultaneously account for 1) the fact that words vary in difficulty, as well as 2) the variability in children’s rate of exposure to and 3) ability to integrate word occurrences.
+We show how these models relate to Item-Response Theory models, which are widely-used in psychometrics.
+Unfortunately, this formal connection also makes it clear that the currently available datasets cannot allow us to fully test the model, illustrating a critical need for theory in shaping new data collection.
+In particular, to compare across populations we need to collect data that use absolute measurement units, like words heard per hour and total vocabulary size, rather than unitless measures like effect size.
+The assumptions of this nascent “standard model” already implicitly guide much work in the field, but making these assumptions explicit brings into sharp relief questions about the applicability of this model to diverse populations.
 
-keywords          : "keywords"
+keywords          : "early language learning; language acquisition; vocabulary development"
 wordcount         : "X"
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["references.bib"]
 
@@
 
-In physics, the "Standard Model" is a widely-accepted theory completed in the 1970s that describes all known elementary particles, along with three of the four known fundamental forces in the universe, explaining a wide variety of empirical phenomena.
-What fundamental building blocks and mechanisms are needed to construct an analogous standard model of early language learning?
+# Introduction
 
-Understanding children's language environment has been of central interest to both theoretical and applied and research on early language learning.
-In particular, word learning has come to be viewed as central to measuring and influencing language learning.
-Important applied area, interest in sociodemographic variation
+Early language learning is a key challenge for cognitive science research: how do we go from speechless, wordless infants to children who can use language expressively and creatively? 
+A better understanding of this process has implications for both policy and technology. 
+Yet despite rapid progress in this area, early language is still often portrayed as an area mired in controversies, for example around issues of innateness. 
+On the contrary, based on new theoretical and empirical work on the growth of vocabulary – and the deep connections between vocabulary and language more broadly – we see an emerging synthesis. 
+This synthesis provides a testable, quantitative model of the way that language is learned, connecting individual experiences to variability and growth over developmental time. 
+The goal of the current paper is to present this synthesis as a "standard model" of early language: a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work even as its assumptions are still the subject of rigorous evaluation.
 
-Word learning is one of the rare areas in developmental work in which the construct of individual differences is also seen as an epistemic construct:
+[^1]: Note that "standard" here is not meant to imply that it is normative, but rather that the model is an accepted starting point for a description of early language learning.
 
-# The Standard Model
+In physics, the "standard model" is a widely-accepted theory completed in the 1970s that describes all known elementary particles along with three of the four known fundamental forces in the universe.
+Although it is known to be incomplete and even in places incorrect, physics’ Standard Model nonetheless explains a wide variety of empirical phenomena, allowing scientists and engineers to model physical interactions with great precision and accuracy – and to design tests of the theory that inform its revision.
+Psychology in general has been criticized for lacking such formal theories to inform and drive empirical research [@muthukrishna2019problem], and we feel this critique can be applied fairly to the field of language acquisition.
+Although a substantial literature (that the current authors have contributed to) concerns the use of computational models to simulate acquisition, these models are often strikingly disconnected from the quantitative – not to mention qualitative – data gathered by most working acquisition researchers.
 
-To build a "standard model", we begin by specifying the minimal assumptions made by any model of word learning.
-First, assume that words accumulate through repeated exposure at a rate controlled by a variety of factors, including 1) learning rate, 2) frequency of exposure, 3) diversity of exposure, and 4) leveraged learning.
+But in fact, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
+Language input in the environment is processed in the moment and then accumulates via repeated exposure into learning. 
+This framework underpins much of the broader policy discussion of individual variability in language learning and its links to environmental stimulation. 
+And this basic model does in fact correspond quite nicely to a class of computational models that has been explored in the word learning literature: accumulator models.
 
-Combine to create a "macro-economics of early language" 
-Input distributional patterns that match corpus dynamics (Montag etc)
-Identify critical uptake patterns to aim for (CDI quantiles?)
+Our aim here is to make this connection explicit, using item-response theory from psychometric testing as a framework for connecting between empirical data and previous accumulator models. 
+Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal verbal theories and evaluate them quantitatively [@vanrooij2020theory].
+Importantly, the goal is not to produce “the correct model” but rather to explore the ways that model assumptions lead to predictions. 
+In fact, the greatest value of such models is likely the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation (Tauber et al., 2016).
 
-## Extensions
-ambiguity and disambiguation
-Cross-situational extensions (Smith, Blythe) - most word learning situations are ambiguous and require disambiguation across time (using mechanisms of learning and memory)
-Socio-pragmatic extensions - most word learning instances are disambiguated by social and pragmatic cues (and are unambiguous)
-Experimental evidence from the "associations vs. hypotheses" debate
+In the remainder of this paper, we introduce accumulator models in more detail and then briefly present how they connect with IRT models. 
+We then review evidence consistent with this sketch of a “standard model”, as well as some empirical challenges for the future development.
 
-# Procedure
+```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", fig.cap=c("A schematic of the “standard” relationships between variables assumed in the literature on early language learning.
+It is generally assumed that child-directed speech (red) is more valuable than overheard speech (blue)."), echo=FALSE}
+knitr::include_graphics("./figs/fig1_standard_model.pdf")
+```
 
-## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
+# Theorizing About the Accumulation of Language
 
-# Results
+We are interested in children’s vocabulary, which we define as the set of words[^2] they know. 
+Although vocabulary has sometimes been thought of as disconnected from other – often more “interesting” – parts of language learning, in fact the language system is tightly woven [@wordbankbook]. 
+Children’s vocabulary size predicts a huge proportion of the concurrent variance in their production of complex grammatical utterances and gestures and is predictively related to early speech perception [@bates1994developmental; @wordbankbook; @cristia2014predicting].
+Further, measures of children’s early language are highly correlated with one another both concurrently and predictively, suggesting a single latent variable – language ability – that is quite stable across early childhood [@bornstein2013language].
+
+[^2]: What makes a word is its own separate and difficult question, especially as we look across languages.
+Are “dog” and “dogs” two separate words?
+If not, what about “mouse” and “mice”?
+These questions are relatively inconsequential in morphologically simple languages like English and Mandarin; yet they are quite a bit more difficult in morphologically more complex languages like Turkish or even agglutinating languages like Inuktitut.
+Here we sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating something – whether or not we call it a word per se, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words are.
+For some initial data bearing out this supposition by comparing Turkish with other languages, see @braginsky2019.
+
+In early theories of language acquisition, children’s environment was assumed to “trigger” specific innate developmental schemas (e.g., Gibson & Wexler, 1994). 
+Driven by influential observations of the tight relationship between environmental input and children’s learning (e.g., Huttenlocher et al., 1991; Hart & Risley, 1995), more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
+One such class of models is accumulator models, which assume that linguistic experiences with words accumulate in separate registers (pictured as “buckets” in Figure 2), with those registers exceeding a particular threshold being learned.
+
+Formalized theories of accumulative vocabulary growth have already contributed significantly to theoretical debate. 
+For example, @mcmurray2007 elegantly demonstrated that children’s “vocabulary explosion”  – an acceleration in the growth of children’s vocabulary in their second year of life – can result from the steady accumulation of word tokens without any change in the environment or learning mechanism.
+Other work has examined similar models using more realistic distributions of words (i.e., Zipf-distributed), developmental change in learning mechanisms, and quantitative comparison to children’s aggregate vocabulary growth [@mayor2010; @hidaka2013; @mollica2017].
+Although this work is exciting, these models have not yet been used to make direct predictions about learning in individual children (Bergelson, 2020).
+
+One difficulty in connecting such models to data is that the relevant variables are often measured in relative, rather than absolute, units.
+This practice is common – for example, measures of intelligence are given on a standardized scale defined by population variability.
+Yet in early language, we have – relatively unusually in a psychological domain – access to absolute units of description. 
+We can count how many words a child hears and express this as a rate (e.g., words per hour; Cristia et al., 2019), and we can similarly estimate how many words they know [@wordbankbook]. 
+These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations (Dailey & Bergelson, in press).
+As we describe below, however, few datasets have measurements of the relevant variables. 
+This may be a case where theory outstrips data and further data collection efforts are necessary.
+
+```{r, out.width="0.7\\linewidth", include=TRUE, fig.align="center", fig.cap=c("An illustration of an accumulator model: each bucket represents the child’s knowledge about a particular word, and each token is a drop in the corresponding bucket.
+Some words are more difficult than others (i.e., have larger buckets).
+When a bucket is full, the corresponding word is considered to be learned.
+In theory, language input coming from child-directed speech (red) may count more than tokens in overheard speech (blue)."), echo=FALSE}
+knitr::include_graphics("./figs/fig2_standard_model.pdf")
+```
+
+# Parsing the Literature Using the Standard Model
+
+Since seminal work by Hart & Risley (1996), the connection between children’s language input and the growth of vocabulary has been a topic of intense interest, and numerous studies have reported associations between these variables [e.g., @hoff2003specificity; @weisleder2013].
+The direction of these associations is of course congruent with the standard model (and indeed is generally an inspiration). 
+But such correlations are also partially explained by a number of other factors, including socioeconomic [@hoff2003specificity; dailey2021language] and genetic [@hayiou2014language] variables, and their magnitude varies across studies [@wang2020meta].
+Though such studies are costly and difficult to conduct, randomized interventions are the gold standard for measuring causal effects of parental language input on language learning. 
+When such studies are conducted, they tend to show modest but reliable effects [e.g., suskind2016project]. 
+These positive results provide support for a causal connection between language input and outcomes, but – because they only estimate the effect of the change in input on variation in outcome – they provide only a relative, rather than an absolute, estimate of how important input is to vocabulary learning.
+
+Indeed, reasoning from first principles, you cannot learn the word “table” if you do not hear it. Input must be a predictor of learning. 
+Correlational studies do not assess this relationship fully – rather, they assess whether variation in input relates to variation in learning. 
+A second line of work that comes closer to assessing the importance of language input focuses not on differences between children but instead on differences between words. 
+These studies use regression models to predict which words are easier or more difficult, typically averaging across children (e.g., Goodman, Dale, & Li, 2008; Braginsky et al., 2018). 
+Such studies show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
+Such models can also be used to assess the value of other proposals for learning mechanisms in the literature (e.g., Fourtassi et al., 2020).
+
+More broadly, this literature illustrates the value of a quantitative process model: under the standard model and its variants, we should be able to compute the expected relation between measures of input and vocabulary size.
+To the extent this model accounts for the data, it will support researchers’ current theory, but likely more informative will be the model’s pattern of misfit to the data, which should provide grist for the twin mills of theory development and empirical study.
+
+# Toward a “Standard” Accumulator Model
+
+As we have argued above, the core of the standard view of language learning is that it is a process of accumulation – individual experiences with words lead to their eventual acquisition. 
+The more of these experiences a child receives, the faster their vocabulary grows. 
+However, both children and words vary: children may learn faster or slower, and words can be more or less difficult to learn. 
+Combining these ideas, the basic hypothesis is that a child’s vocabulary at a particular time should be predicted by their cumulative language exposure and learning rate, combined with the breadth and individual difficulties of the words that they have been exposed to.
+
+The key contribution of such a baseline theory is that it establishes the link between the observable rates of words that children hear and their eventual knowledge of those words. 
+For example, in principle the model could establish how many occurrences of the word “dog” are needed by the average child in order to start understanding (or producing) “dog.” 
+Moreover, given a child’s vocabulary at a certain age, the model would provide an estimate of the child’s language ability.
+
+This model outlines an approach that is very similar to Item-Response Theory (IRT; Embretson & Reise, 2000), an influential and popular psychometric modeling framework for understanding the performance of test-takers (in our case language learners whose vocabulary is being assessed) as they are assessed with a particular set of test items (the particular words we ask about).
+In particular, IRT models jointly infer both how hard specific test items are and how accurate test takers are, recovering these latent parameters from the observed data. 
+In the basic Rasch (or 1-parameter logistic) IRT model, a person i responds correctly to item j with probability determined by their ability ($\theta_i$) and the difficulty of item j ($d_j$).
+
+1PL eqn
+
+These parameters can easily be mapped onto the kind of accumulator model we have been describing: items are words (e.g.
+$d_j$ is the difficulty of word j), and $\theta_i$ is child $i$’s estimated latent language ability.
+Although in a typical IRT model both item difficulties and person abilities are standardized (assumed to be normally-distributed and centered on 0) and unit-free, in principle it is possible to map these scores to real-world distributions – as we would like to do – incorporating measured word frequencies and rates of children’s experienced input.
+With this mapping into absolute units, this kind of model then can provide a quantitative linking hypothesis between measurements of input and learning. 
+Recognizing the parallels between the standard model desiderata and IRT theory, we can use standard extensions of the 1PL (Rasch) IRT model to include both person-level and item-level covariates.
+Item-level (word-level) covariates of central interest in the standard model are estimates of word frequency in child-directed speech as well as lexical class [whether they are a noun, a predicate, or function word; @goodman2008does; @braginsky2019].
+Person-level covariates of interest in a standard model might include sex, socioeconomic status variables, or other demographic information.
+
+This model as described above formalizes an important and testable claim: is the amount of language experience the important predictor of learning for a child, or are there other age-related factors beyond experience?
+Put simply, is a 2-year-old different than a 1-year-old, aside from having twice as much experience with each word? 
+This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is included as a person-level covariate, and one in which age is instead included as a scalar on the prevalence of each item (i.e., in which the frequency of “dog” for a 2-year-old is twice that of “dog” for a 1-year-old).
+Using the IRT framework to formalize theoretical questions and assess them via model comparison is a strength of this approach. 
+Below we report the results of this model comparison, fit to parent-reported vocabulary data from 5,492 English-speaking children in Wordbank. 
+Each model estimated parameters for each child’s ability, and for the 680 words on the CDI Words and Sentences form. 
+As covariates we included the lexical class of each word (function words, adjective, noun, verb, or other), and the estimated numbers of tokens per word a child is likely to hear each month (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech (MacWhinney, 2000).
+An average child in this model (with input rate based on reported empirical means) is expected to receive 1,200 words per hour, 12 hours each day, for a total of 438,300 tokens per month, of which 285,200 tokens (65.1%) are accounted for by tokens of CDI words.
+
+In Model 1, we included age as a person-level covariate, allowing for interactions of age (in months), lexical class, and monthly tokens per word. 
+In Model 2, instead of including the child’s age as a person-level covariate, we scaled the expected tokens per month for each word by the child’s age, yielding an estimate of their lifetime tokens per word, and also allowed for interactions with lexical class.
+Model selection metrics (AIC and BIC) preferred the more complex Model 1, with a person-level age covariate separate from the item prevalence covariate. 
+Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, prevalence (left), as well as for specific
+
+then outline a series of candidate “standard” models to test down the road--some of which will require new data sources in order to differentiate.
+
+```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from the fitted English “standard model” as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 CDI words (right)."), echo=FALSE}
+knitr::include_graphics("./figs/fig3_predicted_age_LC_and_items.pdf")
+```
 

### 2021-02-27 paper/paper.Rmd
adding refs

@@
 
-In early theories of language acquisition, children’s environment was assumed to “trigger” specific innate developmental schemas (e.g., Gibson & Wexler, 1994). 
-Driven by influential observations of the tight relationship between environmental input and children’s learning (e.g., Huttenlocher et al., 1991; Hart & Risley, 1995), more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
+In early theories of language acquisition, children’s environment was assumed to “trigger” specific innate developmental schemas [e.g., @gibson1994triggers]. 
+Driven by influential observations of the tight relationship between environmental input and children’s learning [e.g., huttenlocher1991early; @hart1995), more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
 One such class of models is accumulator models, which assume that linguistic experiences with words accumulate in separate registers (pictured as “buckets” in Figure 2), with those registers exceeding a particular threshold being learned.
@@
 Other work has examined similar models using more realistic distributions of words (i.e., Zipf-distributed), developmental change in learning mechanisms, and quantitative comparison to children’s aggregate vocabulary growth [@mayor2010; @hidaka2013; @mollica2017].
-Although this work is exciting, these models have not yet been used to make direct predictions about learning in individual children (Bergelson, 2020).
+Although this work is exciting, these models have not yet been used to make direct predictions about learning in individual children [@bergelson2020comprehension].
 
@@
 We can count how many words a child hears and express this as a rate (e.g., words per hour; Cristia et al., 2019), and we can similarly estimate how many words they know [@wordbankbook]. 
-These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations (Dailey & Bergelson, in press).
+These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
 As we describe below, however, few datasets have measurements of the relevant variables. 
@@
 A second line of work that comes closer to assessing the importance of language input focuses not on differences between children but instead on differences between words. 
-These studies use regression models to predict which words are easier or more difficult, typically averaging across children (e.g., Goodman, Dale, & Li, 2008; Braginsky et al., 2018). 
+These studies use regression models to predict which words are easier or more difficult, typically averaging across children [e.g., @goodman2008does; @braginsky2019). 
 Such studies show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
@@
 
-This model outlines an approach that is very similar to Item-Response Theory (IRT; Embretson & Reise, 2000), an influential and popular psychometric modeling framework for understanding the performance of test-takers (in our case language learners whose vocabulary is being assessed) as they are assessed with a particular set of test items (the particular words we ask about).
+This model outlines an approach that is very similar to Item-Response Theory [IRT; @embretson2013item], an influential and popular psychometric modeling framework for understanding the performance of test-takers (in our case language learners whose vocabulary is being assessed) as they are assessed with a particular set of test items (the particular words we ask about).
 In particular, IRT models jointly infer both how hard specific test items are and how accurate test takers are, recovering these latent parameters from the observed data. 
@@
 Each model estimated parameters for each child’s ability, and for the 680 words on the CDI Words and Sentences form. 
-As covariates we included the lexical class of each word (function words, adjective, noun, verb, or other), and the estimated numbers of tokens per word a child is likely to hear each month (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech (MacWhinney, 2000).
+As covariates we included the lexical class of each word (function words, adjective, noun, verb, or other), and the estimated numbers of tokens per word a child is likely to hear each month (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech [@macwhinney2000childes].
 An average child in this model (with input rate based on reported empirical means) is expected to receive 1,200 words per hour, 12 hours each day, for a total of 438,300 tokens per month, of which 285,200 tokens (65.1%) are accounted for by tokens of CDI words.

### 2021-03-04 paper/paper.Rmd
some writing

@@
 keywords          : "early language learning; language acquisition; vocabulary development"
-wordcount         : "X"
+wordcount         : "3334"
 
@@
 In early theories of language acquisition, children’s environment was assumed to “trigger” specific innate developmental schemas [e.g., @gibson1994triggers]. 
-Driven by influential observations of the tight relationship between environmental input and children’s learning [e.g., huttenlocher1991early; @hart1995), more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
+Driven by influential observations of the tight relationship between environmental input and children’s learning [e.g., @huttenlocher1991early; @hart1995], more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
 One such class of models is accumulator models, which assume that linguistic experiences with words accumulate in separate registers (pictured as “buckets” in Figure 2), with those registers exceeding a particular threshold being learned.
@@
 In particular, IRT models jointly infer both how hard specific test items are and how accurate test takers are, recovering these latent parameters from the observed data. 
-In the basic Rasch (or 1-parameter logistic) IRT model, a person i responds correctly to item j with probability determined by their ability ($\theta_i$) and the difficulty of item j ($d_j$).
+In the basic Rasch (or 1-parameter logistic) IRT model, a person i responds correctly to item j with probability determined by their ability ($\theta_i$) and the difficulty of item j ($d_j$):
 
-1PL eqn
+$$ P(y_{i,j} = 1 | \theta_i, d_j) = \frac{1}{1+e^{\theta_i + d_j}} $$
 
@@
 Model selection metrics (AIC and BIC) preferred the more complex Model 1, with a person-level age covariate separate from the item prevalence covariate. 
-Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, prevalence (left), as well as for specific
+Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, and prevalence (left), showing for example that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
+Moreover, the model predicts item-level acquisition curves (Figure 3, right), for example predicting that “ball” is learned earlier than “dog”, and that “go” is easier than “have.” Model 1 represents a new baseline standard model, which might be extended to include person-level groups such as SES and sex, as well as item-level covariates such as word length.
 
-then outline a series of candidate “standard” models to test down the road--some of which will require new data sources in order to differentiate.
+Any model in this framework also estimates the latent language ability of each child in the dataset -- a value correlated with age, and with a child’s total vocabulary size as estimated by the CDI, but that may be better correlated with the child’s quantity/quality of input received, or with their processing speed: relations that will require new datasets with concurrent measures of these variables in order to distinguish.
+For one lesson we have learned in synthesizing the standard model framework is that based on current datasets, some of the theories are currently indistinguishable from one another, the models unidentifiable.
+Thus, we conclude by briefly discussing the next steps needed to test and build upon the standard model.
 
-```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from the fitted English “standard model” as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 CDI words (right)."), echo=FALSE}
+```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from the fitted English “standard model” as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 words on the CDI (right)."), echo=FALSE}
 knitr::include_graphics("./figs/fig3_predicted_age_LC_and_items.pdf")
@@
 
-# Discussion
+# Onward from the Standard Model
+
+## Leveraged learning and the role of processing
+
+As the model results above show, older children accumulate language from their input faster than younger children. 
+Why? One potential reason is that independent developmental or maturational factors simply make older children better at remembering their experiences. 
+Our model suggests that domain general developmental differences in learning and memory would be important to measure independently to see how the ability to learn and remember words (or word-object associations) changes with age.
+
+Another possibility – not mutually exclusive – is that children “leverage” their knowledge of language to learn faster [@mitchell2009]. 
+This leverage idea is consistent with many proposed mechanisms for language learning.
+For example, if children reason about new vocabulary items by exclusion (e.g., Markman et al., 1988), they could get better at learning new words just by having more potential items to exclude from consideration.
+Certainly children can perform this kind of reasoning robustly across a range of ages, though the magnitude of its contributions to the broader accumulation of vocabulary is still unknown (Lewis et al., 2020).
+
+Yet another possibility – again not mutually exclusive to others – is that children are getting better at using the words they know. 
+Several recent studies have found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words (e.g., Fernald, Perfors & Marchman, 2006; Marchman, Admas, Loi, Fernald, & Feldman, 2015).
+Children’s processing speed in these studies is measured with the looking-while-listening (LWL) paradigm (Fernald, Pinto, Swingley, Weinberg, McRoberts, 1998), in which children are presented with an onscreen array of familiar objects (e.g., a shoe and a ball), and their latency to look at a named target object (ball) is measured.
+Much of this work has been correlational, however. 
+Thus, it is unclear whether children with larger vocabularies are better at processing familiar words – as one might expect based on adult processing speed results, which find faster and more accurate processing of words in adults with larger vocabularies (Mainz, Shao, Brysbaert, & Meyer, 2017), or if the ability to quickly process familiar words generalizes to being faster to learn new words, as is sometimes speculated (e.g., Law & Edwards, 2015; Peter et al., 2019).
+This latter would be a form of leverage as well.
+
+In sum, given that...
+
+## Beyond Vocabulary
+
+The “Standard Model” that we have proposed is a model of the accumulation of words, but language is rich, complex..
+
+Empirically, The language system is tightly woven. 
+	Grammar Lex correlations
+	Gestures lex 
+	Phonology lex (cristia)
+Bornstein
+Theoretically, accumulator models are generic models of skill acquisition. One perspective on language is that it is in fact a skill [@chater2018language]. 
+Dovetails with construction grammar theory
+A Theory for Understanding Diverse Outcomes
+
+# Conclusions
+
+We have proposed here a way of implementing the theory that implicitly drives much of the research (and policy-making) on early language acquisition. 
+This “standard” model synthesizes measures of language input, processing, and vocabulary growth, allowing us to formalize, test, and iteratively improve our theoretical understanding of early language acquisition.
+Moreover, by situating the standard model in the classic IRT psychometric framework, we have identified specific knowledge gaps in current empirical approaches that must be closed in future work in order to inform future theory development.
 

### 2021-03-04 paper/paper.Rmd
adding refs

@@
 
-Since seminal work by Hart & Risley (1996), the connection between children’s language input and the growth of vocabulary has been a topic of intense interest, and numerous studies have reported associations between these variables [e.g., @hoff2003specificity; @weisleder2013].
+Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest, and numerous studies have reported associations between these variables [e.g., @hoff2003specificity; @weisleder2013].
 The direction of these associations is of course congruent with the standard model (and indeed is generally an inspiration). 
-But such correlations are also partially explained by a number of other factors, including socioeconomic [@hoff2003specificity; dailey2021language] and genetic [@hayiou2014language] variables, and their magnitude varies across studies [@wang2020meta].
+But such correlations are also partially explained by a number of other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables, and their magnitude varies across studies [@wang2020meta].
 Though such studies are costly and difficult to conduct, randomized interventions are the gold standard for measuring causal effects of parental language input on language learning. 
-When such studies are conducted, they tend to show modest but reliable effects [e.g., suskind2016project]. 
+When such studies are conducted, they tend to show modest but reliable effects [e.g., @suskind2016project]. 
 These positive results provide support for a causal connection between language input and outcomes, but – because they only estimate the effect of the change in input on variation in outcome – they provide only a relative, rather than an absolute, estimate of how important input is to vocabulary learning.
@@
 A second line of work that comes closer to assessing the importance of language input focuses not on differences between children but instead on differences between words. 
-These studies use regression models to predict which words are easier or more difficult, typically averaging across children [e.g., @goodman2008does; @braginsky2019). 
+These studies use regression models to predict which words are easier or more difficult, typically averaging across children [e.g., @goodman2008does; @braginsky2019]. 
 Such studies show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
-Such models can also be used to assess the value of other proposals for learning mechanisms in the literature (e.g., Fourtassi et al., 2020).
+Such models can also be used to assess the value of other proposals for learning mechanisms in the literature [e.g., @fourtassi2020growth].
 
@@
 Although in a typical IRT model both item difficulties and person abilities are standardized (assumed to be normally-distributed and centered on 0) and unit-free, in principle it is possible to map these scores to real-world distributions – as we would like to do – incorporating measured word frequencies and rates of children’s experienced input.
-With this mapping into absolute units, this kind of model then can provide a quantitative linking hypothesis between measurements of input and learning. 
+With this mapping into absolute units, this kind of model can provide a quantitative linking hypothesis between measurements of input and learning. 
 Recognizing the parallels between the standard model desiderata and IRT theory, we can use standard extensions of the 1PL (Rasch) IRT model to include both person-level and item-level covariates.
@@
 Each model estimated parameters for each child’s ability, and for the 680 words on the CDI Words and Sentences form. 
-As covariates we included the lexical class of each word (function words, adjective, noun, verb, or other), and the estimated numbers of tokens per word a child is likely to hear each month (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech [@macwhinney2000childes].
+As covariates we included the lexical class of each word (function words, adjective, noun, verb, or other), and the estimated number of monthly tokens per word a child is likely to hear (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech [@macwhinney2000childes].
 An average child in this model (with input rate based on reported empirical means) is expected to receive 1,200 words per hour, 12 hours each day, for a total of 438,300 tokens per month, of which 285,200 tokens (65.1%) are accounted for by tokens of CDI words.

### 2021-03-04 paper/paper.Rmd
adding refs

@@
 Yet another possibility – again not mutually exclusive to others – is that children are getting better at using the words they know. 
-Several recent studies have found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words (e.g., Fernald, Perfors & Marchman, 2006; Marchman, Admas, Loi, Fernald, & Feldman, 2015).
+Several recent studies have found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words (e.g., @fernald2006picking; @marchman2016early).
 Children’s processing speed in these studies is measured with the looking-while-listening (LWL) paradigm (Fernald, Pinto, Swingley, Weinberg, McRoberts, 1998), in which children are presented with an onscreen array of familiar objects (e.g., a shoe and a ball), and their latency to look at a named target object (ball) is measured.

### 2021-03-04 paper/paper.Rmd
last sections

@@
     affiliation   : "1"
+    corresponding : yes 
+    address       : "450 Jane Stanford Way, Stanford, CA 94305"
+    email         : "kachergis@stanford.edu"
+  - name          : "Virginia A. Marchman"
+    affiliation   : "1"
     corresponding : no 
-    address       : "420 Jane Stanford Way"
-    email         : "kachergis@stanford.edu"
+    address       : "450 Jane Stanford Way, Stanford, CA 94305"
+    email         : "mcfrank@stanford.edu"
   - name          : "Michael C. Frank"
     affiliation   : "1"
-    corresponding : yes 
-    address       : "420 Jane Stanford Way"
+    corresponding : no 
+    address       : "450 Jane Stanford Way, Stanford, CA 94305"
     email         : "mcfrank@stanford.edu"
@@
 
-# Toward a “Standard” Accumulator Model
+# Toward a "Standard" Accumulator Model
 
@@
 
-## Leveraged learning and the role of processing
+The goal of a computational theory like the "standard model" described here is to describe the quantitative predictions that arise from a set of assumptions.
+Often, however, it is the *failure* of such a model to predict observed patterns of data that is most useful, as these failures point the way forward towards future refinements [@vanrooij2020theory].
+We discuss three potential failures here as they point towards ways forward.
 
-As the model results above show, older children accumulate language from their input faster than younger children. 
+**Leveraged learning and the role of processing**. As the model results above show, older children accumulate language from their input faster than younger children. 
 Why? One potential reason is that independent developmental or maturational factors simply make older children better at remembering their experiences. 
-Our model suggests that domain general developmental differences in learning and memory would be important to measure independently to see how the ability to learn and remember words (or word-object associations) changes with age.
+Another possibility -- not mutually exclusive -– is that children *leverage* their knowledge of language to learn faster [@mitchell2009]. 
+This leverage idea is consistent with many proposed mechanisms for language learning.
+For example, if children reason about new vocabulary items by exclusion [@markman1988], they could get better at learning new words just by having more potential items to exclude from consideration.
+Yet a third possibility –- again not mutually exclusive with the others -– is that children are getting better at using the words they know. 
+Several recent studies have found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words [e.g., @fernald2006picking; @marchman2016early].
+Disentangling these distinct explanations will require careful measurement of both vocabulary and the potential mechanisms of leverage [cf. @lewis2020].
 
-Another possibility – not mutually exclusive – is that children “leverage” their knowledge of language to learn faster [@mitchell2009]. 
-This leverage idea is consistent with many proposed mechanisms for language learning.
-For example, if children reason about new vocabulary items by exclusion (e.g., Markman et al., 1988), they could get better at learning new words just by having more potential items to exclude from consideration.
-Certainly children can perform this kind of reasoning robustly across a range of ages, though the magnitude of its contributions to the broader accumulation of vocabulary is still unknown (Lewis et al., 2020).
+**Beyond vocabulary**.
+The "standard model" that we have proposed is a model of the accumulation of words, but language is rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings.
+In early conceptions of language learning, syntax and morphology were thought of as distinct and unconnected, but this conception has not been borne out empirically.
+Instead, evidence from large-scale data show again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@frank2021].
+Theoretically, accumulator models like the standard model are generic models of skill acquisition.
+Given the perspective that language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards exensions of the standard model to the accumulation of broader units of language like constructions.
 
-Yet another possibility – again not mutually exclusive to others – is that children are getting better at using the words they know. 
-Several recent studies have found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words (e.g., @fernald2006picking; @marchman2016early).
-Children’s processing speed in these studies is measured with the looking-while-listening (LWL) paradigm (Fernald, Pinto, Swingley, Weinberg, McRoberts, 1998), in which children are presented with an onscreen array of familiar objects (e.g., a shoe and a ball), and their latency to look at a named target object (ball) is measured.
-Much of this work has been correlational, however. 
-Thus, it is unclear whether children with larger vocabularies are better at processing familiar words – as one might expect based on adult processing speed results, which find faster and more accurate processing of words in adults with larger vocabularies (Mainz, Shao, Brysbaert, & Meyer, 2017), or if the ability to quickly process familiar words generalizes to being faster to learn new words, as is sometimes speculated (e.g., Law & Edwards, 2015; Peter et al., 2019).
-This latter would be a form of leverage as well.
-
-In sum, given that...
-
-## Beyond Vocabulary
-
-The “Standard Model” that we have proposed is a model of the accumulation of words, but language is rich, complex..
-
-Empirically, The language system is tightly woven. 
-	Grammar Lex correlations
-	Gestures lex 
-	Phonology lex (cristia)
-Bornstein
-Theoretically, accumulator models are generic models of skill acquisition. One perspective on language is that it is in fact a skill [@chater2018language]. 
-Dovetails with construction grammar theory
-A Theory for Understanding Diverse Outcomes
+**A theory for understanding acquisition in diverse contexts**.
+The "standard model" formalizes an implict assumption of the literature: namely, that early language learning follows the same general model for children in radically different circumstances.
+Yet this assumption could very well be false.
+Some children might learn more from overheard speech and others might learn more from child-directed speech [@sperry2019].
+The kind of data necessary to test this assumption directly are only now being collected, for example in studies that rigorously track learning outcomes and amount of language input in diverse populations [e.g., comparisons between children growing up in low-income, rural, indigenous communities and those in higher-income Western contexts; @casillas2020].
 
@@
 
-We have proposed here a way of implementing the theory that implicitly drives much of the research (and policy-making) on early language acquisition. 
-This “standard” model synthesizes measures of language input, processing, and vocabulary growth, allowing us to formalize, test, and iteratively improve our theoretical understanding of early language acquisition.
-Moreover, by situating the standard model in the classic IRT psychometric framework, we have identified specific knowledge gaps in current empirical approaches that must be closed in future work in order to inform future theory development.
+An implicit theory drives much research and policy-making on early language acquisition: early language accumulates through a set of discrete experiences with individual words. 
+The more experiences, the faster the words are learned. 
+This implicit theory can be expressed as a simple computational model that makes quantitative predictions. 
+By situating this model in a common psychometric framework, we show how it can be used to connect to large-scale, cross-linguistic datasets. 
+This “standard model" synthesizes measures of language input and vocabulary growth, allowing us to formalize, test, and iteratively improve our theoretical understanding of early language acquisition.
+Moreover, this formalization allows us to identify specific gaps in current empirical approaches that must be closed to inform future theory development.
 

### 2021-03-04 paper/paper.Rmd
minor edits, last refs
we're now at 3463 words and 33 refs (goal: 3000 and 30)

@@
 keywords          : "early language learning; language acquisition; vocabulary development"
-wordcount         : "3334"
+wordcount         : "3463"
 
@@
 Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, and prevalence (left), showing for example that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
-Moreover, the model predicts item-level acquisition curves (Figure 3, right), for example predicting that “ball” is learned earlier than “dog”, and that “go” is easier than “have.” Model 1 represents a new baseline standard model, which might be extended to include person-level groups such as SES and sex, as well as item-level covariates such as word length.
+Moreover, the model predicts item-level acquisition curves (Figure 3, right), for example predicting that “ball” is learned earlier than “dog”, and that “go” is easier than “have.” 
+Model 1 represents a new baseline standard model, which might be extended to include person-level groups such as SES and sex, as well as item-level covariates such as word length.
 
@@
 This leverage idea is consistent with many proposed mechanisms for language learning.
-For example, if children reason about new vocabulary items by exclusion [@markman1988], they could get better at learning new words just by having more potential items to exclude from consideration.
+For example, if children reason about new vocabulary items by exclusion [@markman1988children], they could get better at learning new words just by having more potential items to exclude from consideration.
 Yet a third possibility –- again not mutually exclusive with the others -– is that children are getting better at using the words they know. 
 Several recent studies have found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words [e.g., @fernald2006picking; @marchman2016early].
-Disentangling these distinct explanations will require careful measurement of both vocabulary and the potential mechanisms of leverage [cf. @lewis2020].
+Disentangling these distinct explanations will require careful measurement of both vocabulary and the potential mechanisms of leverage [cf. @lewis2020role].
 
@@
 In early conceptions of language learning, syntax and morphology were thought of as distinct and unconnected, but this conception has not been borne out empirically. 
-Instead, evidence from large-scale data show again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@frank2021].
+Instead, evidence from large-scale data show again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@wordbankbook].
 Theoretically, accumulator models like the standard model are generic models of skill acquisition. 
-Given the perspective that language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards exensions of the standard model to the accumulation of broader units of language like constructions.
+Given the perspective that language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards extensions of the standard model to the accumulation of broader units of language like constructions.
 
@@
 ```{r create_r-references}
-#r_refs(file = "references.bib")
+#r_refs(file = "references.bib") # 31 refs right now
 ```

### 2021-03-09 paper/paper.Rmd
tweaks

--- 2021-03-04 paper/paper.Rmd
@@
 ---
-title             : "Towards a ``Standard Model'' of Early Language Learning"
+title             : "Towards a \"Standard Model\" of Early Language Learning"
 shorttitle        : "Standard Model of Language Learning"
@@
 
-floatsintext      : no
+floatsintext      : yes
 figurelist        : no

### 2021-03-09 paper/paper.Rmd
minor edits

--- 2021-03-04 paper/paper.Rmd
@@
 Often, however, it is the *failure* of such a model to predict observed patterns of data that is most useful, as these failures point the way forward towards future refinements [@vanrooij2020theory]. 
-We discuss three potential failures here as they point towards ways forward.
+We discuss three potential failures here as signposts for ways forward.
 
@@
 **Beyond vocabulary**.
-The "standard model" that we have proposed is a model of the accumulation of words, but language is rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings.
+The "standard model" that we have proposed is a model of the accumulation of words, but language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings.
 In early conceptions of language learning, syntax and morphology were thought of as distinct and unconnected, but this conception has not been borne out empirically. 
-Instead, evidence from large-scale data show again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@wordbankbook].
+Instead, evidence from large-scale data shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@wordbankbook].
 Theoretically, accumulator models like the standard model are generic models of skill acquisition. 

### 2021-03-09 paper/paper.Rmd
incorporating Virginia's edits

@@
 Several recent computational models of early vocabulary learning all share the intuition that individual words are learned through an accumulation of environmental input, a potential match with empirical work that emphasizes links between input and learning.
-However, models need to simultaneously account for 1) the fact that words vary in difficulty, as well as 2) the variability in children’s rate of exposure to and 3) ability to integrate word occurrences.
-We show how these models relate to Item-Response Theory models, which are widely-used in psychometrics.
+However, models need to simultaneously account for variation in: 1) word difficulty, 2) children’s rate of exposure to words, and 3) children’s ability to integrate word occurrences.
+We show how these models relate to Item-Response Theory models, widely-used in psychometrics.
 Unfortunately, this formal connection also makes it clear that the currently available datasets cannot allow us to fully test the model, illustrating a critical need for theory in shaping new data collection.
-In particular, to compare across populations we need to collect data that use absolute measurement units, like words heard per hour and total vocabulary size, rather than unitless measures like effect size.
+In particular, to compare across populations, we need to collect data that use absolute measurement units, like words heard per hour and total vocabulary size, rather than unitless measures like effect size.
 The assumptions of this nascent “standard model” already implicitly guide much work in the field, but making these assumptions explicit brings into sharp relief questions about the applicability of this model to diverse populations.
@@
 
-Early language learning is a key challenge for cognitive science research: how do we go from speechless, wordless infants to children who can use language expressively and creatively? 
+Early language learning is a key challenge for cognitive science: how do we go from speechless, wordless infants to children who can use language expressively and creatively? 
 A better understanding of this process has implications for both policy and technology. 
-Yet despite rapid progress in this area, early language is still often portrayed as an area mired in controversies, for example around issues of innateness. 
-On the contrary, based on new theoretical and empirical work on the growth of vocabulary – and the deep connections between vocabulary and language more broadly – we see an emerging synthesis. 
-This synthesis provides a testable, quantitative model of the way that language is learned, connecting individual experiences to variability and growth over developmental time. 
-The goal of the current paper is to present this synthesis as a "standard model" of early language: a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work even as its assumptions are still the subject of rigorous evaluation.
+Despite rapid progress in this area, early language is still often portrayed as mired in controversies, for example, around issues of innateness. 
+However, based on new theoretical and empirical work on the growth of vocabulary, and the deep connections between vocabulary and language more broadly, we see an emerging synthesis. 
+This synthesis provides a testable, quantitative model language learning, connecting individual experiences to variability and growth over developmental time. 
+Our goal is to present this synthesis as a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work even as its assumptions are still the subject of rigorous evaluation.
 
@@
 **Beyond vocabulary**.
-The "standard model" that we have proposed is a model of the accumulation of words, but language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings.
+The "standard model" proposed here is a model of the accumulation of words, but language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings.
 In early conceptions of language learning, syntax and morphology were thought of as distinct and unconnected, but this conception has not been borne out empirically. 

### 2021-03-09 paper/paper.Rmd
incorporating Virginia's edits

@@
 This synthesis provides a testable, quantitative model language learning, connecting individual experiences to variability and growth over developmental time. 
-Our goal is to present this synthesis as a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work even as its assumptions are still the subject of rigorous evaluation.
+Our goal is to present this synthesis as a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work, even if its assumptions still require rigorous evaluation.
 
@@
 
-In physics, the "standard model" is a widely-accepted theory completed in the 1970s that describes all known elementary particles along with three of the four known fundamental forces in the universe.
-Although it is known to be incomplete and even in places incorrect, physics’ Standard Model nonetheless explains a wide variety of empirical phenomena, allowing scientists and engineers to model physical interactions with great precision and accuracy – and to design tests of the theory that inform its revision.
-Psychology in general has been criticized for lacking such formal theories to inform and drive empirical research [@muthukrishna2019problem], and we feel this critique can be applied fairly to the field of language acquisition.
+In physics, the "standard model" is a widely-accepted theory from the 1970s that describes all known elementary particles along with three of the four known fundamental forces in the universe. 
+Although it is known to be incomplete and even incorrect in places, physics’ Standard Model nonetheless explains a wide variety of empirical phenomena, allowing scientists and engineers to model physical interactions with great precision and accuracy and to design tests that inform theory revision.
+Psychology in general has been criticized for lacking such formal theories that inform and drive empirical research [@muthukrishna2019problem].
+We feel this critique can be applied fairly to the field of language acquisition.
 Although a substantial literature (that the current authors have contributed to) concerns the use of computational models to simulate acquisition, these models are often strikingly disconnected from the quantitative – not to mention qualitative – data gathered by most working acquisition researchers.
 
-But in fact, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
-Language input in the environment is processed in the moment and then accumulates via repeated exposure into learning. 
+In fact, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
+Language input is processed in the moment and then accumulates via repeated exposure, resulting in learning. 
 This framework underpins much of the broader policy discussion of individual variability in language learning and its links to environmental stimulation. 
-And this basic model does in fact correspond quite nicely to a class of computational models that has been explored in the word learning literature: accumulator models.
+This basic model does in fact correspond nicely to a class of computational models that has been explored in the word learning literature: accumulator models.
 
-Our aim here is to make this connection explicit, using item-response theory from psychometric testing as a framework for connecting between empirical data and previous accumulator models. 
+Our aim is to make this connection explicit, using item-response theory (IRT) from psychometric testing as a framework for connecting empirical data and previous accumulator models. 
 Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal verbal theories and evaluate them quantitatively [@vanrooij2020theory].
-Importantly, the goal is not to produce “the correct model” but rather to explore the ways that model assumptions lead to predictions. 
-In fact, the greatest value of such models is likely the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation (Tauber et al., 2016).
-
-In the remainder of this paper, we introduce accumulator models in more detail and then briefly present how they connect with IRT models. 
-We then review evidence consistent with this sketch of a “standard model”, as well as some empirical challenges for the future development.
+Importantly, the goal is not to produce “the correct model,” but rather to explore the ways that model assumptions lead to predictions. 
+In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation (Tauber et al., 2016).
 
@@
 We are interested in children’s vocabulary, which we define as the set of words[^2] they know. 
-Although vocabulary has sometimes been thought of as disconnected from other – often more “interesting” – parts of language learning, in fact the language system is tightly woven [@wordbankbook]. 
-Children’s vocabulary size predicts a huge proportion of the concurrent variance in their production of complex grammatical utterances and gestures and is predictively related to early speech perception [@bates1994developmental; @wordbankbook; @cristia2014predicting].
-Further, measures of children’s early language are highly correlated with one another both concurrently and predictively, suggesting a single latent variable – language ability – that is quite stable across early childhood [@bornstein2013language].
+Although vocabulary has sometimes been thought of as disconnected from other, often more “interesting,” parts of language learning, in fact the language system is tightly woven [@wordbankbook]. 
+Children’s vocabulary size predicts a substantial proportion of concurrent variance in their production of complex grammatical utterances and gestures and is predicted by early speech perception [@bates1994developmental; @wordbankbook; @cristia2014predicting].
+Further, measures of early language are highly correlated with one another, both concurrently and predictively, suggesting a single latent variable, i.e.
+language ability, that is quite stable across early childhood [@bornstein2013language].
 
@@
 These questions are relatively inconsequential in morphologically simple languages like English and Mandarin; yet they are quite a bit more difficult in morphologically more complex languages like Turkish or even agglutinating languages like Inuktitut.
-Here we sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating something – whether or not we call it a word per se, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words are.
+Here we sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating *something* – whether or not we call it a word *per se*, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words are.
 For some initial data bearing out this supposition by comparing Turkish with other languages, see @braginsky2019.
 
-In early theories of language acquisition, children’s environment was assumed to “trigger” specific innate developmental schemas [e.g., @gibson1994triggers]. 
-Driven by influential observations of the tight relationship between environmental input and children’s learning [e.g., @huttenlocher1991early; @hart1995], more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
-One such class of models is accumulator models, which assume that linguistic experiences with words accumulate in separate registers (pictured as “buckets” in Figure 2), with those registers exceeding a particular threshold being learned.
+In early theories, linguistic input was assumed to “trigger” specific innate developmental schemas [e.g., @gibson1994triggers]. 
+Driven by influential observations of the tight relationship between input and children’s learning [e.g., @huttenlocher1991early; @hart1995], more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
+One such class of models is *accumulator* models, which assume that linguistic experiences with words accumulate in separate registers (depicted as “buckets” in Figure 2), with those registers exceeding a particular threshold being learned.
 
 Formalized theories of accumulative vocabulary growth have already contributed significantly to theoretical debate. 
-For example, @mcmurray2007 elegantly demonstrated that children’s “vocabulary explosion”  – an acceleration in the growth of children’s vocabulary in their second year of life – can result from the steady accumulation of word tokens without any change in the environment or learning mechanism.
+For example, @mcmurray2007 elegantly demonstrated that children’s “vocabulary explosion” – an acceleration in vocabulary growth in the second year – can result from the steady accumulation of word tokens without changes in the environment or learning mechanism.
 Other work has examined similar models using more realistic distributions of words (i.e., Zipf-distributed), developmental change in learning mechanisms, and quantitative comparison to children’s aggregate vocabulary growth [@mayor2010; @hidaka2013; @mollica2017].
-Although this work is exciting, these models have not yet been used to make direct predictions about learning in individual children [@bergelson2020comprehension].
+Although this work is exciting, these models have not yet made direct predictions about learning in individual children [@bergelson2020comprehension].
 
 One difficulty in connecting such models to data is that the relevant variables are often measured in relative, rather than absolute, units. 
-This practice is common – for example, measures of intelligence are given on a standardized scale defined by population variability.
-Yet in early language, we have – relatively unusually in a psychological domain – access to absolute units of description. 
-We can count how many words a child hears and express this as a rate (e.g., words per hour; Cristia et al., 2019), and we can similarly estimate how many words they know [@wordbankbook]. 
+This practice is common, for example, measures of intelligence are given on a standardized scale defined by population variability. 
+Yet in early language, we have access to absolute units of description, relatively unusually in a psychological domain. 
+We can count how many words a child hears and express this as a rate (e.g., words per hour; Cristia et al., 2019), and we can similarly estimate how many words they know [e.g., by parents report of vocabulary size, @wordbankbook].
 These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
@@
 
-Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest, and numerous studies have reported associations between these variables [e.g., @hoff2003specificity; @weisleder2013].
-The direction of these associations is of course congruent with the standard model (and indeed is generally an inspiration). 
-But such correlations are also partially explained by a number of other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables, and their magnitude varies across studies [@wang2020meta].
-Though such studies are costly and difficult to conduct, randomized interventions are the gold standard for measuring causal effects of parental language input on language learning. 
-When such studies are conducted, they tend to show modest but reliable effects [e.g., @suskind2016project]. 
-These positive results provide support for a causal connection between language input and outcomes, but – because they only estimate the effect of the change in input on variation in outcome – they provide only a relative, rather than an absolute, estimate of how important input is to vocabulary learning.
+Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest. 
+Numerous studies have reported associations between these variables [e.g., @hoff2003specificity; @weisleder2013]. 
+The direction of these associations is, of course, congruent with the standard model (and indeed is generally an inspiration). 
+Moreover, such correlations are also partially explained by other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables, and their magnitude varies across studies [@wang2020meta].
+Randomized interventions are the gold standard for measuring causal effects of parental language input on language learning, although such studies are costly and difficult to conduct. 
+When such studies are conducted, they show modest but reliable effects [e.g., @suskind2016project], providing support for a causal connection between language input and outcomes. 
+Critically, however, these studies only estimate the effect of the change in input on variation in outcome, hence, providing only a relative, rather than an absolute, estimate of how important input is to vocabulary learning.
 
-Indeed, reasoning from first principles, you cannot learn the word “table” if you do not hear it. Input must be a predictor of learning. 
-Correlational studies do not assess this relationship fully – rather, they assess whether variation in input relates to variation in learning. 
-A second line of work that comes closer to assessing the importance of language input focuses not on differences between children but instead on differences between words. 
-These studies use regression models to predict which words are easier or more difficult, typically averaging across children [e.g., @goodman2008does; @braginsky2019]. 
+Reasoning from first principles, you cannot learn the word “table” if you do not hear it: input *must* predict learning. 
+Yet correlational studies do not assess this relationship fully, and instead assess whether *variation* in input relates to *variation* in learning. 
+Thus, a second line of work that comes closer to assessing the importance of language input focuses, not on differences between children, but on differences between words. 
+These studies use regression models to predict which *words* are easier or more difficult, averaging across children [e.g., @goodman2008does; @braginsky2019]. 
 Such studies show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
-Such models can also be used to assess the value of other proposals for learning mechanisms in the literature [e.g., @fourtassi2020growth].
+Such models are also used to evaluate other proposals for learning mechanisms [e.g., @fourtassi2020growth].
 
+<!--
 More broadly, this literature illustrates the value of a quantitative process model: under the standard model and its variants, we should be able to compute the expected relation between measures of input and vocabulary size.
-To the extent this model accounts for the data, it will support researchers’ current theory, but likely more informative will be the model’s pattern of misfit to the data, which should provide grist for the twin mills of theory development and empirical study.
+To the extent this model accounts for the data, it will support researchers’ current theory, but the model’s pattern of misfit to the data will likely be more informative, providing grist for the twin mills of theory development and empirical study.
+-->
 
@@
 
-As we have argued above, the core of the standard view of language learning is that it is a process of accumulation – individual experiences with words lead to their eventual acquisition. 
+The core of the standard view of language learning is that it is a process of accumulation – individual experiences with words lead to their eventual acquisition. 
 The more of these experiences a child receives, the faster their vocabulary grows. 

### 2021-03-10 paper/paper.Rmd
finished a round of edits: 3133 words

@@
 keywords          : "early language learning; language acquisition; vocabulary development"
-wordcount         : "3463"
+wordcount         : "3133"
 
@@
 However, both children and words vary: children may learn faster or slower, and words can be more or less difficult to learn. 
-Combining these ideas, the basic hypothesis is that a child’s vocabulary at a particular time should be predicted by their cumulative language exposure and learning rate, combined with the breadth and individual difficulties of the words that they have been exposed to.
+Combining these ideas, the basic hypothesis is that a child’s vocabulary at a particular time should be predicted by their cumulative language exposure and learning rate, combined with the breadth and individual difficulties of the words to which have been exposed.
 
 The key contribution of such a baseline theory is that it establishes the link between the observable rates of words that children hear and their eventual knowledge of those words. 
-For example, in principle the model could establish how many occurrences of the word “dog” are needed by the average child in order to start understanding (or producing) “dog.” 
+For example, in principle, the model could establish how many occurrences of the word “dog” are needed by the average child in order to start understanding (or producing) “dog.” 
 Moreover, given a child’s vocabulary at a certain age, the model would provide an estimate of the child’s language ability.
 
-This model outlines an approach that is very similar to Item-Response Theory [IRT; @embretson2013item], an influential and popular psychometric modeling framework for understanding the performance of test-takers (in our case language learners whose vocabulary is being assessed) as they are assessed with a particular set of test items (the particular words we ask about).
+This model outlines an approach similar to Item-Response Theory [IRT; @embretson2013item], an influential and popular psychometric modeling framework for understanding the performance of test-takers (in our case language learners) as they are assessed with a particular set of test items (particular words).
 In particular, IRT models jointly infer both how hard specific test items are and how accurate test takers are, recovering these latent parameters from the observed data. 
@@
 $d_j$ is the difficulty of word j), and $\theta_i$ is child $i$’s estimated latent language ability.
-Although in a typical IRT model both item difficulties and person abilities are standardized (assumed to be normally-distributed and centered on 0) and unit-free, in principle it is possible to map these scores to real-world distributions – as we would like to do – incorporating measured word frequencies and rates of children’s experienced input.
+In typical IRT models both item difficulties and person abilities are standardized (assumed to be normally-distributed and centered on 0) and unit-free. 
+However, in principle, it is possible to map these scores to real-world distributions incorporating measured word frequencies and rates of children’s experienced input. 
 With this mapping into absolute units, this kind of model can provide a quantitative linking hypothesis between measurements of input and learning. 
 Recognizing the parallels between the standard model desiderata and IRT theory, we can use standard extensions of the 1PL (Rasch) IRT model to include both person-level and item-level covariates. 
-Item-level (word-level) covariates of central interest in the standard model are estimates of word frequency in child-directed speech as well as lexical class [whether they are a noun, a predicate, or function word; @goodman2008does; @braginsky2019].
-Person-level covariates of interest in a standard model might include sex, socioeconomic status variables, or other demographic information.
+Item-level (word-level) covariates of central interest are estimates of word frequency in child-directed speech, as well as lexical class [e.g., noun, predicate, or function word; @goodman2008does; @braginsky2019].
+Person-level covariates of interest might include sex, socioeconomic status variables, or other demographic information.
 
@@
 Put simply, is a 2-year-old different than a 1-year-old, aside from having twice as much experience with each word? 
-This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is included as a person-level covariate, and one in which age is instead included as a scalar on the prevalence of each item (i.e., in which the frequency of “dog” for a 2-year-old is twice that of “dog” for a 1-year-old).
+This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is included as a person-level covariate, and one in which age is included as a scalar on the prevalence of each item (i.e., in which the frequency of “dog” for a 2-year-old is twice that of “dog” for a 1-year-old).
 Using the IRT framework to formalize theoretical questions and assess them via model comparison is a strength of this approach.
+
 Below we report the results of this model comparison, fit to parent-reported vocabulary data from 5,492 English-speaking children in Wordbank. 
 Each model estimated parameters for each child’s ability, and for the 680 words on the CDI Words and Sentences form. 
-As covariates we included the lexical class of each word (function words, adjective, noun, verb, or other), and the estimated number of monthly tokens per word a child is likely to hear (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech [@macwhinney2000childes].
-An average child in this model (with input rate based on reported empirical means) is expected to receive 1,200 words per hour, 12 hours each day, for a total of 438,300 tokens per month, of which 285,200 tokens (65.1%) are accounted for by tokens of CDI words.
+Covariates were the lexical class of each word (function word, adjective, noun, verb, or other), and the estimated number of monthly tokens per word a child hears (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech [@macwhinney2000childes].
+This model expects an average child to receive 1,200 words per hour, 12 hours each day, for a total of 438,300 tokens per month, of which 285,200 tokens (65.1%) are accounted for by tokens of CDI words.
 
 In Model 1, we included age as a person-level covariate, allowing for interactions of age (in months), lexical class, and monthly tokens per word. 
-In Model 2, instead of including the child’s age as a person-level covariate, we scaled the expected tokens per month for each word by the child’s age, yielding an estimate of their lifetime tokens per word, and also allowed for interactions with lexical class.
+Model 2 dropped the age covariate and instead scaled the expected tokens per month for each word by child age, yielding an estimate of lifetime tokens per word, and also allowed for interactions with lexical class.
 Model selection metrics (AIC and BIC) preferred the more complex Model 1, with a person-level age covariate separate from the item prevalence covariate. 
-Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, and prevalence (left), showing for example that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
-Moreover, the model predicts item-level acquisition curves (Figure 3, right), for example predicting that “ball” is learned earlier than “dog”, and that “go” is easier than “have.” 
+Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, and prevalence (left), revealing that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
+Moreover, the model predicts item-level acquisition curves (Figure 3, right): e.g., “ball” is learned earlier than “dog”, and “go” is easier than “have.” 
 Model 1 represents a new baseline standard model, which might be extended to include person-level groups such as SES and sex, as well as item-level covariates such as word length.
 
-Any model in this framework also estimates the latent language ability of each child in the dataset -- a value correlated with age, and with a child’s total vocabulary size as estimated by the CDI, but that may be better correlated with the child’s quantity/quality of input received, or with their processing speed: relations that will require new datasets with concurrent measures of these variables in order to distinguish.
-For one lesson we have learned in synthesizing the standard model framework is that based on current datasets, some of the theories are currently indistinguishable from one another, the models unidentifiable.
-Thus, we conclude by briefly discussing the next steps needed to test and build upon the standard model.
+These models also estimate a latent language ability for each child in the dataset. 
+However, this ability is no doubt correlated with age, sex (females have larger vocabularies at young ages), and likely with features related to input quality. 
+Estimating the contributions of these factors and determining whether/how language ability changes over developmental time is not possible with current datasets.
+Thus, we conclude by discussing the next steps needed to test and build upon the standard model.
 
@@
 Often, however, it is the *failure* of such a model to predict observed patterns of data that is most useful, as these failures point the way forward towards future refinements [@vanrooij2020theory]. 
-We discuss three potential failures here as signposts for ways forward.
+We discuss three potential failures here.
 
 **Leveraged learning and the role of processing**. As the model results above show, older children accumulate language from their input faster than younger children. 
-Why? One potential reason is that independent developmental or maturational factors simply make older children better at remembering their experiences. 
-Another possibility -- not mutually exclusive -– is that children *leverage* their knowledge of language to learn faster [@mitchell2009]. 
-This leverage idea is consistent with many proposed mechanisms for language learning.
-For example, if children reason about new vocabulary items by exclusion [@markman1988children], they could get better at learning new words just by having more potential items to exclude from consideration.
-Yet a third possibility –- again not mutually exclusive with the others -– is that children are getting better at using the words they know. 
-Several recent studies have found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words [e.g., @fernald2006picking; @marchman2016early].
+There are at least three non-mutually exclusive potential reasons for this.
+First, older children may be better at remembering their experiences due to independent developmental/maturational factors. 
+Another possibility is that children *leverage* their knowledge of language to learn faster through any of several proposed mechanisms [@mitchell2009]. 
+For example, if children reason about new words by exclusion [@markman1988children], they get better at learning new words just by having more potential items to exclude from consideration. 
+A third possibility is that children are getting better at using the words they know. 
+Several recent studies found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words [e.g., @fernald2006picking; @marchman2016early].
 Disentangling these distinct explanations will require careful measurement of both vocabulary and the potential mechanisms of leverage [cf. @lewis2020role].
 
-**Beyond vocabulary**.
-The "standard model" proposed here is a model of the accumulation of words, but language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings.
+**Beyond vocabulary**. The "standard model" proposed here models the accumulation of words.
+Yet, of course, language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings. 
 In early conceptions of language learning, syntax and morphology were thought of as distinct and unconnected, but this conception has not been borne out empirically. 
@@
 Some children might learn more from overheard speech and others might learn more from child-directed speech [@sperry2019]. 
-The kind of data necessary to test this assumption directly are only now being collected, for example in studies that rigorously track learning outcomes and amount of language input in diverse populations [e.g., comparisons between children growing up in low-income, rural, indigenous communities and those in higher-income Western contexts; @casillas2020].
+The kind of data necessary to test this assumption directly are only now being collected, for example, in studies that rigorously track learning outcomes and amount of language input in diverse populations [e.g., comparisons between children growing up in low-income, rural, indigenous communities and those in higher-income Western contexts; @casillas2020].
 

### 2021-03-11 paper/paper.Rmd
small edits

--- 2021-03-10 paper/paper.Rmd
@@
 
-These models also estimate a latent language ability for each child in the dataset. 
-However, this ability is no doubt correlated with age, sex (females have larger vocabularies at young ages), and likely with features related to input quality. 
+These models also estimate a latent language ability for each child, but currently do not specify whether/how this value changes over time. 
+However, this ability is no doubt correlated with age, as well as with features related to input quality. 
 Estimating the contributions of these factors and determining whether/how language ability changes over developmental time is not possible with current datasets.

### 2021-03-11 paper/paper.Rmd
mike update through first two sectons

@@
   A "standard model" is a theoretical framework that synthesizes observables into a quantitative consensus.
-We assess our progress towards such a synthesis for children’s early language learning.
-Several recent computational models of early vocabulary learning all share the intuition that individual words are learned through an accumulation of environmental input, a potential match with empirical work that emphasizes links between input and learning.
-However, models need to simultaneously account for variation in: 1) word difficulty, 2) children’s rate of exposure to words, and 3) children’s ability to integrate word occurrences.
-We show how these models relate to Item-Response Theory models, widely-used in psychometrics.
-Unfortunately, this formal connection also makes it clear that the currently available datasets cannot allow us to fully test the model, illustrating a critical need for theory in shaping new data collection.
-In particular, to compare across populations, we need to collect data that use absolute measurement units, like words heard per hour and total vocabulary size, rather than unitless measures like effect size.
-The assumptions of this nascent “standard model” already implicitly guide much work in the field, but making these assumptions explicit brings into sharp relief questions about the applicability of this model to diverse populations.
+Have we made progress towards this kind of synthesis for children’s early language learning?
+A number of computational models of early vocabulary learning assume that individual words are learned through an accumulation of environmental input.
+This assumption is also implicit in empirical work that emphasizes links between home language input and learning outcomes.
+However, models have typically focused on average performance, while empirical work has focused on variability.
+To capture individual variability, we relate accumulator models to Item-Response Theory models from psychometrics.
+Unfortunately, this formal connection also makes it clear that the currently available datasets cannot allow us to fully test these models, illustrating a critical need for theory in shaping new data collection.
+We end by considering the how this framework could shed light on to the role of language processing in acquistion, the connections of word learning to language learning more broadly, and the sources of variation in early language outcomes.
 
@@
 Early language learning is a key challenge for cognitive science: how do we go from speechless, wordless infants to children who can use language expressively and creatively? 
-A better understanding of this process has implications for both policy and technology. 
-Despite rapid progress in this area, early language is still often portrayed as mired in controversies, for example, around issues of innateness. 
-However, based on new theoretical and empirical work on the growth of vocabulary, and the deep connections between vocabulary and language more broadly, we see an emerging synthesis. 
-This synthesis provides a testable, quantitative model language learning, connecting individual experiences to variability and growth over developmental time. 
-Our goal is to present this synthesis as a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work, even if its assumptions still require rigorous evaluation.
+The field of early language is often portrayed as mired in controversies, for example, around issues of innateness. 
+We see a new synthesis emerging, however, based on new theoretical and empirical work on the growth of vocabulary. 
+Our goal is to present this synthesis as the beginnings of a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work, even if its assumptions still require rigorous evaluation.
 
@@
 We feel this critique can be applied fairly to the field of language acquisition.
-Although a substantial literature (that the current authors have contributed to) concerns the use of computational models to simulate acquisition, these models are often strikingly disconnected from the quantitative – not to mention qualitative – data gathered by most working acquisition researchers.
 
-In fact, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
-Language input is processed in the moment and then accumulates via repeated exposure, resulting in learning. 
+Although a substantial literature uses computational models to simulate acquisition, these models (ours included) are often only loosely connected to the quantitative data gathered by working acquisition researchers.
+Yet despite this lack of formal connections, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
+The core of this framework is that language input  accumulates via repeated exposure, resulting in learning.
 This framework underpins much of the broader policy discussion of individual variability in language learning and its links to environmental stimulation.
-This basic model does in fact correspond nicely to a class of computational models that has been explored in the word learning literature: accumulator models.
 
+This basic framework corresponds nicely with a class of computational models that has been explored in the word learning literature: accumulator models. 
 Our aim is to make this connection explicit, using item-response theory (IRT) from psychometric testing as a framework for connecting empirical data and previous accumulator models. 
-Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal verbal theories and evaluate them quantitatively [@vanrooij2020theory].
-Importantly, the goal is not to produce “the correct model,” but rather to explore the ways that model assumptions lead to predictions. 
-In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation (Tauber et al., 2016).
+Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal verbal theories and evaluate them quantitativelys.
+Importantly, the goal is not to produce 'the correct model," but rather to explore the ways that model assumptions lead to predictions about specific patterns of data. 
+In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation [@tauber2016].
 
-```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", fig.cap=c("A schematic of the “standard” relationships between variables assumed in the literature on early language learning.
+```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", fig.cap=c("A schematic of the "standard" relationships between variables assumed in the literature on early language learning.
 It is generally assumed that child-directed speech (red) is more valuable than overheard speech (blue)."), echo=FALSE}
@@
 
-# Theorizing About the Accumulation of Language
+# Accumulator Models Are An Important Formalism For Describing Word Learning
 
-We are interested in children’s vocabulary, which we define as the set of words[^2] they know. 
-Although vocabulary has sometimes been thought of as disconnected from other, often more “interesting,” parts of language learning, in fact the language system is tightly woven [@wordbankbook]. 
-Children’s vocabulary size predicts a substantial proportion of concurrent variance in their production of complex grammatical utterances and gestures and is predicted by early speech perception [@bates1994developmental; @wordbankbook; @cristia2014predicting].
-Further, measures of early language are highly correlated with one another, both concurrently and predictively, suggesting a single latent variable, i.e.
-language ability, that is quite stable across early childhood [@bornstein2013language].
-
-[^2]: What makes a word is its own separate and difficult question, especially as we look across languages.
-Are “dog” and “dogs” two separate words?
-If not, what about “mouse” and “mice”?
+We are interested in children’s vocabulary, which we define as the set of words they know.^[
+What makes a word is its own separate and difficult question, especially as we look across languages. Are "dog" and "dogs" two separate words? 
+If not, what about "mouse" and "mice"?
 These questions are relatively inconsequential in morphologically simple languages like English and Mandarin; yet they are quite a bit more difficult in morphologically more complex languages like Turkish or even agglutinating languages like Inuktitut.
-Here we sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating *something* – whether or not we call it a word *per se*, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words are.
-For some initial data bearing out this supposition by comparing Turkish with other languages, see @braginsky2019.
-
-In early theories, linguistic input was assumed to “trigger” specific innate developmental schemas [e.g., @gibson1994triggers]. 
-Driven by influential observations of the tight relationship between input and children’s learning [e.g., @huttenlocher1991early; @hart1995], more recent models have emphasized the accumulation of specific linguistic experiences in driving learning and generalization.
-One such class of models is *accumulator* models, which assume that linguistic experiences with words accumulate in separate registers (depicted as “buckets” in Figure 2), with those registers exceeding a particular threshold being learned.
+We sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating *something* – whether or not we call it a word *per se*, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words are.
+For some initial data bearing out this supposition by comparing Turkish with other languages, see @braginsky2019.]
+Driven by influential observations of the strong relationship between language input and children’s word learning [e.g., @huttenlocher1991early; @hart1995], computational modelers have created a range of hypotheses about how linguistic experiences drive learning.
+One class of models is *accumulator* models, which assume that linguistic experiences with words accumulate in separate registers (depicted as "buckets" in Figure 2), with those registers exceeding a particular threshold being learned.
 
 Formalized theories of accumulative vocabulary growth have already contributed significantly to theoretical debate. 
-For example, @mcmurray2007 elegantly demonstrated that children’s “vocabulary explosion” – an acceleration in vocabulary growth in the second year – can result from the steady accumulation of word tokens without changes in the environment or learning mechanism.
+For example, @mcmurray2007 elegantly demonstrated that children’s "vocabulary explosion" – an acceleration in vocabulary growth in the second year – can result from the steady accumulation of word tokens without changes in the environment or learning mechanism.
 Other work has examined similar models using more realistic distributions of words (i.e., Zipf-distributed), developmental change in learning mechanisms, and quantitative comparison to children’s aggregate vocabulary growth [@mayor2010; @hidaka2013; @mollica2017].
@@
 Yet in early language, we have access to absolute units of description, relatively unusually in a psychological domain. 
-We can count how many words a child hears and express this as a rate (e.g., words per hour; Cristia et al., 2019), and we can similarly estimate how many words they know [e.g., by parents report of vocabulary size, @wordbankbook].
+We can count how many words a child hears and express this as a rate [e.g., words per hour; @cristia2019], and we can similarly estimate how many words they know [e.g., by parents report of vocabulary size, @wordbankbook].
 These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
@@
 
-# Parsing the Literature Using the Standard Model
+# Accumulator Models are Presupposed in the Empirical Literature
 
 Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest. 
-Numerous studies have reported associations between these variables [e.g., @hoff2003specificity; @weisleder2013]. 
-The direction of these associations is, of course, congruent with the standard model (and indeed is generally an inspiration). 
-Moreover, such correlations are also partially explained by other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables, and their magnitude varies across studies [@wang2020meta].
+Numerous studies have reported positive associations between these variables [e.g., @hoff2003specificity; @weisleder2013] - as predicted by accumulator models - though their magnitude varies across studies [@wang2020meta].
+Such correlations are also partially moderated by other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables.
 Randomized interventions are the gold standard for measuring causal effects of parental language input on language learning, although such studies are costly and difficult to conduct. 
@@
 
-Reasoning from first principles, you cannot learn the word “table” if you do not hear it: input *must* predict learning. 
+Reasoning from first principles, you cannot learn the word "table if you do not hear it: input *must* predict learning. 
 Yet correlational studies do not assess this relationship fully, and instead assess whether *variation* in input relates to *variation* in learning. 
@@
 Such studies show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
-Such models are also used to evaluate other proposals for learning mechanisms [e.g., @fourtassi2020growth].
+This general finding provides convergent support for accumulator models, absent the confound of differences between individuals.
 
@@
 The key contribution of such a baseline theory is that it establishes the link between the observable rates of words that children hear and their eventual knowledge of those words. 
-For example, in principle, the model could establish how many occurrences of the word “dog” are needed by the average child in order to start understanding (or producing) “dog.” 
+For example, in principle, the model could establish how many occurrences of the word "dog" are needed by the average child in order to start understanding (or producing) "dog." 
 Moreover, given a child’s vocabulary at a certain age, the model would provide an estimate of the child’s language ability.
@@
 Put simply, is a 2-year-old different than a 1-year-old, aside from having twice as much experience with each word? 
-This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is included as a person-level covariate, and one in which age is included as a scalar on the prevalence of each item (i.e., in which the frequency of “dog” for a 2-year-old is twice that of “dog” for a 1-year-old).
+This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is included as a person-level covariate, and one in which age is included as a scalar on the prevalence of each item (i.e., in which the frequency of "dog" for a 2-year-old is twice that of "dog" for a 1-year-old).
 Using the IRT framework to formalize theoretical questions and assess them via model comparison is a strength of this approach.
@@
 Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, and prevalence (left), revealing that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
-Moreover, the model predicts item-level acquisition curves (Figure 3, right): e.g., “ball” is learned earlier than “dog”, and “go” is easier than “have.” 
+Moreover, the model predicts item-level acquisition curves (Figure 3, right): e.g., "ball" is learned earlier than "dog", and "go" is easier than "have." 
 Model 1 represents a new baseline standard model, which might be extended to include person-level groups such as SES and sex, as well as item-level covariates such as word length.
@@
 
-```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from the fitted English “standard model” as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 words on the CDI (right)."), echo=FALSE}
+```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from the fitted English "standard model" as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 words on the CDI (right)."), echo=FALSE}
 knitr::include_graphics("./figs/fig3_predicted_age_LC_and_items.pdf")
@@
 In early conceptions of language learning, syntax and morphology were thought of as distinct and unconnected, but this conception has not been borne out empirically. 
-Instead, evidence from large-scale data shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@wordbankbook].
+Instead, evidence from large-scale data shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language,@wordbankbook].
 Theoretically, accumulator models like the standard model are generic models of skill acquisition. 
@@
 By situating this model in a common psychometric framework, we show how it can be used to connect to large-scale, cross-linguistic datasets. 
-This “standard model" synthesizes measures of language input and vocabulary growth, allowing us to formalize, test, and iteratively improve our theoretical understanding of early language acquisition.
+This "standard model" synthesizes measures of language input and vocabulary growth, allowing us to formalize, test, and iteratively improve our theoretical understanding of early language acquisition.
 Moreover, this formalization allows us to identify specific gaps in current empirical approaches that must be closed to inform future theory development.

### 2021-03-11 paper/paper.Rmd
caption fix

@@
 
-```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", fig.cap=c("A schematic of the "standard" relationships between variables assumed in the literature on early language learning.
+```{r, out.width="0.95\\linewidth", include=TRUE, fig.align="center", fig.cap=c("A schematic of the standard relationships between variables assumed in the literature on early language learning.
 It is generally assumed that child-directed speech (red) is more valuable than overheard speech (blue)."), echo=FALSE}
@@
 
-```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from the fitted English "standard model" as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 words on the CDI (right)."), echo=FALSE}
+```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from a fitted English IRT model as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 words on the CDI (right)."), echo=FALSE}
 knitr::include_graphics("./figs/fig3_predicted_age_LC_and_items.pdf")

### 2021-03-11 paper/paper.Rmd
VM - another pass

@@
 Although it is known to be incomplete and even incorrect in places, physics’ Standard Model nonetheless explains a wide variety of empirical phenomena, allowing scientists and engineers to model physical interactions with great precision and accuracy and to design tests that inform theory revision.
-Psychology in general has been criticized for lacking such formal theories that inform and drive empirical research [@muthukrishna2019problem]. 
+Psychology, in general, has been criticized for lacking such formal theories that inform and drive empirical research [@muthukrishna2019problem]. 
 We feel this critique can be applied fairly to the field of language acquisition.
@@
 Our aim is to make this connection explicit, using item-response theory (IRT) from psychometric testing as a framework for connecting empirical data and previous accumulator models. 
-Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal verbal theories and evaluate them quantitativelys.
-Importantly, the goal is not to produce 'the correct model," but rather to explore the ways that model assumptions lead to predictions about specific patterns of data. 
+Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal theories and evaluate them quantitatively.
+Importantly, the goal is not to produce the "correct" model, but rather to explore the ways that model assumptions lead to predictions about specific patterns of data. 
 In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation [@tauber2016].
@@
 These questions are relatively inconsequential in morphologically simple languages like English and Mandarin; yet they are quite a bit more difficult in morphologically more complex languages like Turkish or even agglutinating languages like Inuktitut.
-We sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating *something* – whether or not we call it a word *per se*, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words are.
+We sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating *something* – whether or not we call it a word *per se*, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words.
 For some initial data bearing out this supposition by comparing Turkish with other languages, see @braginsky2019.]
 Driven by influential observations of the strong relationship between language input and children’s word learning [e.g., @huttenlocher1991early; @hart1995], computational modelers have created a range of hypotheses about how linguistic experiences drive learning.
-One class of models is *accumulator* models, which assume that linguistic experiences with words accumulate in separate registers (depicted as "buckets" in Figure 2), with those registers exceeding a particular threshold being learned.
+One class of models is *accumulator* models, which assume that linguistic experiences with words accumulate in separate registers (depicted as "buckets" in Figure 2), and that those registers exceeding a particular threshold are learned.
 
@@
 One difficulty in connecting such models to data is that the relevant variables are often measured in relative, rather than absolute, units. 
-This practice is common, for example, measures of intelligence are given on a standardized scale defined by population variability. 
+This practice is common. For example, measures of intelligence are given on a standardized scale defined by population variability. 
 Yet in early language, we have access to absolute units of description, relatively unusually in a psychological domain. 
-We can count how many words a child hears and express this as a rate [e.g., words per hour; @cristia2019], and we can similarly estimate how many words they know [e.g., by parents report of vocabulary size, @wordbankbook].
+We can count how many words a child hears and express this as a rate [e.g., words per hour; @cristia2019], and we can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size, @wordbankbook].
 These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
@@
 
-Reasoning from first principles, you cannot learn the word "table if you do not hear it: input *must* predict learning. 
-Yet correlational studies do not assess this relationship fully, and instead assess whether *variation* in input relates to *variation* in learning. 
+Reasoning from first principles, you cannot learn the word "table" if you do not hear it: input *must* predict learning. 
+Yet correlational studies do not assess this relation fully, and instead assess whether *variation* in input relates to *variation* in learning. 
 Thus, a second line of work that comes closer to assessing the importance of language input focuses, not on differences between children, but on differences between words.
@@
 
-This model outlines an approach similar to Item-Response Theory [IRT; @embretson2013item], an influential and popular psychometric modeling framework for understanding the performance of test-takers (in our case language learners) as they are assessed with a particular set of test items (particular words).
-In particular, IRT models jointly infer both how hard specific test items are and how accurate test takers are, recovering these latent parameters from the observed data. 
+This model outlines an approach similar to Item-Response Theory [IRT; @embretson2013item], an influential and popular psychometric modeling framework for understanding the performance of test takers (in our case language learners) as they are assessed with a particular set of test items (particular words).
+In particular, IRT models jointly infer both the difficulty of specific test items and the accuracy of specific test takers, recovering these latent parameters from the observed data. 
 In the basic Rasch (or 1-parameter logistic) IRT model, a person i responds correctly to item j with probability determined by their ability ($\theta_i$) and the difficulty of item j ($d_j$):
@@
 
-These parameters can easily be mapped onto the kind of accumulator model we have been describing: items are words (e.g.
+These parameters can easily be mapped onto the accumulator model we have been describing: items are words (e.g.
 $d_j$ is the difficulty of word j), and $\theta_i$ is child $i$’s estimated latent language ability.
-In typical IRT models both item difficulties and person abilities are standardized (assumed to be normally-distributed and centered on 0) and unit-free. 
+In typical IRT models, both item difficulties and person abilities are standardized (assumed to be normally-distributed and centered on 0) and unit-free. 
 However, in principle, it is possible to map these scores to real-world distributions incorporating measured word frequencies and rates of children’s experienced input. 
 With this mapping into absolute units, this kind of model can provide a quantitative linking hypothesis between measurements of input and learning. 
-Recognizing the parallels between the standard model desiderata and IRT theory, we can use standard extensions of the 1PL (Rasch) IRT model to include both person-level and item-level covariates. 
+Recognizing the parallels between the standard model desiderata and IRT theory, we can use standard extensions of the 1PL (Rasch) model to include both item-level and person-level covariates. 
 Item-level (word-level) covariates of central interest are estimates of word frequency in child-directed speech, as well as lexical class [e.g., noun, predicate, or function word; @goodman2008does; @braginsky2019].
@@
 Put simply, is a 2-year-old different than a 1-year-old, aside from having twice as much experience with each word? 
-This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is included as a person-level covariate, and one in which age is included as a scalar on the prevalence of each item (i.e., in which the frequency of "dog" for a 2-year-old is twice that of "dog" for a 1-year-old).
+This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is a person-level covariate, and one in which age is a scalar on the prevalence of each item (i.e., the frequency of "dog" for a 2-year-old is twice that of a 1-year-old).
 Using the IRT framework to formalize theoretical questions and assess them via model comparison is a strength of this approach.
 
-Below we report the results of this model comparison, fit to parent-reported vocabulary data from 5,492 English-speaking children in Wordbank. 
-Each model estimated parameters for each child’s ability, and for the 680 words on the CDI Words and Sentences form. 
-Covariates were the lexical class of each word (function word, adjective, noun, verb, or other), and the estimated number of monthly tokens per word a child hears (range: 0.2 - 19,286 tokens), based on estimates from the CHILDES corpus of child-directed speech [@macwhinney2000childes].
-This model expects an average child to receive 1,200 words per hour, 12 hours each day, for a total of 438,300 tokens per month, of which 285,200 tokens (65.1%) are accounted for by tokens of CDI words.
+We report the results of this model comparison, fit to parent-reported vocabulary data from 5,492 English-speaking children in Wordbank. 
+Each model estimated parameters for each child’s ability, and for the 680 words on the CDI Words and Sentences form [@fenson2007]. 
+Covariates were each word's lexical class (function word, adjective, noun, verb, or other) and the estimated number of monthly tokens a child hears (range: 0.2 - 19,286 tokens), based on child-directed speech from CHILDES [@macwhinney2000childes].
+This model expects an average child to receive 1,200 words/hour, 12 hours/day, for a total of 438,300 tokens/month, of which 285,200 tokens (65.1%) are  tokens of CDI words.
 
 In Model 1, we included age as a person-level covariate, allowing for interactions of age (in months), lexical class, and monthly tokens per word. 
-Model 2 dropped the age covariate and instead scaled the expected tokens per month for each word by child age, yielding an estimate of lifetime tokens per word, and also allowed for interactions with lexical class.
-Model selection metrics (AIC and BIC) preferred the more complex Model 1, with a person-level age covariate separate from the item prevalence covariate. 
-Predicted acquisition curves from Model 1 are shown in Figure 3, broken down by age, lexical class, and prevalence (left), revealing that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
-Moreover, the model predicts item-level acquisition curves (Figure 3, right): e.g., "ball" is learned earlier than "dog", and "go" is easier than "have." 
+Model 2 dropped the age covariate and instead scaled the expected tokens/month for each word by child age, yielding an estimate of lifetime tokens per word and also allowed for interactions with lexical class.
+Model selection metrics (AIC and BIC) preferred the more complex Model 1. 
+Predicted acquisition curves from Model 1 are shown in Figure 3 by age, lexical class, and prevalence (left), revealing that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
+Moreover, the model predicts item-level acquisition curves (Figure 3, right): e.g., "ball" is learned earlier than "dog" and "go" is easier than "have." 
 Model 1 represents a new baseline standard model, which might be extended to include person-level groups such as SES and sex, as well as item-level covariates such as word length.
@@
 
-The goal of a computational theory like the "standard model" described here is to describe the quantitative predictions that arise from a set of assumptions. 
+The goal of a computational theory like the "standard model" is to describe the quantitative predictions that arise from a set of assumptions. 
 Often, however, it is the *failure* of such a model to predict observed patterns of data that is most useful, as these failures point the way forward towards future refinements [@vanrooij2020theory]. 
@@
 
-**Leveraged learning and the role of processing**. As the model results above show, older children accumulate language from their input faster than younger children. 
+**Leveraged learning and the role of processing**. As the model results show, older children accumulate language from their input faster than younger children. 
 There are at least three non-mutually exclusive potential reasons for this.
@@
 Yet, of course, language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings. 
-In early conceptions of language learning, syntax and morphology were thought of as distinct and unconnected, but this conception has not been borne out empirically. 
-Instead, evidence from large-scale data shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language,@wordbankbook].
+In early views, syntax and morphology were conceptualized as distinct and unconnected, but this conception has not been borne out empirically. 
+Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language,@wordbankbook].
 Theoretically, accumulator models like the standard model are generic models of skill acquisition. 
-Given the perspective that language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards extensions of the standard model to the accumulation of broader units of language like constructions.
+If language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards extensions of the standard model to the accumulation of broader units of language-like constructions.
 
 **A theory for understanding acquisition in diverse contexts**.
-The "standard model" formalizes an implict assumption of the literature: namely, that early language learning follows the same general model for children in radically different circumstances.
+The "standard model" formalizes an implict assumption: namely, that early language learning follows the same general model for all children, even those in radically different circumstances.
 Yet this assumption could very well be false. 
 Some children might learn more from overheard speech and others might learn more from child-directed speech [@sperry2019]. 
-The kind of data necessary to test this assumption directly are only now being collected, for example, in studies that rigorously track learning outcomes and amount of language input in diverse populations [e.g., comparisons between children growing up in low-income, rural, indigenous communities and those in higher-income Western contexts; @casillas2020].
+The kind of data necessary to test this assumption directly are only now being collected, for example, in studies that rigorously track learning outcomes and amount of language input in diverse populations [e.g., comparisons between children in low-income, rural, indigenous communities and those in higher-income Western contexts; @casillas2020].
 
@@
 
-An implicit theory drives much research and policy-making on early language acquisition: early language accumulates through a set of discrete experiences with individual words. 
+An implicit theory drives much research and policy-making on early language acquisition: early language accumulates through discrete experiences with individual words. 
 The more experiences, the faster the words are learned. 

### 2021-03-11 paper/paper.Rmd
fixed email address

@@
     address       : "450 Jane Stanford Way, Stanford, CA 94305"
-    email         : "mcfrank@stanford.edu"
+    email         : "marchman@stanford.edu"
   - name          : "Michael C. Frank"

### 2021-03-13 paper/paper.Rmd
last refs fixed; 32 refs, 2900 words

@@
 keywords          : "early language learning; language acquisition; vocabulary development"  
-wordcount         : "3133"
+wordcount         : "2896"
 
@@
 Importantly, the goal is not to produce the "correct" model, but rather to explore the ways that model assumptions lead to predictions about specific patterns of data. 
-In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation [@tauber2016].
+In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation [@tauber2017bayesian].
 
@@
 Yet in early language, we have access to absolute units of description, relatively unusually in a psychological domain. 
-We can count how many words a child hears and express this as a rate [e.g., words per hour; @cristia2019], and we can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size, @wordbankbook].
+We can count how many words a child hears and express this as a rate [e.g., words per hour; @cristia2020accuracy], and we can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size, @wordbankbook].
 These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
@@
 In early views, syntax and morphology were conceptualized as distinct and unconnected, but this conception has not been borne out empirically. 
-Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language,@wordbankbook].
+Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language; @wordbankbook].
 Theoretically, accumulator models like the standard model are generic models of skill acquisition. 

### 2021-03-18 paper/paper.Rmd
in progress updates to model section

@@
 ---
-title             : "Towards a \"Standard Model\" of Early Language Learning"
+title             : "Toward a \"Standard Model\" of Early Language Learning"
 shorttitle        : "Standard Model of Language Learning"
@@
 However, models have typically focused on average performance, while empirical work has focused on variability.
-To capture individual variability, we relate accumulator models to Item-Response Theory models from psychometrics.
-Unfortunately, this formal connection also makes it clear that the currently available datasets cannot allow us to fully test these models, illustrating a critical need for theory in shaping new data collection.
-We end by considering the how this framework could shed light on to the role of language processing in acquistion, the connections of word learning to language learning more broadly, and the sources of variation in early language outcomes.
+To capture individual variability, we relate the tradition of research on accumulator models to Item-Response Theory models from psychometrics.
+This formal connection also makes it clear that the currently available datasets cannot allow us to fully test these models, illustrating a critical need for theory in shaping new data collection.
 
@@
 We see a new synthesis emerging, however, based on new theoretical and empirical work on the growth of vocabulary. 
-Our goal is to present this synthesis as the beginnings of a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and guide to future work, even if its assumptions still require rigorous evaluation.
+Our goal is to present this synthesis as the beginnings of a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and should guide future work, even if its assumptions still require rigorous evaluation.
 
-[^1]: Note that "standard" here is not meant to imply that it is normative, but rather that the model is an accepted starting point for a description of early language learning.
+[^1]: Note that "standard" here is not meant to imply that the model is a normative prescription, but rather that the model is an accepted starting point for a description of early language learning.
 
@@
 Although a substantial literature uses computational models to simulate acquisition, these models (ours included) are often only loosely connected to the quantitative data gathered by working acquisition researchers.
-Yet despite this lack of formal connections, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
+Despite this lack of formal connections, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
 The core of this framework is that language input accumulates via repeated exposure, resulting in learning.
@@
 
-This basic framework corresponds nicely with a class of computational models that has been explored in the word learning literature: accumulator models. 
-Our aim is to make this connection explicit, using item-response theory (IRT) from psychometric testing as a framework for connecting empirical data and previous accumulator models. 
+This basic framework corresponds nicely with a group of computational models that has been explored in the word learning literature, which we refer to jointly as "accumulator models" [@mcmurray2007;@hidaka2013;@mollica2017].
+Our aim is to make an explicit connection between these models and the broader but less formal discussion of the role of language input in learning. 
+Since discussions of language input are typically focused on variation between children, we use item-response theory (IRT) from psychometric testing as a framework for connecting empirical data with     accumulator models.
 Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal theories and evaluate them quantitatively.
@@
 Driven by influential observations of the strong relationship between language input and children’s word learning [e.g., @huttenlocher1991early; @hart1995], computational modelers have created a range of hypotheses about how linguistic experiences drive learning.
-One class of models is *accumulator* models, which assume that linguistic experiences with words accumulate in separate registers (depicted as "buckets" in Figure 2), and that those registers exceeding a particular threshold are learned.
+Accumulator models, which we focus on here, assume that linguistic experiences with words accumulate in separate registers (depicted as "buckets" in Figure 2), and that those registers that exceed a particular threshold are learned.
 
-Formalized theories of accumulative vocabulary growth have already contributed significantly to theoretical debate. 
-For example, @mcmurray2007 elegantly demonstrated that children’s "vocabulary explosion" – an acceleration in vocabulary growth in the second year – can result from the steady accumulation of word tokens without changes in the environment or learning mechanism.
+Accumulator models have already contributed significantly to several theoretical issues. 
+For example, @mcmurray2007 elegantly demonstrated that children’s "vocabulary explosion" -- an acceleration in vocabulary growth in the second year -- can result from the steady accumulation of word tokens without changes in the environment or learning mechanism.
 Other work has examined similar models using more realistic distributions of words (i.e., Zipf-distributed), developmental change in learning mechanisms, and quantitative comparison to children’s aggregate vocabulary growth [@mayor2010; @hidaka2013; @mollica2017].
@@
 One difficulty in connecting such models to data is that the relevant variables are often measured in relative, rather than absolute, units. 
-This practice is common. For example, measures of intelligence are given on a standardized scale defined by population variability. 
-Yet in early language, we have access to absolute units of description, relatively unusually in a psychological domain. 
-We can count how many words a child hears and express this as a rate [e.g., words per hour; @cristia2020accuracy], and we can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size, @wordbankbook].
+This practice is common. 
+For example, measures of intelligence are given on a standardized scale defined by population variability. 
+Yet in early language -- unusually in a psychological domain -- we have access to absolute units.
+We can count how many words a child hears and express this number as a rate [e.g., words per hour; @cristia2020accuracy]. 
+We can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size, @wordbankbook]. 
 These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
-As we describe below, however, few datasets have measurements of the relevant variables. 
-This may be a case where theory outstrips data and further data collection efforts are necessary.
+Few datasets have measurements of the relevant variables, however. 
+Thus, one upshot of our argument here is that future data collection efforts should attempt whenever possible to report measurements in absolute units.
 
@@
 Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest. 
-Numerous studies have reported positive associations between these variables [e.g., @hoff2003specificity; @weisleder2013] - as predicted by accumulator models - though their magnitude varies across studies [@wang2020meta].
+Numerous studies have reported positive associations between these variables [e.g., @hoff2003specificity; @weisleder2013] -- as predicted by accumulator models -- though their magnitude varies across studies [@wang2020meta].
 Such correlations are also partially moderated by other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables.
@@
 These studies use regression models to predict which *words* are easier or more difficult, averaging across children [e.g., @goodman2008does; @braginsky2019]. 
-Such studies show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
+They typically show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
 This general finding provides convergent support for accumulator models, absent the confound of differences between individuals.
@@
 
-# Toward a "Standard" Accumulator Model
+# Connecting Accumulator Models with Psychometric Models
 
-The core of the standard view of language learning is that it is a process of accumulation – individual experiences with words lead to their eventual acquisition. 
+The core of the view that we are describing is that language learning is a process of accumulation. 
+Individual experiences with words lead to their eventual acquisition. 
 The more of these experiences a child receives, the faster their vocabulary grows. 
 However, both children and words vary: children may learn faster or slower, and words can be more or less difficult to learn. 
-Combining these ideas, the basic hypothesis is that a child’s vocabulary at a particular time should be predicted by their cumulative language exposure and learning rate, combined with the breadth and individual difficulties of the words to which have been exposed.
+Combining these ideas, the basic hypothesis is that a child’s vocabulary at a particular time should be predicted by their cumulative language exposure and learning rate, combined with the breadth of the sample of words to which have been exposed and those words' individual difficulties.
 
-The key contribution of such a baseline theory is that it establishes the link between the observable rates of words that children hear and their eventual knowledge of those words. 
-For example, in principle, the model could establish how many occurrences of the word "dog" are needed by the average child in order to start understanding (or producing) "dog." 
-Moreover, given a child’s vocabulary at a certain age, the model would provide an estimate of the child’s language ability.
-
-This model outlines an approach similar to Item-Response Theory [IRT; @embretson2013item], an influential and popular psychometric modeling framework for understanding the performance of test takers (in our case language learners) as they are assessed with a particular set of test items (particular words).
-In particular, IRT models jointly infer both the difficulty of specific test items and the accuracy of specific test takers, recovering these latent parameters from the observed data. 
-In the basic Rasch (or 1-parameter logistic) IRT model, a person i responds correctly to item j with probability determined by their ability ($\theta_i$) and the difficulty of item j ($d_j$):
+This hypothesis describes an approach similar to Item-Response Theory [IRT; @embretson2013item], an influential and popular psychometric modeling framework. 
+IRT is commonly used for estimating the ability of "test takers" (in our case language learners) as they are assessed with a particular set of "test items" (particular words). 
+IRT models jointly infer both the difficulty of specific test items and the accuracy of specific test takers, recovering these latent parameters from the observed data. 
+In the basic Rasch (or 1-parameter logistic) IRT model, a person $i$ responds correctly to item $j$ with probability determined by their ability ($\theta_i$) and the difficulty of item $j$ ($d_j$):
 
@@
 These parameters can easily be mapped onto the accumulator model we have been describing: items are words (e.g.
-$d_j$ is the difficulty of word j), and $\theta_i$ is child $i$’s estimated latent language ability.
+$d_j$ is the difficulty of word $j$), and $\theta_i$ is child $i$’s estimated latent language ability.
+
 In typical IRT models, both item difficulties and person abilities are standardized (assumed to be normally-distributed and centered on 0) and unit-free. 
@@
 
-This model as described above formalizes an important and testable claim: is the amount of language experience the important predictor of learning for a child, or are there other age-related factors beyond experience?
+# Comparing Model Variants as a Method for Evaluating Theories
+
+Using the well-known IRT framework to formalize theoretical questions and assess them via model comparison is a benefit over more ad-hoc frameworks. 
+We give one example here. 
+We can use this framework to test whether the amount of language experience is the important predictor of learning for a child or whether there are other age-related factors beyond experience. 
 Put simply, is a 2-year-old different than a 1-year-old, aside from having twice as much experience with each word? 
-This theoretical question can be answered by simple model comparison between two versions of the standard model, one in which age is a person-level covariate, and one in which age is a scalar on the prevalence of each item (i.e., the frequency of "dog" for a 2-year-old is twice that of a 1-year-old).
-Using the IRT framework to formalize theoretical questions and assess them via model comparison is a strength of this approach.
+This theoretical question can be answered by simple model comparison between two related models.
+
+In Model 1, we included age as a person-level covariate, allowing for interactions of age, lexical class, and monthly tokens per word. 
+Model 2 dropped the age covariate and instead scaled the expected tokens/month for each word by child age, yielding an estimate of lifetime tokens per word and also allowed for interactions with lexical class.
+
+The first includes age is a person-level covariate, indicating 
+The second includes age as a scalar multiplier on the prevalence of each item (i.e., the frequency of "dog" for a 2-year-old is twice that of a 1-year-old).
 
@@
 
-In Model 1, we included age as a person-level covariate, allowing for interactions of age (in months), lexical class, and monthly tokens per word. 
-Model 2 dropped the age covariate and instead scaled the expected tokens/month for each word by child age, yielding an estimate of lifetime tokens per word and also allowed for interactions with lexical class.
 Model selection metrics (AIC and BIC) preferred the more complex Model 1. 

### 2021-03-19 paper/paper.Rmd
up through model variants

@@
 To capture individual variability, we relate the tradition of research on accumulator models to Item-Response Theory models from psychometrics.
-This formal connection also makes it clear that the currently available datasets cannot allow us to fully test these models, illustrating a critical need for theory in shaping new data collection.
+This formal connection also makes it clear that the currently available datasets cannot allow us to fully test these models, illustrating a critical need for theory in shaping new data collection and creating and testing an eventual "standard model."
 
@@
 
-[^1]: Note that "standard" here is not meant to imply that the model is a normative prescription, but rather that the model is an accepted starting point for a description of early language learning.
+[^1]: "Standard" here is not meant to imply a normative prescription, but rather that the model is an accepted starting point for a description of early language learning.
 
@@
 In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation [@tauber2017bayesian].
+Perhaps such iterative work will lead to a true "standard model" of language learning.
 
@@
 We are interested in children’s vocabulary, which we define as the set of words they know.^[
-What makes a word is its own separate and difficult question, especially as we look across languages. Are "dog" and "dogs" two separate words? 
+What makes a word is its own separate and difficult question, especially as we look across languages. 
+Are "dog" and "dogs" two separate words? 
 If not, what about "mouse" and "mice"? 
-These questions are relatively inconsequential in morphologically simple languages like English and Mandarin; yet they are quite a bit more difficult in morphologically more complex languages like Turkish or even agglutinating languages like Inuktitut.
-We sidestep this issue by noting that the Turkish or Inuktitut learner is accumulating *something* – whether or not we call it a word *per se*, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics and information sources as words.
-For some initial data bearing out this supposition by comparing Turkish with other languages, see @braginsky2019.]
+These questions are relatively inconsequential in morphologically simple languages like English and Mandarin but much more difficult in morphologically complex languages like Turkish or agglutinating languages like Inuktitut.
+Yet the Turkish or Inuktitut learner is accumulating *something*. 
+Whether or not we call it a word *per se*, our hypothesis is that this unit is being accumulated and that its accumulation will be subject to many of the same dynamics as words.]
 Driven by influential observations of the strong relationship between language input and children’s word learning [e.g., @huttenlocher1991early; @hart1995], computational modelers have created a range of hypotheses about how linguistic experiences drive learning.
-Accumulator models, which we focus on here, assume that linguistic experiences with words accumulate in separate registers (depicted as "buckets" in Figure 2), and that those registers that exceed a particular threshold are learned.
+Accumulator models, which we focus on here as the foundation for a potential "standard model", assume that linguistic experiences with words accumulate in separate registers (depicted as "buckets" in Figure 2), and that those registers that exceed a particular threshold are learned.
 
@@
 
-Reasoning from first principles, you cannot learn the word "table" if you do not hear it: input *must* predict learning. 
-Yet correlational studies do not assess this relation fully, and instead assess whether *variation* in input relates to *variation* in learning. 
-Thus, a second line of work that comes closer to assessing the importance of language input focuses, not on differences between children, but on differences between words.
-These studies use regression models to predict which *words* are easier or more difficult, averaging across children [e.g., @goodman2008does; @braginsky2019]. 
-They typically show a strong association between word frequency and the age at which children typically acquire particular words, especially for object labels. 
+Reasoning from first principles, you cannot learn the word "table" if you do not hear it: input must predict learning. 
+Yet correlational studies do not assess this relation fully, and instead assess whether variation in input relates to variation in learning. 
+Thus, a second line of work focuses on differences between words, not children. 
+These studies use regression models to predict which words are easier or more difficult, averaging across children [e.g., @goodman2008does; @braginsky2019]. 
+They typically show a strong association between word frequency and the age at which children acquire particular words on average, especially for object labels. 
 This general finding provides convergent support for accumulator models, absent the confound of differences between individuals.
-
-<!--
-More broadly, this literature illustrates the value of a quantitative process model: under the standard model and its variants, we should be able to compute the expected relation between measures of input and vocabulary size.
-To the extent this model accounts for the data, it will support researchers’ current theory, but the model’s pattern of misfit to the data will likely be more informative, providing grist for the twin mills of theory development and empirical study.
--->
+In sum, accumulator models provide the conceptual underpinning of the relationships between input (freuqency for words, quantity for children) and learning.
 
@@
 IRT is commonly used for estimating the ability of "test takers" (in our case language learners) as they are assessed with a particular set of "test items" (particular words). 
-IRT models jointly infer both the difficulty of specific test items and the accuracy of specific test takers, recovering these latent parameters from the observed data. 
+IRT models provide a convenient and broadly-used framework within which to describe and compare different model variants, which in turn represent different sets of theoretical assumptions.
+
+In their formal structure, IRT models describe individual item responses as a function of both the difficulty of specific words and the language abilities of individual children. 
+These latent parameters can be inferred from an observed dataset. 
 In the basic Rasch (or 1-parameter logistic) IRT model, a person $i$ responds correctly to item $j$ with probability determined by their ability ($\theta_i$) and the difficulty of item $j$ ($d_j$):
@@
 With this mapping into absolute units, this kind of model can provide a quantitative linking hypothesis between measurements of input and learning.
-Recognizing the parallels between the standard model desiderata and IRT theory, we can use standard extensions of the 1PL (Rasch) model to include both item-level and person-level covariates. 
-Item-level (word-level) covariates of central interest are estimates of word frequency in child-directed speech, as well as lexical class [e.g., noun, predicate, or function word; @goodman2008does; @braginsky2019].
-Person-level covariates of interest might include sex, socioeconomic status variables, or other demographic information.
+
+We can then use standard extensions of IRT models to explore different assumptions.
+For example, we can include both item-level covariates -- e.g., estimates of word frequency or lexical class [@braginsky2019] -- and person-level covariates -- e.g., sex, socioeconomic status variables, or other demographic information.
+We can also consider whether language ability is uni-dimensionality or whether a multi-factor model is necessary [@frank2021]. 
+A final benefit of the IRT framework is that we can use a standardized set of tools for comparison of models on their fit to data, a major benefit over more ad-hoc frameworks.
 
@@
 
-Using the well-known IRT framework to formalize theoretical questions and assess them via model comparison is a benefit over more ad-hoc frameworks. 
-We give one example here. 
+We next give an example of how IRT accumulator models can be fit to data and hence how different theoretical assumptions can be compared empirically in this framework. 
 We can use this framework to test whether the amount of language experience is the important predictor of learning for a child or whether there are other age-related factors beyond experience. 
@@
 
-The goal of a computational theory like the "standard model" is to describe the quantitative predictions that arise from a set of assumptions. 
+The goal of any computational theory is to derive the predictions arising from a specific set of assumptions. 
 Often, however, it is the *failure* of such a model to predict observed patterns of data that is most useful, as these failures point the way forward towards future refinements [@vanrooij2020theory]. 
-We discuss three potential failures here.
+We discuss three potential failures of standard accumulator models like those described above.
 
-**Leveraged learning and the role of processing**. As the model results show, older children accumulate language from their input faster than younger children. 
+**Leveraged learning and the role of processing**. As our results show, older children accumulate language from their input faster than younger children. 
 There are at least three non-mutually exclusive potential reasons for this.
 First, older children may be better at remembering their experiences due to independent developmental/maturational factors. 
-Another possibility is that children *leverage* their knowledge of language to learn faster through any of several proposed mechanisms [@mitchell2009]. 
+Second, children may *leverage* their knowledge of language to learn faster through any of several proposed mechanisms [@mitchell2009]. 
 For example, if children reason about new words by exclusion [@markman1988children], they get better at learning new words just by having more potential items to exclude from consideration. 
@@
 
-**Beyond vocabulary**. The "standard model" proposed here models the accumulation of words.
-Yet, of course, language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings. 
+**Beyond vocabulary**. 
+The "standard model" proposed here models the accumulation of words.
+Yet language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings. 
 In early views, syntax and morphology were conceptualized as distinct and unconnected, but this conception has not been borne out empirically. 
@@
 The more experiences, the faster the words are learned. 
-This implicit theory can be expressed as a simple computational model that makes quantitative predictions. 
-By situating this model in a common psychometric framework, we show how it can be used to connect to large-scale, cross-linguistic datasets. 
-This "standard model" synthesizes measures of language input and vocabulary growth, allowing us to formalize, test, and iteratively improve our theoretical understanding of early language acquisition.
+This implicit theory can be expressed within a family of computational models that make quantitative predictions. 
+By situating these models in a common psychometric framework, we show how they can be used to connect to large-scale, cross-linguistic datasets. 
+This modeling framework synthesizes measures of language input and vocabulary growth, allowing us to formalize, test, and iteratively improve our understanding.
 Moreover, this formalization allows us to identify specific gaps in current empirical approaches that must be closed to inform future theory development.
+Perhaps one day soon, these developments will lead to a true "standard model" of language learning.
 

### 2021-03-19 paper/paper.Rmd
full pass

@@
 We can count how many words a child hears and express this number as a rate [e.g., words per hour; @cristia2020accuracy]. 
-We can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size, @wordbankbook]. 
+We can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size; @frank2021]. 
 These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
@@
 
-We next give an example of how IRT accumulator models can be fit to data and hence how different theoretical assumptions can be compared empirically in this framework. 
-We can use this framework to test whether the amount of language experience is the important predictor of learning for a child or whether there are other age-related factors beyond experience. 
-Put simply, is a 2-year-old different than a 1-year-old, aside from having twice as much experience with each word? 
-This theoretical question can be answered by simple model comparison between two related models.
+We next give an example of how IRT accumulator models can be fit to data and hence how different theoretical assumptions can be compared empirically in the framework described above. 
+We ask, is a 2-year-old different than a 1-year-old, aside from having twice as much experience with each word? 
+That is, is the amount of language experience the only developmental change we predict (as assumed by the simplest version of an accumulator model)? 
+Or are there other age-related changes that distinguish younger and older children's learning?
 
-In Model 1, we included age as a person-level covariate, allowing for interactions of age, lexical class, and monthly tokens per word. 
-Model 2 dropped the age covariate and instead scaled the expected tokens/month for each word by child age, yielding an estimate of lifetime tokens per word and also allowed for interactions with lexical class.
+To answer this question, we fit IRT models to parent report data about the words that children produce. 
+We downloaded data from 5,429 monolingual, English-learning children between 16 and 30 months from Wordbank [@frank2021]. 
+These data were collected using the MacArthur-Bates Communicative Development Inventory (CDI), a reliable and valid parent report instrument; they include 680 binary responses per child about whether a particular child produced a particular word at a given age.
+Each model estimated parameters for each child's linguistic ability and each word's difficulty.
 
-The first includes age is a person-level covariate, indicating 
-The second includes age as a scalar multiplier on the prevalence of each item (i.e., the frequency of "dog" for a 2-year-old is twice that of a 1-year-old).
+We also included the frequency of each word as the key parameter controlling its accumulation, treating frequency as an item-level covariate of difficulty.
+We do not have access to word frequencies in individual children's environments, however. 
+Instead, we estimated average word frequencies by using the American English corpora in CHILDES [@macwhinney2000childes], following previous work [@goodman2008does;@braginsky2019].
+<!-- IS THIS ^^ CORRECT? ALL AMERICAN ENGLISH? WHAT AGE CUTOFF? CHILDES-DB? -->
+We estiamted that  an average child receives 1,200 words/hour, 12 hours/day, for a total of 438,300 tokens/month, of which 285,200 tokens (65.1%) are tokens of words appearing on the CDI. 
+<!-- HOW WAS THIS ^^ COMPUTED! NEED AN APPENDIX OR NOTES IN THE REPO OR SOMETHING -->
+We then included frequency estimates for each word of the tokens per month heard by a typical child (range: 0.2 - 19,286 tokens).
 
-We report the results of this model comparison, fit to parent-reported vocabulary data from 5,492 English-speaking children in Wordbank. 
-Each model estimated parameters for each child’s ability, and for the 680 words on the CDI Words and Sentences form [@fenson2007]. 
-Covariates were each word's lexical class (function word, adjective, noun, verb, or other) and the estimated number of monthly tokens a child hears (range: 0.2 - 19,286 tokens), based on child-directed speech from CHILDES [@macwhinney2000childes].
-This model expects an average child to receive 1,200 words/hour, 12 hours/day, for a total of 438,300 tokens/month, of which 285,200 tokens (65.1%) are  tokens of CDI words.
+In our simple accumulator model, we used as our covariate for each child the  expected total tokens they had heard given their age. 
+We also included an interaction between frequency and lexical category, as previous work has suggested that frequency is more predictive for nouns than for other word classes [@goodman2008does;@braginsky2019].
+In contrast, in our more complex model, we included age as a person-level covariate, allowing for interactions between the child's age and the frequency and lexical class predictors.
 
-Model selection metrics (AIC and BIC) preferred the more complex Model 1. 
-Predicted acquisition curves from Model 1 are shown in Figure 3 by age, lexical class, and prevalence (left), revealing that nouns are learned more rapidly than verbs or function words, and that there are interactions of prevalence and word type (e.g., more frequent nouns are learned earlier, but there is little effect of prevalence for verbs and function words).
+The more complex model showed a better fit to the data and were preferred over under several model selection metrics (AIC and BIC).
+Predicted acquisition curves from this model are shown in Figure 3 by age, lexical class, and prevalence (left), revealing that nouns are learned more rapidly than verbs or function words, and that more frequent nouns are learned earlier, but there is little effect of frequency for verbs and function words.
 Moreover, the model predicts item-level acquisition curves (Figure 3, right): e.g., "ball" is learned earlier than "dog" and "go" is easier than "have." 
-Model 1 represents a new baseline standard model, which might be extended to include person-level groups such as SES and sex, as well as item-level covariates such as word length.
-
-These models also estimate a latent language ability for each child, but currently do not specify whether/how this value changes over time. 
-However, this ability is no doubt correlated with age, as well as with features related to input quality. 
-Estimating the contributions of these factors and determining whether/how language ability changes over developmental time is not possible with current datasets.
-Thus, we conclude by discussing the next steps needed to test and build upon the standard model.
+Further, its better fit provides support for age-related changes in the accumulation mechanism. 
+More generally, this simulation demonstrates the possibility of using large-scale datasets to test hypotheses about the nature of learning mechanisms.
 
@@
 
-# Onward from the Standard Model
+# Onward Towards a Standard Model
 
@@
 Often, however, it is the *failure* of such a model to predict observed patterns of data that is most useful, as these failures point the way forward towards future refinements [@vanrooij2020theory]. 
-We discuss three potential failures of standard accumulator models like those described above.
+We discuss three potential failures of simple accumulator models.
 
-**Leveraged learning and the role of processing**. As our results show, older children accumulate language from their input faster than younger children. 
-There are at least three non-mutually exclusive potential reasons for this.
-First, older children may be better at remembering their experiences due to independent developmental/maturational factors. 
-Second, children may *leverage* their knowledge of language to learn faster through any of several proposed mechanisms [@mitchell2009]. 
-For example, if children reason about new words by exclusion [@markman1988children], they get better at learning new words just by having more potential items to exclude from consideration. 
-A third possibility is that children are getting better at using the words they know.
-Several recent studies found that a surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which individual children are able to process familiar words [e.g., @fernald2006picking; @marchman2016early].
-Disentangling these distinct explanations will require careful measurement of both vocabulary and the potential mechanisms of leverage [cf. @lewis2020role].
+**Leveraged learning and the role of processing**. 
+As our results above show, older children accumulate language from their input faster than younger children.
+Why?
+Are older children simply better at remembering words, leading to "bigger drops" in the bucket for older children? 
+Or children *leverage* their knowledge of language to learn faster [@mitchell2009]? 
+They could do this by reasoning about new words by exclusion [@markman1988children]. 
+Or their increasing fluency with the words they know could help them learn new words.
+A surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which they process familiar words [e.g., @fernald2006picking; @marchman2016early].
+These theories predictions about which words should faciliate which others predictions that could be tested using the models we describe.
+
+**A theory for understanding acquisition in diverse contexts**. 
+The framework above formalizes an implicit assumption: namely, children learn the same way in all circumstances.
+Yet this assumption could very well be false. 
+For example, some children might learn more from overheard speech and others might learn more from child-directed speech [@sperry2019]. 
+The kind of data necessary to test this assumption directly are only now being collected, for example, in studies that rigorously track learning outcomes and amount of language input in diverse populations [e.g., comparisons between children in low-income, rural, indigenous communities and those in higher-income Western contexts; @casillas2020].
 
 **Beyond vocabulary**. 
-The "standard model" proposed here models the accumulation of words.
+We have described a model of the accumulation of words.
 Yet language is a rich, complex system in which words are inflected morphologically and composed syntactically to express compositional meanings. 
 In early views, syntax and morphology were conceptualized as distinct and unconnected, but this conception has not been borne out empirically. 
-Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language; @wordbankbook].
+Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language; @frank2021].
 Theoretically, accumulator models like the standard model are generic models of skill acquisition. 
 If language learning is a form of skill acquisition [@chater2018language], such connections could lead the way towards extensions of the standard model to the accumulation of broader units of language-like constructions.
-
-**A theory for understanding acquisition in diverse contexts**. 
-The "standard model" formalizes an implict assumption: namely, that early language learning follows the same general model for all children, even those in radically different circumstances. 
-Yet this assumption could very well be false. 
-Some children might learn more from overheard speech and others might learn more from child-directed speech [@sperry2019]. 
-The kind of data necessary to test this assumption directly are only now being collected, for example, in studies that rigorously track learning outcomes and amount of language input in diverse populations [e.g., comparisons between children in low-income, rural, indigenous communities and those in higher-income Western contexts; @casillas2020].
 

### 2021-03-19 paper/paper.Rmd
more tweaks

@@
 For example, @mcmurray2007 elegantly demonstrated that children’s "vocabulary explosion" -- an acceleration in vocabulary growth in the second year -- can result from the steady accumulation of word tokens without changes in the environment or learning mechanism.
-Other work has examined similar models using more realistic distributions of words (i.e., Zipf-distributed), developmental change in learning mechanisms, and quantitative comparison to children’s aggregate vocabulary growth [@mayor2010; @hidaka2013; @mollica2017].
+Other work has examined similar models using more realistic distributions of words, developmental change in learning mechanisms, and comparison to children’s aggregate vocabulary growth [@mitchell2009; @hidaka2013; @mollica2017].
 Although this work is exciting, these models have not yet made direct predictions about learning in individual children [@bergelson2020comprehension].
@@
 We can similarly estimate how many words they know [e.g., by parents' reports of vocabulary size; @frank2021]. 
-These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations [@dailey2021language]. 
+These absolute units give the potential for a model to make powerfully general quantitative predictions, which could be tested across different situations and populations. 
 Few datasets have measurements of the relevant variables, however. 
@@
 Since seminal work by @hart1995, the connection between children’s language input and the growth of vocabulary has been a topic of intense interest. 
-Numerous studies have reported positive associations between these variables [e.g., @hoff2003specificity; @weisleder2013] -- as predicted by accumulator models -- though their magnitude varies across studies [@wang2020meta].
-Such correlations are also partially moderated by other factors, including socioeconomic [@hoff2003specificity; @dailey2021language] and genetic [@hayiou2014language] variables.
+Numerous studies have reported positive associations between these variables [e.g., @hoff2003specificity] -- as predicted by accumulator models -- though their magnitude varies across studies [@wang2020meta].
+Such correlations are also partially moderated by other factors, including socioeconomic [@hoff2003specificity] and genetic [@hayiou2014language] variables.
 Randomized interventions are the gold standard for measuring causal effects of parental language input on language learning, although such studies are costly and difficult to conduct. 
@@
 To answer this question, we fit IRT models to parent report data about the words that children produce.
+In the simple setting we describe, these models can be written as multi-level logistic regression models [@de-boeck2011], albeit with a clearer theoretical interpretation than typical regression models.
 We downloaded data from 5,429 monolingual, English-learning children between 16 and 30 months from Wordbank [@frank2021]. 
@@
 
-```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves from a fitted English IRT model as a function of lexical class and the number of expected tokens per month (left), and for a sample of the 680 words on the CDI (right)."), echo=FALSE}
+```{r, out.width="1.0\\linewidth", include=TRUE, fig.align="center", fig.cap=c("Predicted acquisition curves derived from a fitted age-dependent accumulator model (see text) as a function of lexical class and the number of expected tokens per month (left), and for a sample of specific (right)."), echo=FALSE}
 knitr::include_graphics("./figs/fig3_predicted_age_LC_and_items.pdf")
@@
 Or their increasing fluency with the words they know could help them learn new words.
-A surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which they process familiar words [e.g., @fernald2006picking; @marchman2016early].
+A surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which they process familiar words [e.g., @marchman2016early].
 These theories predictions about which words should faciliate which others predictions that could be tested using the models we describe.
@@
 In early views, syntax and morphology were conceptualized as distinct and unconnected, but this conception has not been borne out empirically. 
-Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@bornstein2013language; @frank2021].
+Instead, evidence shows again and again that the language system is "tightly woven," with extremely tight correlations between the acquisition of words, morphology, and syntax [@frank2021]. 
 Theoretically, accumulator models like the standard model are generic models of skill acquisition. 

### 2021-03-19 paper/paper.Rmd
edits

@@
 keywords          : "early language learning; language acquisition; vocabulary development"  
-wordcount         : "2896"
+wordcount         : "2921"
 
@@
 
-Reasoning from first principles, you cannot learn the word "table" if you do not hear it: input must predict learning. 
-Yet correlational studies do not assess this relation fully, and instead assess whether variation in input relates to variation in learning. 
-Thus, a second line of work focuses on differences between words, not children. 
+Reasoning from first principles, you cannot learn the word "table" if you do not hear it: input must predict learning of individual words. 
+Yet correlational studies do not assess this word-level relation fully, and instead assess whether variation in input relates to variation in learning across children. 
+However, a second line of work focuses on differences between words, not children. 
 These studies use regression models to predict which words are easier or more difficult, averaging across children [e.g., @goodman2008does; @braginsky2019]. 
@@
 This general finding provides convergent support for accumulator models, absent the confound of differences between individuals.
-In sum, accumulator models provide the conceptual underpinning of the relationships between input (freuqency for words, quantity for children) and learning.
+In sum, accumulator models provide the conceptual underpinning of the relationships between input (frequency for words, quantity for children) and learning.
 
@@
 However, both children and words vary: children may learn faster or slower, and words can be more or less difficult to learn. 
-Combining these ideas, the basic hypothesis is that a child’s vocabulary at a particular time should be predicted by their cumulative language exposure and learning rate, combined with the breadth of the sample of words to which have been exposed and those words' individual difficulties.
+Combining these ideas, the basic hypothesis is that a child’s vocabulary at a particular time should be predicted by their cumulative language exposure and learning rate, combined with the breadth of the sample of words to which they have been exposed and those words' individual difficulties.
 
@@
 We do not have access to word frequencies in individual children's environments, however. 
-Instead, we estimated average word frequencies by using the American English corpora in CHILDES [@macwhinney2000childes], following previous work [@goodman2008does;@braginsky2019].
-<!-- IS THIS ^^ CORRECT? ALL AMERICAN ENGLISH? WHAT AGE CUTOFF? CHILDES-DB? -->
-We estiamted that  an average child receives 1,200 words/hour, 12 hours/day, for a total of 438,300 tokens/month, of which 285,200 tokens (65.1%) are tokens of words appearing on the CDI. 
+Instead, we estimated average word frequencies using the American English corpora in CHILDES [@macwhinney2000childes] retrieved from childes-db [@sanchez2019childes], following previous work [@goodman2008does;@braginsky2019].
+<!-- need to specify age cut-off? ToDo GK - believe it was all data -->
+We estimated that an average child receives 1,200 words/hour, 12 hours/day, for a total of 438,300 tokens/month, of which 285,200 tokens (65.1%) are tokens of words appearing on the CDI (see Appendix).
 <!-- HOW WAS THIS ^^ COMPUTED! NEED AN APPENDIX OR NOTES IN THE REPO OR SOMETHING -->
@@
 Why?
-Are older children simply better at remembering words, leading to "bigger drops" in the bucket for older children? 
-Or children *leverage* their knowledge of language to learn faster [@mitchell2009]? 
+Are older children simply better at remembering words, leading to "bigger drops" in their buckets? 
+Or do older children *leverage* their knowledge of language to learn faster than younger children [@mitchell2009]? 
 They could do this by reasoning about new words by exclusion [@markman1988children]. 
-Or their increasing fluency with the words they know could help them learn new words.
+Or could their increasing fluency with the words they know help them learn new words? 
 A surprising proportion of variance in the rate of young children’s vocabulary growth is accounted for by the speed with which they process familiar words [e.g., @marchman2016early].
-These theories predictions about which words should faciliate which others predictions that could be tested using the models we describe.
+These theoretical proposals about which words should facilitate the learning of other words yield predictions that could be tested using the models we describe.
 

### 2021-08-13 model_interventions.Rmd
added refs, cleaned up model code a bit
looking at book vs. CHILDES vs. SUBTLEX frequencies for CDI/non-CDI words

@@
 ```{r, summary-tables, echo=F}
+# Table 1 O’Reilly, d’Apice, von Stumm
+# Adult word count N=107 Mean=54065.00 SD=17135.33 Min=15038.02 Max=99684.85 Skew=0.37 Kurtosis=0.05
+
 # Gilkerson et al. 2017 Table 2: standardized AWC
@@
 
+# log-normal distribution?
+require(fitdistrplus)
+mc_tok = df$mother_child_tokens_hr[!is.na(df$mother_child_tokens_hr)]
+ac_tok = df$adult_child_tokens_hr[!is.na(df$adult_child_tokens_hr)]
+all_tok = df$all_tokens_hr[!is.na(df$all_tokens_hr)]
+plotdist(mc_tok, histo=T, demp=T) # 113
+plotdist(ac_tok, histo=T, demp=T) # 42
+plotdist(all_tok, histo=T, demp=T) # 83
+
+# based on skew vs. kurtosis, look like beta distributions
+descdist(df$mother_child_tokens_hr[!is.na(df$mother_child_tokens_hr)], boot=1000)
+descdist(df$adult_child_tokens_hr[!is.na(df$adult_child_tokens_hr)], boot=1000)
+descdist(df$all_tokens_hr[!is.na(df$all_tokens_hr)], boot=1000)
+
+fb <- fitdist(mc_tok/max(mc_tok+1e-9), method="mme", "beta") # beta: vals must be in [0-1], so I scaled...
+# llik = -7.76
+
+summary(fitdist(mc_tok/max(mc_tok+1e-9), "weibull")) # llik: 23.4 ?
+
+fw <- fitdist(mc_tok, "weibull") # llik: -902
+fg <- fitdist(mc_tok, "gamma", method="mme") # llik: -902 fails with method="mle" error code 100
+fln <- fitdist(mc_tok, "lnorm") # llik: -907.9991
+fn <- fitdist(mc_tok, "norm") # llik: -920.7385
+plot.legend <- c("Weibull","log-normal", "gamma")
+par(mfrow=c(2,2))
+denscomp(list(fw, fln, fg), legendtext=plot.legend)
+qqcomp(list(fw, fln, fg), legendtext=plot.legend)
+cdfcomp(list(fw, fln, fg), legendtext=plot.legend)
+ppcomp(list(fw, fln, fg), legendtext=plot.legend)
+
 # should this be a Poisson distribution? not normal

### 2021-08-13 app.R

@@ -9,7 +9,7 @@ library(shinythemes)
 #source("model-nonsampling.R")
 source("model-childesfreqs.R")
 
-theme_set(theme_classic())
+theme_set(theme_classic(base_size=16))
 # Define UI
 ui <- fluidPage(
     theme = shinytheme("spacelab"),

### 2021-08-13 word_freqs_book_vs_childes.R

@@ -1,0 +1,213 @@
+require(tidyverse)
+require(ggrepel)
+require(here)
+
+# SUBTLEX-US subtitle word frequencies (adult speech standard)
+subf <- read_csv(here("data/SUBTLEX-US.csv")) %>%
+  rename(word_count_norm = SUBTLWF, # frequency per million words
+         word_count = SLUSfreq) %>% 
+  mutate(source = "SUBTLEX",
+         prob = word_count / sum(word_count)) %>%
+  arrange(desc(word_count_norm)) %>%
+  mutate(rank = 1:n()) %>%
+  select(word, word_count, word_count_norm, prob, source, rank)
+# hapaxes already removed  
+
+# CHILDES corpus (hapaxes already removed, but 6295 word_count==2)
+chf <- read_csv(here("data/childes_english_word_freq_cleaned_noHapaxes.csv")) %>% select(-X1) %>%
+  mutate(source = "CHILDES") %>% 
+  arrange(desc(word_count)) %>%
+  mutate(rank = 1:n())
+
+# 165 UK picture book corpus: https://osf.io/b2qd8/
+# from: Dawson, N., Hsiao, Y., Tan, A.W.M., Banerji, N., & Nation, K. (2021). 
+# Features of lexical richness in children’s books: Comparisons with child-directed speech. Language Development Research.
+bbf <- read_csv(here("alvin-books/picturebook_ref.csv")) %>%
+  filter(PoS_type=="lexical") %>%  # just use 9074 lexical items? (2360 non-lexical excluded)
+  arrange(desc(n)) %>%
+  rename(word_count = n,
+         word = lemma) %>%
+  mutate(rank = 1:n()) 
+# only 5550 unique words (unique are lemma x PoS combos)
+
+# Dawson2021 uses keyness scores (ratio of normalized frequency in focus corpus to norm. freq in reference corpus),
+# with +10 to each words frequency 
+
+# book corpus (Montag et al 2016)
+bf <- read.csv(here("book-corpus/100out.txt"), sep=' ') %>% 
+  mutate(word = ifelse(word=="i", "I", word),
+         word_count_norm = word_count * (1e6 / sum(word_count)), # count per million tokens
+         prob = word_count / sum(word_count),
+         source = "Books"
+  ) %>%
+  arrange(desc(word_count)) %>%
[…]

### 2021-09-24 childrens-input.Rmd
estimating total language input

---
title: "Children's language input"
author: "George"
date: "9/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

# Goal

Use average hourly input rates from three sources (overheard speech, child-directed speech, and child-read books) to estimate how many tokens a 'standard' child will experience over their early development.

## Overheard and Child-directed speech

Using children's median number of waking hours, we extrapolate from our best estimates of hourly rates for overheard speech (600 words/hr) and child-directed speech (1198 words/hr) to the monthly amounts of speech heard by children.
(Should these rates change for school-aged children?)

```{r waking-hours}
# https://www.sleepfoundation.org/children-and-sleep/how-much-sleep-do-kids-need
waking_hours_per_day = c(7,8,9, # 14-17 hrs of sleep
                           seq(10, 12, length.out=9), # infant: 12-15 hrs of sleep
                           rep(12, 12), # toddler: 11-14 hrs of sleep
                           rep(13, 3*12), # preschool (3-5): 10-13 hrs of sleep
                           rep(14, 8*12)) # school-aged: 9-11 hrs of sleep
```

```{r hourly-rates}
overheard_speech_rate = rep(600, 13*12) # 13 years - increase at school age?
child_directed_speech_rate = rep(1198, 13*12) # how should this change?
# let's estimate monthly rates based on Hasbrauck and Tindal (2006) norms
reading_rate = 60 * c(rep(0, 12*5), # no reading 0-5
                 seq(0, 53, length.out=12), # 53 correct wpm by end of Grade 1
                 seq(53, 89, length.out=12), # 89 by end of Grade 2
                 seq(89, 107, length.out=12), # 107 by end of Grade 3
                 seq(107, 123, length.out=12), # Grade 4
                 seq(123, 139, length.out=12), # Grade 5
                 seq(139, 150, length.out=12), # Grade 6
                 seq(150, 150, length.out=12), # Grade 7
                 seq(150, 151, length.out=12) # Grade 8
                 ) # first 13 years

# 6TH-8TH GRADE (SPRING) (11-14 YEARS OLD): 150 – 204 WPM
# HIGHSCHOOL (14-18 YEARS OLD): 200 – 300 WPM
# COLLEGE (18-23) YEARS OLD	300 – 350 WPM
# ADULTS	220 – 350 WPM
```

## Reading rates

Many sources recommend that school-aged children read 15-20 minutes per day (7.6 hours/month), although most teachers estimate that most of their students read less than this.
For now we will assume the 'standard' child reads the recommended daily amount, at speeds following the Hasbrauck and Tindal (2006) norms (e.g., 53 correct words/min by end of Grade 1; 151 correct words/min by end of Grade 8).
For reference, silent-reading adults average 238 words/min, while adults that read aloud average 183 words/min (Brysbaert, 2019).

```{r, echo=FALSE}
hours_read_per_month = 30.44*.25 # 15 min/day
tokens_read_per_month = reading_rate * hours_read_per_month
waking_hours_per_month = waking_hours_per_day * 30.44

hours_overheard_speech_per_month = overheard_speech_rate * waking_hours_per_month
hours_child_directed_speech_per_month = child_directed_speech_rate * waking_hours_per_month

input_long_df <- bind_rows(
  tibble(Source = "Reading", Age = 1:length(tokens_read_per_month), 
         `Hourly Rate` = reading_rate,
         `Monthly Tokens` = tokens_read_per_month,
         `Total Tokens` = cumsum(tokens_read_per_month)),
  tibble(Source = "Overheard Speech", Age = 1:length(tokens_read_per_month),
         `Hourly Rate` = overheard_speech_rate,
         `Monthly Tokens` = hours_overheard_speech_per_month,
         `Total Tokens` = cumsum(hours_overheard_speech_per_month)),
  tibble(Source = "Child-directed Speech", Age = 1:length(tokens_read_per_month),
         `Hourly Rate` = child_directed_speech_rate,
         `Monthly Tokens` = hours_child_directed_speech_per_month,
         `Total Tokens` = cumsum(hours_child_directed_speech_per_month))
  )
```

## Monthly input

For reference, the average novel is ~60,000 words.

```{r monthly-input, echo=F}
input_long_df %>%
  mutate(`Age (years)` = Age / 12) %>%
  ggplot(aes(x=`Age (years)`, y=`Monthly Tokens`, fill=Source)) + 
    geom_area() + #ylim(0, 1.5e6) + 
    theme_classic() # + geom_hline(yintercept=1e6, linetype="dashed")
```

## Cumulative input

How many tokens do children accumulate from different sources by the age of 13?

```{r, echo=F}
input_long_df %>% 
  mutate(`Cumulative Tokens (millions)` = `Total Tokens` / 1e6,
         `Age (years)` = Age / 12) %>%
  ggplot(aes(x=`Age (years)`, y=`Cumulative Tokens (millions)`, fill=Source)) + 
    geom_area() + 
    theme_classic() + # + geom_hline(yintercept=1e6, linetype="dashed")
    theme(legend.position = c(.2, .7))
```

(Add TV? "The average across all 170 stations is 2.38 words per second" (143 words/min, 8580 words/hr). The average child (aged 8-12) watches 4-6 hours/day -- 34320 to 51480 words that are not overheard/child-directed speech.)

### 2021-11-02 childrens-input.Rmd
working on standard model v2

@@
 hours_read_per_month = 30.44*.25 # 15 min/day
+
+get_input_df <- function(overheard_speech_rate, child_directed_speech_rate, 
+                         reading_rate, hours_read_per_month, waking_hours_per_day) {
+  hours_overheard_speech_per_month = overheard_speech_rate * waking_hours_per_month
+  hours_child_directed_speech_per_month = child_directed_speech_rate * waking_hours_per_month
+
   tokens_read_per_month = reading_rate * hours_read_per_month
   waking_hours_per_month = waking_hours_per_day * 30.44
-
-hours_overheard_speech_per_month = overheard_speech_rate * waking_hours_per_month
-hours_child_directed_speech_per_month = child_directed_speech_rate * waking_hours_per_month
 
@@
            `Hourly Rate` = reading_rate,
+           `Monthly Hours` = hours_read_per_month,
            `Monthly Tokens` = tokens_read_per_month,
@@
            `Hourly Rate` = overheard_speech_rate,
+           `Monthly Hours` = waking_hours_per_month,
            `Monthly Tokens` = hours_overheard_speech_per_month,
@@
            `Hourly Rate` = child_directed_speech_rate,
+           `Monthly Hours` = waking_hours_per_month,
            `Monthly Tokens` = hours_child_directed_speech_per_month,
@@
     )
+  return(input_long_df)
+}
+
+input_long_df <- get_input_df(overheard_speech_rate, 
+                              child_directed_speech_rate, 
+                              reading_rate, 
+                              hours_read_per_month, 
+                              waking_hours_per_day)
 ```
@@
 
-(Add TV?
-"The average across all 170 stations is 2.38 words per second" (143 words/min, 8580 words/hr).
-The average child (aged 8-12) watches 4-6 hours/day -- 34320 to 51480 words that are not overheard/child-directed speech.)
+```{r}
+age13 = subset(input_long_df, Age==156)
+age13[1,]$`Total Tokens` / sum(age13$`Total Tokens`)
+```
+
+## Strong Reader
+
+What if a child reads an hour per day, and is an accelerated reader (starts reading at age 5, and reads at norms 1 grade level ahead)?
+
+```{r}
+reading_rate_accel = 60 * c(rep(0, 12*4), # no reading 0-4
+                 seq(0, 53, length.out=12), # 53 correct wpm by end of kindergarten
+                 seq(53, 89, length.out=12), # 89 by end of Grade 1
+                 seq(89, 107, length.out=12), # 107 by end of Grade 2
+                 seq(107, 123, length.out=12), # Grade 3
+                 seq(123, 139, length.out=12), # Grade 4
+                 seq(139, 150, length.out=12), # Grade 5
+                 seq(150, 150, length.out=12), # Grade 6
+                 seq(150, 151, length.out=12), # Grade 7
+                 seq(151, 204, length.out=12)
+                 ) # first 13 years
+
+# 6TH-8TH GRADE (SPRING) (11-14 YEARS OLD): 150 – 204 WPM
+# HIGHSCHOOL (14-18 YEARS OLD): 200 – 300 WPM
+# COLLEGE (18-23) YEARS OLD	300 – 350 WPM
+# ADULTS	220 – 350 WPM
+
+input_long_df_accel <- get_input_df(overheard_speech_rate, 
+                              child_directed_speech_rate, 
+                              reading_rate_accel, 
+                              30.44, 
+                              waking_hours_per_day)
+
+```
+
+```{r, echo=F}
+input_long_df_accel %>% 
+  mutate(`Cumulative Tokens (millions)` = `Total Tokens` / 1e6,
+         `Age (years)` = Age / 12) %>%
+  ggplot(aes(x=`Age (years)`, y=`Cumulative Tokens (millions)`, fill=Source)) + 
+    geom_area() + 
+    theme_classic() + # + geom_hline(yintercept=1e6, linetype="dashed")
+    theme(legend.position = c(.2, .7)) + 
+  ggtitle("An Avid Reader")
+```
+
+## TV Watching
+
+"The average across all 170 stations is 2.38 words per second" (143 words/min, 8568 words/hr). 
+Children ages 2 to 5 watched TV on average 32 hours each week (4.6 hrs/day), and children ages 6 to 8 spent 28 hours per week watching television (Nielson Co., 2009).
+The average child (aged 8-12) watches 4-6 hours/day, providing 34320 to 51480 daily words that are not overheard/child-directed speech.
+
+We show cumulative tokens for children watching the average amount of television for their age.
+We also assume TV time eats into CDS and ODS equally (i.e., for each hour of TV watched, children lose 30 minutes of CDS, and 30 minutes of ODS).
+We assume children younger than 3 do not watch any TV.
+
+```{r, echo=F}
+tv_hourly_tokens = 2.38 * 60 * 60
+
+hours_tv_per_day = c(rep(0, 2*12),
+                       rep(4.6, 3*12),
+                       rep(4.0, 2*12),
+                       rep(5.0, 6*12))
+
+tokens_tv_per_month = tv_hourly_tokens * hours_tv_per_day * 30.44
+
+# incorporate this into get_input_df function? or make adjust_for_activity() ?
+input_long_tv_df <- bind_rows(input_long_df,
+    tibble(Source = "TV", Age = 1:length(tokens_tv_per_month), 
+           `Hourly Rate` = tv_hourly_tokens,
+           `Monthly Hours` = hours_tv_per_day * 30.44,
+           `Monthly Tokens` = tokens_tv_per_month,
+           `Total Tokens` = cumsum(tokens_tv_per_month)))
+
+input_wide_tv_df <- input_long_tv_df %>% 
+  pivot_wider(id_cols=Age, names_from=Source, values_from=`Monthly Hours`) %>%
+  mutate(`Overheard Speech` = `Overheard Speech` - TV/2,
+         `Child-directed Speech` = `Child-directed Speech` - TV/2)
+
+input_wide_tv_df
+```
+
+```{r, echo=F}
+input_long_tv_df %>% 
+  mutate(`Cumulative Tokens (millions)` = `Total Tokens` / 1e6,
+         `Age (years)` = Age / 12) %>%
+  ggplot(aes(x=`Age (years)`, y=`Cumulative Tokens (millions)`, fill=Source)) + 
+    geom_area() + 
+    theme_classic() + # + geom_hline(yintercept=1e6, linetype="dashed")
+    theme(legend.position = c(.2, .7))
+```

### 2021-11-02 paper/paper.Rmd
working on standard model v2

@@
 ---
-title             : "Toward a \"Standard Model\" of Early Language Learning"
-shorttitle        : "Standard Model of Language Learning"
+title             : "A \"Standard Model\" of Early Language Learning Requires Absolute Measurement"
+shorttitle        : "Standard Model Measures"
 
@@
 abstract: |
-  A "standard model" is a theoretical framework that synthesizes observables into a quantitative consensus.
-Have we made progress towards this kind of synthesis for children’s early language learning?
+  A "standard model" which formally unites consensus verbal theories to make quantitative predictions is a critical step for improving theories of early language learning.
+Although early attempts at such a standard model have been made, their impact has been limited due to a lack of both 1) a
    A number of computational models of early vocabulary learning assume that individual words are learned through an accumulation of environmental input.
@@
 
-Early language learning is a key challenge for cognitive science: how do we go from speechless, wordless infants to children who can use language expressively and creatively? 
-The field of early language is often portrayed as mired in controversies, for example, around issues of innateness. 
-We see a new synthesis emerging, however, based on new theoretical and empirical work on the growth of vocabulary. 
-Our goal is to present this synthesis as the beginnings of a "standard model": a baseline theory that should be accepted widely in its outlines as a baseline summary[^1] of our beliefs and should guide future work, even if its assumptions still require rigorous evaluation.
-
-[^1]: "Standard" here is not meant to imply a normative prescription, but rather that the model is an accepted starting point for a description of early language learning.
-
-In physics, the "standard model" is a widely-accepted theory from the 1970s that describes all known elementary particles along with three of the four known fundamental forces in the universe. 
-Although it is known to be incomplete and even incorrect in places, physics’ Standard Model nonetheless explains a wide variety of empirical phenomena, allowing scientists and engineers to model physical interactions with great precision and accuracy and to design tests that inform theory revision.
 Psychology, in general, has been criticized for lacking such formal theories that inform and drive empirical research [@muthukrishna2019problem]. 
-We feel this critique can be applied fairly to the field of language acquisition.
-
-Although a substantial literature uses computational models to simulate acquisition, these models (ours included) are often only loosely connected to the quantitative data gathered by working acquisition researchers.
-Despite this lack of formal connections, researchers studying early language learning often presuppose a common framework for early language learning (Figure 1). 
-The core of this framework is that language input accumulates via repeated exposure, resulting in learning.
-This framework underpins much of the broader policy discussion of individual variability in language learning and its links to environmental stimulation.
+Cumulative theories that are iteratively refined,.
 
@@
 Our aim is to make an explicit connection between these models and the broader but less formal discussion of the role of language input in learning.
-Since discussions of language input are typically focused on variation between children, we use item-response theory (IRT) from psychometric testing as a framework for connecting empirical data with accumulator models.
-Sitting at a level of analysis between generic regressions and cognitive process models, data-analytic cognitive models of this type offer a way to implement our verbal theories and evaluate them quantitatively.
-Importantly, the goal is not to produce the "correct" model, but rather to explore the ways that model assumptions lead to predictions about specific patterns of data. 
-In fact, the greatest value of such models is the ability to identify the areas of greatest mismatch between data and model, highlighting areas of the theory requiring further investigation [@tauber2017bayesian].
-Perhaps such iterative work will lead to a true "standard model" of language learning.
+
+For an overview of best practices for collecting long format sensory experience (LFSE) recording, see @casillas2019step.
+
+## Absolute > Relative Measurements
+
+## Measurement Error
+
+accuracy, precision..
+If we only binary code child-directed / other-directed at the level of 5-minute segments, then even assuming that human coders are 100% accurate we have potential error of just less than 2.5 minutes per segment.
+
+Relative (Fractional) Uncertainty = | uncertainty / (measured quantity) |
+Propagation of uncertainty - do an example for all-day LENA recordings?
+<!-- https://www.webassign.net/question_assets/unccolphysmechl1/measurements/manual.html -->
+NIST says we should report total combined uncertainty for any measured values.
+NIST. Essentials of Expressing Measurement Uncertainty. http://physics.nist.gov/cuu/Uncertainty/
+Lichten, William. Data and Error Analysis., 2nd. ed. Prentice Hall: Upper Saddle River, NJ, 1999.
+
+## Reliability
+
+How reliable are human coders, e.g. at distinguishing child-directed from other-directed speech?
+How reliable are measures of children's speech environment (e.g., adult word count per hour), and how do they relate to the duration of the recording?
+
+A variety of reliability coefficients have been devised, which generally range from 0 to 1, and are meant to quantify the proportion of variation in a test score that is due to differences in the true scores of the test takers, and how much is due to measurement error or noise.
+
+Classical Test Theory / True Score Theory: an individual's observed score on a test is X = T + E, where T is their 'true' score and E is measurement error.
+Unfortunately, we only have access to X, and can only estimate the other values.
 
